ATG2A engages RAB1A and ARFGAP1 positive membranes during autophagosome biogenesis

Version 1

Abstract

Abstract

Autophagosomes form from seed membranes that expand through bulk-lipid transport via the bridge-like lipid transporter ATG2. The origins of the seed membranes and their relationship to the lipid transport machinery are poorly understood. Using proximity labeling and a variety of fluorescence microscopy techniques, we show that ATG2A localizes to extra-Golgi ARFGAP1 puncta during autophagosome biogenesis. ARFGAP1 itself is dispensable during macroautophagy, but among other proteins associating to these membranes, we find that RAB1 is essential. ATG2A co-immunoprecipitates strongly with RAB1A, and siRNA-mediated depletion of RAB1 blocks autophagy downstream of LC3B lipidation, similar to ATG2A depletion. Further, when either autophagosome formation or the early secretory pathway is perturbed, ARFGAP1 and RAB1A accumulate at ectopic locations with autophagic machinery. Our results suggest that ATG2A engages a RAB1A complex on select early secretory membranes at an early stage in autophagosome biogenesis.

Introduction

Macroautophagy is a cellular degradative pathway in which a seed vesicle grows to encapsulate a target cargo, closes to form the complete autophagosome, and subsequently fuses with the lysosome. This transient organelle forms de novo in approximately ten minutes and requires the delivery of millions of lipids1. ATG2 proteins function as bridge-like bulk lipid transport proteins2–5 and are essential during the membrane expansion step of autophagosome biogenesis. We propose that ATG2 bridges the ER and the developing phagophore, thereby allowing lipids to cross the cytosol through a long hydrophobic groove3. At the phagophore, lipids are equilibrated into both leaflets of the bilayer via the ATG9 scramblases6–8. As such, ATG9 is essential to macroautophagy and is a key component of the seed membranes that ultimately expand during ATG2-mediated lipid transport9, 10.

In addition to ATG9 vesicles, several other membrane sources have been proposed to be involved in autophagosome biogenesis. In particular, both genetics and biochemistry implicate an essential role for the early secretory pathway. In yeast, some COP-II vesicle and ER exit site (ERES) proteins contribute to starvation-mediated autophagy at a stage that is downstream of formation of the pre-autophagosome structure (PAS)11, 12. Furthermore, yeast ERES naturally colocalize with the autophagic cargo Ape113, and that colocalization increases when various autophagy factors are depleted13, 14. These studies imply that autophagosome initiation events may not be dependent upon the secretory pathway, but subsequent membrane growth or autophagosome maturation requires active ERES structures. Cross linking studies further establish a close proximity between autophagy and ERES proteins15. Finally, the direct integration of yeast COPII-derived membranes into developing phagophores is suggested by the detection of the COP-II cargo Axl2-GFP, a transmembrane protein, at both phagophores and mature autophagosomes16.

In human cell lines, genetic or pharmacological disruption of the COPII-related small GTPase SAR1A reduces the frequency of autophagic events17. Following biochemical fractionation of cellular membranes, the ER-to-Golgi Intermediate Compartment (ERGIC) is especially competent for supporting in vitro lipidation by LC3B, the primary marker of autophagic structures17, 18 and clusters of COPII positive membranes can be seen accumulating adjacent to phagophores19. Like in yeast, these studies can be interpreted as a key role for ERES or ERGIC membranes in forming or expanding autophagosome membranes. As starvation-mediated macroautophagy drives substantial remodeling of both the ERES and ERGIC membranes (Liang Ge review;), including the formation of subdomains with unique protein-dependent organelle-organelle contact sites20, it is possible that the ERES/ERGIC contribution to autophagy represents a distinct subcompartment of these membrane structures.

Presumably, any contribution of an early secretory membrane must ultimately coordinate with recruitment of ATG9 vesicles, in order to ensure that the lipid transport machinery is competent to expand the phagophore membrane. Several studies have postulated an early fusion intermediate in phagophore formation, such as the description of the HyPas compartment where vesicles bearing distinct autophagy proteins (ATG16L1 and FIP200) were observed to coalesce early in autophagosome biogenesis21. Intriguingly, when the lipid transporter ATG2 is genetically depleted from cells, all of the essential autophagy machinery including ATG9 concentrates into large compartments9, 22–24. Correlative light and electron microscopy of these structures indicate they are large assemblies of ATG9 vesicles surrounded by more complex membranes which harbor FIP200 and WIPI2, among other proteins9, 25. These complex membranes often adopt a cup-like shape, leading us to name them “proto-phagophores”9. The origins of the proto-phagophores have been unknown.

In this study, we demonstrate that ATG2A is enriched during autophagosome biogenesis at RAB1/ARFGAP1 positive membranes. Although both proteins can be found within the ERGIC compartment, proximity data and colocalization analyses in secretory pathway disrupted cells suggest that RAB1 and ARFGAP1 membranes involved in ATG2 recruitment are distinct from generic ERGIC membranes. Genetic depletion experiments indicate that RAB1 and ATG2 function at a similar step in autophagosome formation and further these proteins can be isolated together in a complex. Finally, both ARFGAP1 and RAB1 co-localize with proto-phagophores, implying these membranes can recruit much of the key early autophagy machinery and may be dynamically remodeled into cup-like morphologies.

Results

ATG2A is proximal to early secretory membranes

To determine the local proteome adjacent to ATG2A, we used a proximity labeling approach. We fused the biotinylation enzyme APEX2 and a green fluorescent marker to the amino-terminus of ATG2A (APEX2-EGFP-ATG2A) and expressed this construct in ATG2A/B double knockout cells (ATG2 DKO cells). In this way, the APEX2 enzyme will biotinylate cellular proteins in close proximity to the expressed ATG2A fusion protein26. To ensure that the APEX2-EGFP-ATG2A construct is active and correctly localized, we measured the restoration of macroautophagy in these cells by following LC3B flux and p62 clearance (Fig. S1 A-C). While both LC3B-II and p62 dramatically accumulate in the ATG2 DKO cells, this accumulation is almost entirely reversed upon re-expression of APEX2-EGFP-ATG2A.

We also validated that biotinylation in these cells depends upon the presence of APEX-EGFP-ATG2A and a full reaction. In cells expressing APEX2-EGFP-ATG2A, biotin is only observed when both the biotin substrate (biotin phenol) and the reaction promoter (hydrogen peroxide) are included (Fig. S1 D and E). Likewise, streptavidin-mediated immunoprecipitation revealed a strong enrichment of biotinylated material only when the full reaction conditions were employed (Fig. S1F). Furthermore, the anti-biotin and GFP signals colocalized establishing that biotinylation is favored at sites proximal to ATG2A (Fig. S1D).

Mass spectrometry of proteins isolated by streptavidin-mediated immunoprecipitation was performed using a three state quantitative SILAC experiment26 with light, medium, and heavy isotopes as follows: no hydrogen peroxide as a negative control, a full reaction in complete media, and a full reaction in starvation media and 100 nM bafilomycin. When comparing both the heavy and medium isotope states to the light state, mass spectrometry analysis revealed an abundance of autophagy related proteins in both full media and under starvation (Figs. 1A and S1G, respectively). The majority of these proteins were cargo adaptors such as NBR1, p62 (SQSTM1), and TAX1BP1, suggesting that the target cargo is immediately adjacent to the biogenesis machinery during autophagosome maturation. We also observed a clear enrichment of ERES or COPII proteins including SEC16A, SEC23A, SEC23IP, and SEC24B/C. Additional ERES proteins such as TFG and SEC31A were also enriched in some, but not all, of the three replicates. In addition, ARFGAP1, a COPI effector that is primarily localized to the Golgi but also found in earlier secretory membranes, was one of the two most enriched proteins in both full media and starvation. Both ARFGAP1 and the COPII proteins suggest ATG2 is localizing to the early secretory pathway.

ATG2A is proximal to ARFGAP1 positive membranes.

(A) Proximity labeling of HEK293 ATG2 DKO cells stably expressing APEX2-GFP-ATG2A was performed as previously described26. Three biological replicates were averaged and plotted as the log2 ratio between the full labeling reaction in complete media over cells that were not treated with hydrogen peroxide on the x axis, and -log10 of the p-value on the y axis. The red line demonstrates the p-value cutoff of 0.05. The proteomics revealed an enrichment of autophagy (labeled in green) and early secretory proteins (COPII/ER Exit Site proteins are labeled in orange, ARFGAP1 is labeled in red). (B) Strategy to identify phagophores by live cell imaging. Phagophores are defined as sites where both GFP-ATG2A and TagBFP-LC3B co-localize. (C-D) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2A and TagBFP-LC3B and transfected with mRuby-SEC23A or RFP-SEC23IP/p125a or ARFGAP1-TagRFP. The cells were starved for 2-4 hours prior to imaging. The fraction of phagophores (white arrows) that colocalized with the third fluorescent protein (transfected gene) was quantified and plotted in (D). Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled for each condition. Maximum intensity projections of confocal images. (E) Global organelle profiling data from https://organelles.sf.czbiohub.org31. Whole organelle profiles were established by pulling down endogenously tagged proteins and performing quantitative proteomics. The enrichment value for each protein were calculated by taking the difference in median log2 LFQ values between the replicates and the null distribution.

ATG2A is recruited to ARFGAP1 membranes during autophagosome biogenesis

As an orthogonal method to assess the proximity of ATG2A to early secretory components, we performed FLASH-PAINT, a super resolution technique with high multiplexing capabilities27. FLASH-PAINT offers sub-organelle resolution while simultaneously allowing us to screen >10 proteins. ATG2 DKO cells stably expressing GFP-ATG2A were starved and then immunostained for autophagic proteins (GFP-ATG2A, p62, and LC3B), ERES proteins (Tango1, SEC24C, SEC31A, SEC16A, TFG), and markers of COPI-positive membranes (ARFGAP1 and COPI). All proteins were concurrently labeled with primary antibodies that were in turn identified by their association with secondary antibodies with unique DNA barcodes (Fig. S1H). A fluorophore conjugated imager DNA strand and adaptor strands were added to target the imager to the correct secondary antibody. Eraser strands were added in between rounds of imaging to outcompete the adaptor and allow for sequential labeling. In this way, the sub-cellular localization of all ten proteins could be established within the same image.

In these images, the Golgi can easily be identified as a perinuclear ribbon staining strongly for both ARFGAP1 and COP1 (magenta and purple respectively, Fig. S1I). In addition, there are bright puncta throughout the cell (pink arrow heads) stained with TFG and SEC31A, indicating that they are ERES. ARFGAP1 and COPI frequently also accumulated at ERES (made more obvious in the top-right inset, where TFG and SEC31 are removed), consistent with previous reports that COPI vesicles facilitate ER to Golgi transport in human cells28–30. GFP-ATG2A is mostly cytosolic, however small accumulations of ATG2 could be found at ERES and were generally observed to be proximal to ARFGAP1/COP1 (inset; brackets). Note that clusters of LC3B and p62 are also present. Occasionally, non-ERES accumulations of p62 could be observed in close proximity to LC3B and ATG2A (lower insets), and here again ATG2 is proximal to ARFGAP1.

Critically, we can directly measure the nominal proximity of each protein relative to the others under investigation by quantifying the median distance between all ten tested proteins within 200 nm of each individual protein (Fig. S1J). All early secretory pathway proteins closely associate with each other, including the COPI proteins ARFGAP1 and β-COP. As the majority of ATG2A was cytosolic, its average distances from other proteins were greater, however, within this cohort of candidate proteins, we found that ATG2A is most closely associated with P62 and ARFGAP1 (15nm and 16nm respectively; top row of Fig. S1J), exactly consistent with the APEX screen.

The APEX and FLASH-PAINT datasets provided a clear but static result that ATG2A is recruited to ARFGAP1 positive structures. To determine whether this recruitment was coincident with phagophore formation, we stably expressed GFP-ATG2A and TagBFP-LC3B in ATG2 DKO cells in order to identify phagophores. ATG2A primarily localizes to phagophores and lipid droplets, whereas LC3B primarily localizes to phagophores, autophagosomes and autolysosomes. Where these two proteins colocalize strongly suggests a growing phagophore in the process of membrane expansion (Figs. 1B). To assess co-localization with phagophores, we expressed candidate early secretory proteins labeled with red fluorophores (Fig. 1 C and D). ARFGAP1-TagRFP colocalized with ∼90% of actively forming phagophores, whereas only ∼30% of phagophores were coincident with the ERES markers mRuby-SEC23A or RFP-p125a/SEC23IP. Further, while the ARFGAP1 signal was largely concentrated in the Golgi, phagophores most frequently colocalized with extra-Golgi ARFGAP1 puncta.

Finally, we compare our results to the subcellular localization of ARFGAP1 established by global organelle profiling available at the database at https://organelles.sf.czbiohub.org31. ARFGAP1 is enriched on autophagic membranes positive for LC3B as well as endolysosomal and early secretory membranes (Fig. 1E). Strikingly, ARFGAP1 was most highly enriched on membranes isolated via RAB1A immunoprecipitation. The autophagic proteins ATG2A, ATG2B, WDR45, and ATG9A, all of which are involved in membrane expansion of the autophagosome, were also highly enriched on RAB1A membranes. These collective results establish that both endogenous ARFGAP1 (organelle isolation, APEX, FLASH PAINT) and overexpressed ARFGAP1 (live imaging) naturally associate with ATG2A, including during phagophore biogenesis.

Autophagic flux depends upon RAB1 but not ARFGAP1

To look for a potential functional role for ARFGAP1 in autophagy, we measured autophagic flux using the HaloTag dropout assay in which liberation of a free HaloTag (HT) is indicative of lysosome-mediated proteolysis of the full-length (FL) protein32. We stably expressed HaloTag7-mGFP-LC3B in WT HEK293 cells and then used siRNA to knockdown various regulators or potential regulators of macroautophagy initiation (Fig. 2 A and B). Despite efficient knockdown, siRNA against ARFGAP1 had very little effect on flux; the ratio of free HaloTag to full-length HaloTag7-mGFP-LC3B following ARFGAP1 knockdown was similar to both nontransfected (NT) and nontargeting siRNA controls. In contrast, knockdown of the autophagy initiation factor FIP200 dramatically suppressed dropout of the HaloTag. We also looked for an interaction between ARFGAP1 and ATG2A, but even when both proteins were overexpressed, we could not detect significant interaction via co-immunoprecipitation (Fig. 2 E and F). Thus, it is likely that ARFGAP1 localizes to a key membrane involved in autophagosome biogenesis but is not itself a player in initiating autophagosome formation.

Autophagic flux depends upon RAB1 but not ARFGAP1.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that siRNA KD of RAB1 inhibits autophagic flux. In gel fluorescence reveals HaloTag7s moiety bound to the TMR ligand. Indicated bands are full length (FL) HT-mGFP-LC3B protein and HaloTag7 after cleavage (HT) in the lysosome. The ratio between FL and HT demonstrates autophagic flux, quantified in (B). Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were collected for each condition. (C-D) Immunoblot of lysates from three biological replicates demonstrating that siRNA KD of RAB1 results in an accumulation of LC3B-II. The band intensity of LC3B-II was normalized against the intensity of GAPDH for each lane. To normalize each replicate, each individual value was divided by value of the non-transfected sample. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. (E-F) Immunoblot showing CoIP between GFP-RAB1A and 3xFLAG-ATG2A compared to the same reaction with other GFP-tagged early secretory proteins. The cells were incubated with either complete media or EBSS for four hours prior to harvesting. For the CoIP, ∼3 mg of cell lysate was used per condition (slight deviations in total amount between replicates, but not between lanes). To quantify the results, the intensity of the 3xFLAG-ATG2A IP signal was ratioed against the input signal. To normalize each replicate, each individual value was divided by the average of the replicate. Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. The media was not a significant source of variation (p = 0.3535).

As shown in (Fig. 1E), both the ATG2/ATG9 lipid transport apparatus as well as ARFGAP1 are associated to RAB1A-positive membranes. RAB1 proteins mediate ER to Golgi trafficking in the early secretory pathway. The yeast RAB1 homologue, Ypt1, is essential to autophagy33, 34, while in mammals, genetic depletion of either RAB1A or RAB1B modestly reduces autophagic flux35, 36. Thus, we tested the functional and physical interaction of RAB1 proteins with ATG2. Consistent with previous results, siRNA knockdown of either RAB1A or RAB1B in our system partially inhibited flux in the Halo dropout assay (Fig. 2 A and B). Strikingly, knockdown of both homologues simultaneously was nearly as effective as knocking down FIP200, establishing that RAB1 proteins play an essential role in supporting starvation induced autophagic flux.

In addition to following the Halo-tagged LC3B, we also immunoblotted for endogenous LC3B in each of our siRNA experiments. We found that knockdown of either RAB1 isoform, and especially knockdown of both together leads to a clear accumulation of the membrane conjugated form of LC3B, LC3B-II (Fig. 2A). As LC3B overexpression can alter endogenous LC3B patterns, we confirmed this increase in a cell line expressing only HaloTag7-mGFP (Fig. 2 C and D). Knockdown of RAB1A and RAB1B drove a threefold increase of LC3B-II (Fig. 2D).

Increasing LC3B-II levels without autophagic flux is indicative of a block to macroautophagy at a late stage, downstream of initiation, and is therefore easily distinguished from the block induced by depleting FIP200 (where LC3B-II levels go down). Intriguingly, LC3B-II levels also increase when ATG2 proteins are genetically depleted (Fig. S1A,B), suggesting ATG2A and RAB1A might function together. Indeed, overexpressed GFP-RAB1A could strongly immunoprecipitate flag-tagged ATG2A in our cells, indicating these proteins are in a shared complex (Fig. 2 E and F).

We also tested several other proximity hits from the biotinylation analysis for interaction with flag-tagged ATG2A, including GFP-tagged copies of ERES resident proteins SEC23A, SEC31A, SEC16A, Sar1 and all four homologs of Sec24 (Fig. S2 A-D). We found that ATG2A co-immunoprecipitated weakly with many of them, the strongest of which was SEC24D. Note that this SEC24D interaction was barely above the background of pull-down with GFP alone (Fig. 2 E and F). Thus, consistent with proximity and imaging analyses, ATG2 proteins are in the general proximity of the ERES machinery, but protein-protein interactions appear to be limited to RAB1.

Disruption of the early secretory pathway results in the recruitment of autophagic proteins to RAB1A-positive tubular membranes

These results suggest that ATG2 engages an early secretory membrane, distinct from the ERES, in a RAB1-dependent manner. We next tested whether disruption of the early secretory system would impact the localization of ATG2A. To this end, we overexpressed SAR1BH79G, the catalytic dead mutant of the small GTPase SAR1B that recruits the COPII coat. SAR1BH79G binds the membrane and recruits the COPII coat, but is unable to release, thereby trapping COPII onto the membrane. Overexpression of this construct has been shown to reduce autophagy in starved cells17, 35. Consistent with those results, overexpression of SAR1BH79G-mOrange2 diminished autophagic flux of both a bulk cargo (HaloTag7-mGFP reduced ∼20%) and of LC3B (HaloTag7-mGFP-LC3B reduced by ∼30%) by the HaloTag degradation assay32 (Figs 3A/B). Further, LC3B puncta number were lowest in cells expressing high amounts of SAR1BH79G (Figs. S3A/B), consistent with a block to autophagosome formation rather than autophagosome clearance.

ATG2A, LC3B, ARFGAP1, and RAB1A localize to tubules following disruption of the early secretory pathway.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that SAR1BH79G-mOrange2 (mO) overexpression negatively affects autophagic flux. The ratio between the FL and the free HT represents the autophagic flux in each condition (as in Fig. 2 A/B) and is plotted in (B). Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from four biological replicates were collected for each of the HT-mGFP conditions and five replicates were collected for each of the HT-mGFP-LC3B conditions. (C) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2 that were transfected with SAR1BH79G-mOrange2 and starved for two hours. Note that GFP-ATG2A forms long tubules that do not colocalize with the SAR1BH79G-mOrange2 signal. Maximum intensity projection of a confocal image. Gray line was added to show cell outline. (D-E) Live cell imaging demonstrates the localization of ARFGAP1-TagRFP (D) and SNAP-RAB1A (E) on GFP-ATG2A/TagBFP-LC3B positive tubules following SAR1BH79G overexpression. Arrows denote thin GFP-ATG2A positive tubules whereas arrow heads denote thicker tubules that lack most GFP-ATG2A but contain high levels of LC3B, ARFGAP1, and RAB1A. Maximum intensity projections of confocal images. (F) CLEM-FIBSEM of cells prepared as in (C). GFP-ATG2A fluorescence corresponds to a thin tubular membrane that is connected to a thicker tubule. The small inset in the left image shows a single SEM slice. The turquoise arrow is pointing to the tubule which is segmented and displayed in turquoise in the larger image, whereas the yellow arrows are pointing at adjacent ER tubules which are segmented and displayed in the larger image in yellow. At many points, these ER structures come into very close apposition with the tubule which is denoted in red. The ER structures are removed in the right image for better visualization of the ATG2A positive tubule. Note that only the thin section has corresponding GFP-ATG2A fluorescence.

Unexpectedly, in a small subset of transfected cells (1.5±0.8%), overexpression of SAR1BH79G under starvation conditions shifted the localization of GFP-ATG2A from a largely punctate distribution to long, tubular objects that were SAR1BH79G negative (Fig 3C). GFP-ATG2A decorated these tubules with varying intensities approximating a beads-on-a-string pattern (e.g. arrows in Fig. 3 D and E). Fluorescently tagged forms of LC3B, ARFGAP1 and RAB1A also decorated these tubules. However, these three proteins displayed the brightest intensities where ATG2A was dimmer (e.g. arrowheads in Fig. 3 D and E). These three proteins, but not ATG2A, also decorated larger structures that appeared to be contiguous with the tubule, including bulbous regions at the tubule termini.

To determine the ultrastructure of these regions, we subjected GFP-ATG2A-expressing cells to correlative light and focused ion beam scanning electron microscopy (CLEM-FIB-SEM). The GFP-ATG2A fluorescence signal corresponded to thin, tubular membranes (Fig. 3F and S3C), which is consistent with the finding that ATG2A preferentially binds highly curved membranes37. The tubule is surrounded by ER elements, many of which appear to be in very close contact (Fig. 3F, S3C). GFP-ATG2A signal was more intense in regions with closely adjacent ER, consistent with a possible contact site localization. The tubule continued into larger structures at either end, which were GFP-ATG2-negative, and morphologically similar with the bulbous structures harboring ARFGAP1, RAB1A and LC3B in fluorescence images.

Thus, although unusual, these rare tubules afforded an opportunity to demonstrate that ATG2, ARFGAP1 and RAB1A target the same membrane, easily resolvable even at confocal resolution. ATG2 collects where ER is closely adjacent, and LC3B accumulates with RAB1A, implying a topological organization with LC3B and RAB1A decorating the same structure, suggesting that RAB1A membranes may be a component or progenitor of the phagophore membrane.

ARFGAP1 and RAB1A accumulate around pre-autophagic structures in ATG2 DKO cells

We previously published that very small, ATG9- and LC3B-positive vesicles accumulate in large clusters with the cargo-adaptor p62 in ATG2 DKO cells9. This cluster is surrounded by a shell of more complicated membranes which we termed “protophagophores” because they colocalize with other autophagy proteins such as WIPI2 and because many of them appear to adopt a cup-like shape in TEM images. Beyond this protophagophore shell, the entire structure is surrounded by ER (illustrated in Fig. 4A). Our work and others have shown that essentially all proteins involved in autophagosome biogenesis accumulate with these clusters, including FIP200, ATG14, ATG9A, ATG16L, and p629, 22, 23, suggesting they are repositories of autophagosome biogenesis material awaiting the arrival of ATG2 proteins. If early secretory membranes contribute directly in any way to the biogenesis event, they may also concentrate at these sites.

ATG2 deletion results in the accumulation of ARFGAP1 and RAB1A on autophagic membranes.

(A) Schematic of the membranes that accumulate at the ATG2 DKO compartment based off the TEM images in Olivas et al9. The core of the compartment contains small vesicles that stain for ATG9A, LC3B, and P62. The periphery of the compartment is comprised of more complicated cup shaped membranes that stain for WIPI2. The entire compartment is enwrapped by ER tubules. (B) Immunofluorescence microscopy demonstrates that ARFGAP1 colocalizes with P62 in ATG2 DKO cells. Inset 1 highlights a cell undergoing mitosis, in which the colocalization of these two markers persists as denoted by the arrows. Inset 2 shows several cells, a ring-like accumulation of ARFGAP1 around P62 is clearly visible, separate from the Golgi. Maximum intensity projection of a confocal image. (C) Immunofluorescence images demonstrating the distribution of early secretory membranes around the ATG2 DKO compartment which is marked by either P62 (mouse antibody) or LC3B (rabbit antibody). The images presented are single confocal slices of the center of the ATG2 DKO compartment. The gray lines show the cell periphery and the insets focus on the largest accumulation of P62 or LC3B in the cell. (D) Quantification of the images in (C) and in Sup Fig 4A. The enrichment factor is equal to the mean of the given protein at the periphery of the ATG2 DKO compartment compared to its mean throughout the cell. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled together for each of the conditions. (E-F) Immunofluorescence images demonstrating the enrichment of overexpressed GFP-RAB1A (E) or SNAP-RAB1A (F) at the periphery of the ATG2 DKO compartment as determined by the presence of P62 (E) or WIPI2 (F). Presented as a single confocal slice. The gray lines show the periphery of the cell and the insets focus on ATG2 DKO compartments that are distal from the Golgi.

We therefore performed IF against various markers of the early secretory pathway in ATG2 DKO cells to look for colocalization with this compartment (represented by staining for p62 (mouse antibody) or for LC3B (rabbit antibody); Figs. 4B-D). We found that ER exit site markers such as SEC16A, SEC24C, SEC31A, and TFG were not enriched around the compartment, neither was the Golgi protein GM130 (Figs. 4C-D and S4A). ßCOP was modestly enriched, but at levels that by eye, did not suggest encapsulation of the compartment the way autophagy proteins collect. In addition, overexpressed mEmerald-ERGIC53 did not localize to the ATG2 DKO compartment (Fig. S4B). In contrast, ARFGAP1 targeted essentially all of the ATG2 DKO compartments, forming a robust spherical shell at the periphery of the p62 signal (Fig. 4B-D). This distribution was maintained during mitosis in contrast to the ARFGAP1 signal in the Golgi, which dissipated (Fig. 4B).

We were unable to find a strong RAB1A signal by immunofluorescence, however, when overexpressed, GFP-RAB1A also formed a sphere around the ATG2 DKO compartment (Fig. 4E). Both ARFGAP1 and SNAP-RAB1A colocalized with WIPI2 (Fig. 4F), consistent with a model whereby these membranes arrive at the PAS prior to ATG2A recruitment and subsequent autophagosome biogenesis.

Discussion

Here we describe an interaction of ARFGAP1/RAB1-positive membranes with the autophagic biogenesis machinery including the lipid transporter ATG2. These membranes likely derive from a privileged region of the ERGIC consistent with the literature tying these intermediate secretory membranes to autophagy17, 20. In particular, the presence of both ARFGAP1 and RAB1 plus the general proximity to ERES proteins across a variety of assays is consistent with an early secretory origin for these membranes. Intriguingly, ARFGAP1 and RAB1 coalesce around the periphery of the ATG2 DKO compartment exactly where we previously described complex cup-shaped membranes, so called proto-phagophores9, accumulate. We do not detect any other ERES or ERGIC markers at these sites, implying proto-phagophores derive from only a subdomain of the ERGIC which we suspect is remodeled specifically for the purpose of supporting autophagy.

What do the ERGIC-derived membranes contribute to autophagosome formation? We speculate on three possibilities: 1) Local proteins are needed to drive other early autophagy steps. Most notably, in addition to interacting with ATG2A as shown here, RAB1 proteins also interact with other early acting autophagy proteins33, 34, 38 including especially the PI3Kinase, VPS3439, 40.

Thus, a privileged region of the ERGIC may simply best coordinate the machinery driving local protein and lipid remodeling needed to initiate phagophore formation. 2) ERGIC-derived membranes may have a unique membrane composition and structure necessary for autophagosome biogenesis. The RAB1-positive membranes that accumulate around the core of ATG9 vesicles in the DKO compartment are enriched for cup-like morphologies9. We do not yet know the structural basis for these morphologies, but it may be that membranes with a natural degree of plasticity, as has been shown for the complex network of structures emanating from COPI/COPII rich early secretory membranes28, are better adapted to the subsequent remodeling needed to shape and grow around cytoplasm isolates during starvation-mediated macroautophagy. 3) Small ATG9 vesicles may not contribute sufficient volume to grow a phagophore, even considering the very small luminal volume of a mature autophagosome.

Geometric estimates of the surface-to-volume ratio in yeast autophagosomes, gleaned from cryo-electron microscopy tomograms, suggest that as much as 20% of the membrane might derive from fusion rather than lipid transport in order to deliver volume41. Autophagosome biogenesis also depends upon the incorporation of ATG9 scramblases42, which are thought to be delivered from a trans-Golgi compartment to the site of autophagosome biogenesis. As both the ERGIC and ATG9 seed membranes collect to high local concentrations in ATG2 DKO cells, future work may be able to tease out how their separation and eventual coalescence are regulated.

Materials and methods

Plasmids and reagents

The following plasmids were used in this study: pLVX Puro GFP-ATG2A (Previously published3), pLVX Puro APEX2-GFP-ATG2A (This study; a PCR fragment containing flanking EcoRI cutsites around APEX2 was generated from pcDNA3 APEX2-NES (Addgene #49386) and was inserted into GFP-ATG2A by ligation), pCMV 3xFLAG-ATG2A (Previously published3), pLVX Neo TagBFP-LC3B (This study; pLVX puro TagBFP-LC3B3 and pLVX-EGFP-IRES-Neo (Addgene #128660) were cut with EcoRI and BamHI, following which the TagBFP-LC3B fragment was ligated into the pLVX Neo backbone), pDESt47 Sar1-GFP (Addgene #67409), SAR1B-mOrange2 (Addgene #166899), SAR1BH79G2-mOrange2 (Addgene #166900), SAR1BH79G-3xFLAG (This Study; A BamHI-3xFLAG-NotI fragment was generated by oligo annealing, following which it was ligated into the SAR1BH79G-mOrange2 backbone that was cut by the same restriction enzymes), mRuby-SEC23A (Addgene 36158), EGFP-SEC23A (Addgene #66609), pEGFP-SEC24A (This Study; a XhoI-SEC24A-PacI fragment was generated by PCR from HA-SEC24A which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), pEGFP-SEC24B (This Study; a XhoI-SEC24B-PacI fragment was generated by PCR from HA-SEC24B which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), EGFP-SEC24C (This Study; a AgeI-EGFP-XhoI fragment was generated by PCR and was ligated into the backbone of pEYFP-SEC24C (Addgene #66608) which was cut by the same restriction enzymes), pEGFP-SEC24D (Addgene #32678), GFP-SEC31A (Addgene #42124), GFP-SEC16A (Addgene #36155), ARFGAP1-TagRFP (This Study; ARFGAP1-TagRFP was a gift from the Rothman lab. A frame shift and a point mutation were corrected by site directed mutagenesis), ARFGAP1-GFP (Gift from the Antonny Lab), pMRX IB HaloTag7-mGFP (Addgene # 184903), pMRX No HaloTag7-mGFP-LC3B (Addgene # 184901), pLVX GFP (This Study; a fragment containing EGFP was generated from pLVX GFP-ATG2A and was ligated into pLVX puro (Clonetech, 632159) that was cut with EcoRI using Gibson Assembly), GFP-RAB1A (Gift from the Rothman Lab), SNAP-RAB1A (Gift from the Rothman Lab)

Antibodies for immunoblotting were used at 1:1000 and include the following: anti-ATG2A (Cell Signaling Technology, 15011S), anti-P62 (BD BioSciences, 610832), anti-GAPDH (Abcam, ab9484), anti-LC3B (Cell Signaling Technology, 3868S), anti-RB1CC1 (Thermo Fisher Scientific, 172501AP), anti-RAB1A (Cell Signaling Technology, D3X9S), anti-RAB1B (Proteintech, 17824-1-AP), anti-ARFGAP1 (Proteintech, 13571-1-AP), anti-FLAG (Sigma-Aldrich, F1804-200UG), anti-GFP (Cell Signaling Technology, 2956S), anti-mCherry (Invitrogen, PA5-34974), ECL anti-rabbit IgG horseradish peroxidase-linked (GE Healthcare, NA934V), and ECL anti-mouse IgG horseradish peroxidase-linked (GE Healthcare, NA931V).

Antibodies for immunofluorescence were used at 1:500 and include the following: anti-Biotin (Rockland, 200-301098S), anti-GFP nanobody (Massive Photonics), anti-ARFGAP1 (Proteintech, 13571-1-AP;), anti-SEC31A (BD Biosciences, 611280), anti-SEC24C (Cell Signaling Technology, 14676S), anti-MIA3 (Tango1; Sigma, HPA055922), anti-TFG (Abcam, ab156866), anti-COPI (CMIA10; made by Rothman Lab), anti-LC3B (MBL, PM036), anti-P62 (BD BioSciences, 610832), anti-SEC16A (Abcam, ab70722), anti-SEC24D (Cell Signaling Technology, 14687S), anti-GM130 (Proteintech, 11308-1-AP), anti-ERGIC53 (Abcam, EPR6979), Alexa Fluor 405 goat anti-rabbit IgG (Thermo Fisher Scientific, A31556), Alexa Fluor 405 goat anti-mouse IgG (Thermo Fisher Scientific, A31553), Alexa Fluor 594 goat anti-rabbit IgG (Thermo Fisher Scientific, A11037), Alexa Fluor 647 goat anti-rabbit IgG (Thermo Fisher Scientific, A21245), and Alexa Fluor 647 goat anti-mouse IgG (Thermo Fisher Scientific, A21236).

The following reagents were purchased for FLASH-PAINT: Cy3B-modified DNA oligonucleotides were custom-ordered from IDT. Sodium chloride 5 M (cat: AM9759) were obtained from Ambion. Ultrapure water (cat: 10977-015) was purchased from Invitrogen. µ-Slide 8-well chambers (cat: 80807) were purchased from ibidi. Methanol (cat: 9070-05) was purchased from J.T. Baker. Glycerol (cat: G5516-500ml), protocatechuate 3,4-dioxygenase pseudomonas (PCD) (cat: P8279), 3,4-dihydroxybenzoic acid (PCA) (cat: 37580-25G-F) and (+−)-6-hydroxy-2,5,7,8-tetra-methylchromane-2-carboxylic acid (Trolox) (cat: 238813-5 G) were ordered from Sigma. 1× Phosphate Buffered Saline (PBS) pH 7.2 (cat: 10010-023) was purchased from Gibco. Paraformaldehyde (cat: 15710) were obtained from Electron Microscopy Sciences. Bovine serum albumin (cat: 001-000-162) was ordered from Jackson ImmunoResearch. Triton X-100 (cat: T8787-50ML) was purchased from Sigma-Aldrich. DNA labeled Nanobodies were obtained from Massive Photonics.

SiControl, FIP200, RAB1A, RAB1B, and ARFGAP1 were acquired from Horizon Discovery (D-001810-10-05, L-0082383-00-0005, L-008958-01-0005, L-013321-02-005; Horizon Discovery).

Cell culture

HEK293 cells (ATCC, CRL-1573) were cultured at 37°C and 5% CO2 in DMEM (Thermo Fisher Scientific, 11965092) supplemented with 10% FBS (Thermo Fisher Scientific, 10438062) and 1% penicillinstreptomycin (Thermo Fisher Scientific, 15140122). Cells were starved by incubation in Earle’s Balanced Salt Solution (Thermo Fisher Scientific, 24010043) for biochemical assays or in Hank’s Balanced Salt Solution (Thermo Fisher Scientific, 14025092) for live cell imaging assays and treated with 0.1 µM bafilomycin A1 (Enzo, BML-CM110-0100) for 2 h.

Lentivirus production and transduction

HEK293 cells were seeded into a 10-cm plate. At 70% confluence, cells were transfected with 4.3 µg psPAX2 (Addgene #12260), 0.43 µg pCMV-VSV-G (Addgene, #8454), and 4.3 µg target plasmid using 20 μl Lipofectamine 3000 (Thermo Fisher Scientific, L3000008). DNA was added to 500 μl Opti-MEM (Thermo Fisher Scientific, 31985070) and Lipofectamine 3000 to 500 μl Opti-MEM in separate 1.5 ml microcentrifuge tubes. The tubes were then mixed and left for 10 min before adding the mixture dropwise to cells in fresh DMEM. After overnight incubation at 37°C, the medium was replaced with fresh DMEM. Cell medium was then collected after 24-48 h and filtered with a 0.45-µm syringe filter (Pall, 4184). Generated virus was either used immediately or was aliquoted and stored at −80°C. HEK293 cells to be transduced were seeded into a 6-well plate. Undiluted virus was added to cells and incubated at 37°C for 24 h. The medium was replaced with fresh DMEM and left for another 24 h. The cells were then treated with 2 µg/ml puromycin (Clontech, 631306), 2 mg/ml Geneticin (Thermo Fisher Scientific, 10131035), or 10 µg/ml Blasticidin (Gibco, A11139-03) for at least 2 wk. For the case of pMRX No HaloTag7-mGFP-LC3B, which lacked a selection marker, the cells were transduced and subsequently sorted by flow assisted cell sorting to normalize GFP expression levels.

Proximity labeling

Proximity labeling was performed as previously described26. ATG2 DKO cells stably expressing APEX2-EGFP-ATG2A were cultured for two weeks in media containing heavy, medium, or light isotopes of arginine and lysine (Cambridge Isotope Laboratories, CLM-2265-H-PK, DLM-2640-PK, CNLM-291-H-PK; Sigma-Aldrich, 608033). All conditions were expanded to 3×15cm plates prior to proximity labeling. Heavy labeled cells were starved in EBSS and treated with 100 nM Bafilomycin A1 (Enzo, BML-CM110-0100) for 2 hrs prior to labeling. All conditions were incubated with 500 μM Biotin Phenol (Biotinyl Tyramide; Sigma-Aldrich, SML2135) for 30 min at 37°C. The heavy and medium isotope conditions were treated with 1 mM H2O2 (Sigma-Aldrich, H1009-100ML) for 1 min at room temperature. All conditions were then aspirated and scraped in quencher solution containing 10 mM sodium azide (Sigma-Aldrich, S2002-100G), 10 mM sodium ascorbate (Sigma-Aldrich, A4034-100G) and 5 mM Trolox (Sigma-Aldrich, 238813-5G) in DPBS. The cells were then spun down at 77xg for 5 min at 4°C and were resuspended in quencher solution. This process was repeated for a total of three washes. The cells were then lysed in RIPA buffer (50 mM Tris, 150 mM NaCl, 0.1% SDS, 0.5% sodium deoxycholate (Sigma-Aldrich, D6750-10G), 1% NP40 (Roche, 11754599001) and 1x protease inhibitor (Roche, 11836170001)) for 30 min prior to centrifugation for 10 min at 18,407xg at 4°C. The supernatants were collected and protein concentration was assessed via bradford assay. Heavy, medium and light cells were mixed at a 1:1:1 ratio for a combined mass of 3-4 mg of protein. The protein mixtures were subsequently incubated with streptavidin magnetic beads (Pierce, 88817) for 2 hrs at 4°C after which they were washed twice with RIPA lysis buffer, once with 1 M KCl, once with 0.1 M Na2CO3, once with 2 M urea in 10 mM Tris (pH 8.0) and twice with RIPA lysis buffer for each sample. All washes were performed at 4°C. The proteins were eluted off of the beads by boiling the samples in 2x LDS sample buffer (Invitrogen, NP0007). The eluted fraction was run on a gel (Invitrogen, NP0341) for a 10 minutes. The resulting gel plug was stained (Invitrogen, 46-5034) and excised and sent to the Yale Keck facility for mass spectrometry analysis.

Mass spectrometry

MS was performed at Yale Mass Spectrometry & Proteomics Resource of the W.M. Keck Foundation Biotechnology Resource Laboratory. Samples were analyzed in biological triplicates. The data were normalized to internal controls and total spectral counts. For differential analysis between the two groups 10E-04 to was added to all values, the data was log2-transformed, and an unpaired Student t-test was performed. Differentially abundant proteins were considered those with an absolute fold change > 1.5 and p <0.05.

FLASH-PAINT Imaging

Cells were fixed with 4% PFA (Electron Microscopy Services, 15710-S) for 30 min. After three washes in 1xPBS for five minutes each, cells were blocked and permeabilized with 3% BSA and 0.25% Triton X-100 at room temperature for 1 h. Next, cells were incubated with the anti-ARFGAP1 antibody, anti-SEC31A antibody, and the GFP-Nanobody in 3% BSA and 0.1% Triton X-100 at 4 °C overnight. Additionally, all other primary antibodies were pre incubated with the corresponding nanobodies (Tables S1) at 4 °C overnight. The next day, after four washes (30 s, 60 s, 2× 5 min) cells were incubated with the nanobodies corresponding to ARFGAP1 antibody and SEC31A antibody for ∼2 h at room temperature. Next, unlabeled excess secondary nanobodies (to block unlabeled epitopes) were added to pre-incubation antibody-nanobody mixes at room temperature for 5 min. Next, the cells were incubated with the pooled antibody-nanobody mix for ∼2.5 h at room temperature. After four washes (30 s, 60 s, 2× 5 min), the sample was post-fixed with 3% PFA and 0.1% GA for 10 min. Finally, samples were washed three times with 1× PBS for 5 min each before adding the imaging solution.

The imaging solution consisted of 1x PBS and 500 mM NaCl and was supplemented with 1× Trolox, (100× Trolox: 100 mg Trolox, 430 μL 100 % Methanol, 345 μL 1M NaOH in 3.2 mL H2O), 1× PCA (40× PCA: 154 mg PCA, 10 mL water and NaOH were mixed and pH was adjusted 9.0), and 1× PCD (100× PCD: 9.3 mg PCD, 13.3 mL of buffer (100 mM Tris-HCl pH 8, 50 mM KCl, 1 mM EDTA, 50 % Glycerol)). All three stock buffers were frozen and stored at -20 °C.

Fluorescence imaging was carried out on an inverted Nikon Eclipse Ti2 microscope (Nikon Instruments) with the Perfect Focus System, attached to a Andor Dragonfly unit. The Dragonfly was used in the BTIRF mode, applying an objective-type TIRF configuration with an oil-immersion objective (Nikon Instruments, Apo SR TIRF 60×, NA 1.49, Oil). As excitation laser, a 561 nm (1W nominal) was used. The beam was coupled into a multimode fiber going through the Andor Borealis unit reshaping the beam from a gaussian profile to a homogenous flat top. As dichroic mirror a CR-DFLY-DMQD-01 was used. Fluorescence light was spectrally filtered with an emission filter (TR-DFLY-F600-050) and imaged on a scientific complementary metal oxide semiconductor (sCMOS) camera (Sona 4BV6X, Andor Technologies) without further magnification, resulting in an effective pixel size of 108 nm. The power at the objective lens was ∼10 % of the power set at the laser.

Imaging was carried out using the corresponding Imager and Adapter (Table S2, S3, S5) in imaging buffer. 25,000 frames were acquired at 25 ms exposure time. The readout bandwidth was set to 540 MHz. Laser power (@561 nm) was set to 80 mW (measured before the back focal plane (BFP) of the objective), corresponding to ∼1.8 kW/cm2 at the sample plane. After imaging the sample was subsequently washed three times with 200 µL each with 1× PBS (on the microscope) followed by an incubation with the corresponding eraser (Table S4; at 20 nM) for ∼ 5 min. This process was repeated iteratively for ten sequential rounds (Table S5).

DNA-PAINT data was reconstructed, postprocessed (drift correction and alignment of imaging rounds) and rendered with the Picasso package43.

Focused ion beam scanning electron microscopy and correlative light electron microscopy

For focused ion beam scanning electron microscopy (FIBSEM) CLEM, ATG2 DKO HEK293 cells overexpressing GFP-ATG2A were plated on 35-mm MatTek dish (MatTek, P35G-1.5-14-CGRD). The cells were starved for two hours prior to pre-fixation in 4% PFA + 0.25% glutaraldehyde and three washes in 1x PBS. Regions of interest by fluorescence light microscopy were selected and their coordinates on the dish were identified using phase contrast. Cells were further fixed with 2.5% glutaraldehyde in 0.1 M sodium cacodylate buffer, postfixed in 2% OsO4 and 1.5% K4Fe(CN)6 (Sigma-Aldrich) in 0.1 M sodium cacodylate buffer, en bloc stained with 2% aqueous uranyl acetate, dehydrated, embedded in Embed 812, and polymerized at 60°C for 48 hrs. Cells of interest were relocated based on the pre-recorded coordinates. Epon block was glued onto the SEM sample mounting aluminum stub and platinum en bloc coating on the sample surface was carried out with the sputter coater (Ted Pella, Inc.). The cell of interest was relocated under SEM imaging based on pre-recorded coordinates and FIB-SEM imaged in a Crossbeam 550 FIB-SEM workstation operating under SmartSEM (Carl Zeiss Microscopy GmbH) and Atlas 5 engine (Fibics, Inc.). The imaging resolution was set at 7 nm/pixel in the X, Y axis with milling being performed at 7 nm/step along the Z axis to achieve an isotropic resolution of 7 nm/voxel. Images were aligned and exported with Atlas 5 (Fibics, Inc.), further processed and segmented with DragonFly Pro software (Object Research Systems [ORS], Inc.).

Lysis, gel electrophoresis, and immunoblotting

Cells were collected by scraping in PBS. The cells were centrifuged at 500 g for 3 min at 4°C, and lysed in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor cocktail (Sigma-Aldrich, 11873580001) on ice for five minutes. Lysates were collected after centrifugation at 16,000 g for 10 min at 4°C. Total protein concentration was determined by a bradford assay (BioRad, 5000006). 50-100 μM DTT (American Bio, AB00490-00010) and 1xLDS (Invitrogen, NP0007) loading buffer were added to each sample, which were then electrophoresed on precast Bis-Tris gels (Thermo Fisher Scientific, NP0341BOX or NP0322BOX). Gels were subsequently transferred to Immobilon-FL PVDF membranes (Sigma-Aldrich, IPFL00010) at 30-35 V for 90 min.

The membranes were blocked with 5% BSA (Sigma-Aldrich, A9647) in PBS-T (PBS containing 0.1% Tween 20; Thermo Scientific, CAS 9005-64-5) for 1 h at room temperature and incubated with primary antibody (diluted 1:1,000 in PBS-T containing 5% BSA [Sigma-Aldrich, A9647] and 0.03% sodium azide (Sigma-Aldrich, S2002)) overnight at 4°C. Membranes were washed three times in PBS-T before incubation with the secondary antibody which was diluted 1:5,000 in PBS-T containing 0.5% Omniblock (American Bio, AB10109-00100) for 1 h. Membranes were then washed three times in PBST and treated with SuperSignal West Femto substrate (Thermo Fisher Scientific, 34096) for 5 min before imaging with the Bio-Rad VersaDoc imaging system.

Immunoprecipitation

For immunoprecipitation experiments, 15 cm plates were seeded and transfected upon reaching 80-90% confluency. The next day, the cells were collected by scraping, washed 1x in PBS, and lysed as previously described. Following quantification of protein levels by a Bradford assay, equal amounts of protein were added to pre-washed GFP-Trap beads (Proteintech, gtma-20).

The volume was increased to 500 uL and the bead/lysate slurry was incubated at 4°C for two hours while rotating. The beads were pelleted using a magnetic rack and washed five times in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor. The target proteins were eluted by boiling the beads for 10 min in 2x LDS loading buffer.

Densitometry and quantification

Densitometry quantifications of immunoblots and in-gel fluorescence images were performed using ImageJ software. For LC3B and P62 clearance analysis (Fig. S1 A-C), the band intensity of LC3B or P62 for each condition was normalized against the GAPDH intensity of that lane. To normalize each replicate, each value was divided by the mean value of the entire replicate. This normalization technique was also employed in the co-immunoprecipitation experiments.

Statistical significance was determined by two-way ANOVA followed by a Tukey multiple comparison test to compare cell types. To quantify the co-immunoprecipitation reactions, the intensity of the 3xFLAG-ATG2A IP band was ratioed against the intensity of its corresponding input band. Following normalization, statistical significance was determined by a one (Fig. S3 A-D) or two-way (Fig. 2E-F) ANOVA followed by a Dunnett multiple comparison test to comparing each sample to the control lane. For the HaloTag flux assay (Fig. 2A, 2B, 3A, 3B), the Halotag flux rate was calculated as the ratio of the free Halotag band divided by the unprocessed Halotag band. Statistical significance was determined by a one (Fig. 2 A and B) or two-way (Fig 3 A and B) ANOVA followed by a Dunnett multiple comparison test to comparing each sample to the control lane. For one way ANOVA tests, data distributions were assumed to be normal. For two way ANOVA tests, the Geisser-Greenhouse correction was used as equal variability was not assumed. All data are n = 3 or greater and represent biological replicates. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 9 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001. All blots were adjusted in contrast for better visualization of the bands in the figures.

LC3B flux analysis

WT, ATG2 DKO and ATG2 DKO + APEX2-GFP-ATG2A HEK293 cells were cultured in 6-well plates to 70-90% confluence. At the beginning of the assay, the media was replaced with fresh DMEM, DMEM containing 100 nM bafilomycin A1, or EBSS containing 100 nM bafilomycin A1. After a 2 hour incubation, cells were harvested by scraping, washed with 1x PBS, and lysed. 20 μg of protein from each sample were analyzed by SDS-PAGE and immunoblotting for ATG2A, P62, GAPDH, and LC3B.

HaloTag flux assay

HEK293 WT stably expressing either HaloTag7-mGFP or HaloTag7-mGFP-LC3B were cultured in 6-well plates to 60-70% confluence prior to transfection with either SAR1B/SAR1BH79G-mOrange2 for one day (Fig. 3A) or siRNA for two days (Fig. 2A). SAR1B transfection was performed using Lipofectamine 3000 (Thermofisher Scientific, L3000015) according to the manufacturer’s protocol. SiRNA transfection was performed using Dharmafect (Horizon Discovery, T-2001-01) according to the manufacturer’s protocol. The cells were incubated with 100 nm TMR-conjugated HaloTag ligand (Promega, G8251) for 30 min. Following two washes with 1xPBS, the cells were starved in EBSS for four hours. The cells were then collected and lysed. Following SDS-PAGE separation, in gel fluorescence was immediately measured with a ChemiDoc MP Imaging System (BioRad). Images were taken at 30 s and 100 s exposures. The gel was then transferred and immunoblotted against mCherry and GAPDH (Fig 3A) or FIP200, RAB1A, RAB1B, ARFGAP1, GAPDH, and LC3B (Fig 2A).

Immunofluorescence

Cells seeded on coverslips were fixed in 4% PFA (Electron Microscopy Sciences, 15710), washed three times with PBS, and permeabilized in PBS containing 0.1% Saponin (Sigma-Aldrich, 47036) and 3% BSA. Incubation with primary antibodies was done at 4C for 18-24 hrs and secondary antibodies at room temperature for 1 hr at indicated concentrations (see Reagents). After each antibody incubation, cells were washed three times in PBS containing 0.1% Saponin and 3% BSA. The coverslips were mounted on precleaned microscope slides with Fluoromont-G mounting reagent (Southern Biotech, 0100-01). Images were acquired on an inverted Zeiss LSM 880 laser scanning confocal microscope with AiryScan, using Zen acquisition software (Fig S1D) or a Nikon CSU-W1 SoRa 60x oil objective, using the Nikon Elements software (Fig. 1C, 3C-F, S3A, S3C, 4A, 4C, 4E, 4F). All images were processed using ImageJ software. Images are displayed either as maximum intensity projections or single confocal slices as stated in the figure legends.

To assess the enrichment factor of early secretory proteins around the ATG2 DKO compartment (Fig. 4D), the phagophore channel which was denoted by either LC3B or P62 was masked and duplicated. Following outlier removal, one masked image was dilated and the other eroded. Following an XOR function, the resulting signal formed a ring around the ATG2 DKO compartment and the early secretory protein channel(s) were measured using this ROI. Separately, the early secretory protein channel was masked using a manual threshold and measured. The enrichment factor was calculated as (Meancompartment-Mincell)/(Meancell-Mincell). Three biological replicates were combined and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 9 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

Live cell imaging

Cells were plated on poly-d-lysine coated MatTek dishes (MatTek, P35GC-1.5-14-C). The following day the cells were transfected using Lipofectamine 3000 (Thermo Fisher Scientific, L3000015) according to the manufacturer’s protocol for 18-24 hrs. The cells were starved in HBSS (Thermo Fisher Scientific, 14025092) for 2-4 h and imaged in the same solution at 37°C. All images were analyzed using ImageJ and are displayed as maximum intensity projections.

For phagophore quantification (Fig. 1 C and D), each channel was masked using a manual threshold and outliers were removed. The ATG2A and LC3B channels were merged using the AND function and quantified as the total number of phagophores. The phagophore channel was then merged using the AND function with the transfected gene and quantified as the number of positive phagophores. Three biological replicates were combined and the ratio of positive over total was plotted and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 9 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

eLife Assessment

This study provides valuable insights into the crosstalk between ATG2A with components of the early secretory pathway, namely RAB1A and ARFGAP1. The evidence supporting the claims is convincing. However, the manuscript would benefit from a more in-depth exploration of the details of the role of RAB1A in autophagy and the functional implications of its interaction with ATG2A. In addition, the molecular details of the role of ARFGAP1 in this complex need further clarification

Reviewer #1 (Public review):

Summary:

D. Fuller et al. set out to study the molecular partners that cooperate with ATG2A, a lipid transfer protein essential for phagophore elongation, during the process of autophagy. Through a series of experiments combining microscopy and biochemistry, the authors identify ARFGAP1 and Rab1A as components of early autophagic membranes, which accumulate at the periphery of aberrant pre-autophagosomal structures induced by loss of ATG2. While ARFGAP1 has no apparent function in autophagy, the authors show that RAB1A is implicated in autophagy, although the mechanisms are not explored in the manuscript.

Strengths:

The work presented by Fuller et al. provides new insights into the composition of early autophagic membranes. The authors provide a series of MS experiments identifying proteins in close proximity to ATG2A, which is a valuable dataset for the field. Furthermore, they show for the first time the interaction between ATG2A and RAB1A, both in fed and starved conditions, which extends the characterisation of the pre-autophagosomal structures observed in ATG2 DKO cells.

Weaknesses:

The authors claim that this study elucidates the role of early secretory membranes in phagophore formation. However, this work is largely observational, which presents compelling evidence on the association between RAB1A GTPase and ATG2A without providing mechanistic insights into the functional relevance of this interaction. It remains unclear whether Rab1A depletion phenocopies ATG2A depletion in terms of autophagy progression or accumulation of pre-autophagosomal structures.

Furthermore, this research is conducted exclusively in HEK293 cells. Including at least one additional cell line would significantly strengthen the main findings (i.e., effects on LC3-II accumulation observed for RAB1A/B knockdown, given the previously published data on this topic).

A notable weakness of this manuscript, in this reviewer's opinion, lies in the discussion of the data in the context of existing literature. The discussion is rather short, mostly focused on the phenotype observed in ATG2 DKO cells. While this phenotype is certainly intriguing, it feels the discussion overlooks some important aspects, as outlined in the comments to the authors.

Reviewer #2 (Public review):

The mechanisms governing autophagic membrane expansion remain incompletely understood. ATG2 is known to function as a lipid transfer protein critical for this process; however, how ATG2 is coordinated with the broader autophagic machinery and endomembrane systems has remained elusive. In this study, the authors employ an elegant proximity labeling approach and identify two ER-Golgi intermediate compartment (ERGIC)-localized proteins-Rab1 and ARFGAP1-as novel regulators of ATG2 during autophagic membrane expansion.

Their findings support a model in which autophagosome formation occurs within a specialized subdomain of the ER that is enriched in both ER exit sites (ERES) and ERGIC, providing valuable mechanistic insight. The overall study is well-executed and offers an important contribution to our understanding of autophagy.

Specific Comments

(1) Integration with Prior Literature

The data convincingly implicate the ERES-ERGIC interface in autophagosome biogenesis. It would strengthen the manuscript to discuss previous studies reporting ERES and ERGIC remodeling and formation of ERERS-ERGIC contact sites (PMID: 34561617; PMID: 28754694) in the context of the current findings.

(2) Experimental Conditions

In Figures 2A-C and Figure 4, it is unclear how the cells were treated. Were they starved in EBSS? This information should be included in the corresponding figure legends.

(3) LC3 Lipidation vs. Cleavage

In Figure 2A, ARFGAP1 knockdown appears to reduce LC3 lipidation without affecting Halo-LC3 cleavage. Clarifying this observation would help readers better understand the functional specificity of ARFGAP1 in the pathway.

(4) Use of HT-mGFP in Figure 2C

It should be clarified whether the assay in Figure 2C was performed in the presence of HT-mGFP. Explaining the rationale would aid the interpretation of the results.

(5) COPII Inhibition Strategy

The authors used the dominant-active SAR1(H79G) mutant to inhibit COPII function. While this is effective in in vitro budding assays, the GDP-locked mutant SAR1(T39N) has been shown to be more effective in blocking COPII-mediated trafficking in cells. Including SAR1(T39N) in the analysis would provide stronger support for the conclusions.

Reviewer #3 (Public review):

The manuscript by Fuller et al describes a crosstalk between ARTG2A with components of the early secretory pathway, namely RAB1A and ARFGAP1. They show that ATG2A is recruited to membranes positive for RAB1A, which they also show to interact with ATG2A. In agreement with earlier findings by other groups, silencing RAB1A negatively affects autophagy. While ARFGAP1 was also found on ATG2A-positive membranes, silencing ARFGAP1 had no impact on autophagy. Notably, these ARFGAP1-positive membranes are not Golgi membranes.

The findings are interesting, and in general, the data are of good quality; however, I have outstanding questions. An answer to any of these questions might strengthen the manuscript:

(1) Are the membranes to which ATG2A is recruited a form of ERGIC?

(2) Figure 3A/B: Is it possible to show a better example? The difference is barely detectable by eye. Since immunoblotting is not really a quantitative method, I think that such a weak effect is prone to be wrong. Is there another tool/assay to validate this result?

(3) Is the curvature-sensitive region of ARFGAP1 required for its co-localization with ATG2A?

(4) What does Rab1A do? What is its effector? Or does the GTPase itself remodel the membrane?

(5) What about Arf1? It appears that the role of ARFGAP1 is unrelated to Arf1 and COPI? Thus, one would predict that Arf1 does not localize to these structures and does not affect ATG2A function.

(6) Does ARFGAP1 promote fission of the membrane from its donor compartment?

(7) What are ARFGAP1 and Rab1A recruited to? What is the lipid composition or protein that recruits these two players to regulate autophagy?

Author response:

The following is the authors’ response to the original reviews.

We thank the reviewers and editors for their thoughtful comments, which substantially improved the quality and clarity of our manuscript. We have attempted to address each major concern with either new experiments or significant textual revisions.

Reviewer 1 noted that “this research is conducted exclusively in HEK293 cells… including at least one additional cell line would significantly strengthen the main findings.” To directly address this concern, we repeated our RAB1A/B double-knockdown experiments in H4 neuroglioma cells, which endogenously express a tandem fluorescent-tagged LC3B reporter. Using flow cytometry to quantify autophagic flux, we confirmed that RAB1 depletion in H4 cells recapitulates the flux defects observed in HEK293 cells, thereby validating the generality of our findings across distinct lineages.

To validate the robustness of the ATG2 DKO phenotype and the localization of ARFGAP1-positive membranes, we acquired an ATG2 double knockout HeLa cell line. We confirmed the presence of the characteristic large ATG2-deficient PAS compartment in HeLa cells, and the recruitment of ARFGAP1 membranes, but note that ARFGAP1 displays a solid distribution through the compartment in these cells, in contrast to the more peripheral enrichment observed in HEK293 cells. These data are now included and discussed in the revised manuscript.

Multiple reviewers asked for greater clarity around the interaction between ATG2A and RAB1A. Although our original data showed that these proteins co-immunoprecipitate in cells, we had not established whether their association was direct. In response, we attempted in vitro co-immunoprecipitations from purified components. As we could not detect interactions in this simplified system, we now speculate that the ATG2A–RAB1A interaction is indirect. This clarification is now incorporated into the results section.

Multiple reviewers also raised questions regarding the nature of the membranes recruiting ARFGAP1 and the potential relationship to Arf1 and Golgi trafficking. In particular, Reviewer 3 asked: “(5) What about Arf1? … one would predict that Arf1 does not localize to these structures and does not affect ATG2A function.” To examine whether ARFGAP1 recruitment depends on Golgi integrity or Arf1-regulated trafficking, we perturbed the Golgi using three mechanistically distinct methods: Brefeldin A, mitotic entry, and SidM expression, each of which dissolves Golgi architecture. In each condition, ARFGAP1 localization to the enlarged PAS compartment in ATG2 DKO cells was unchanged. These results indicate that ARFGAP1 recruitment is independent of Golgi structure and provide indirect support for the notion that Arf1 does not participate in this process. Reviewer 3 also asked: “Is the curvature-sensitive region of ARFGAP1 required for its co-localization with ATG2A?” To address this, we generated ARFGAP1 mutants lacking either GAP catalytic activity or the ALPS curvature-sensing domain. When expressed in ATG2 DKO cells, all mutants retained full recruitment to the PAS compartment. Thus, neither GAP activity nor ALPS-mediated curvature sensing is required for ARFGAP1 localization in this context.

Response to Reviewer 3 -“(2) Figure 3A/B: … is there another tool/assay to validate this result?”—we quantified autophagic flux following SAR1B(H79G) overexpression using the flow-cytometry tandem-fluorescent LC3 assay. These experiments confirmed that SAR1B(H79G) causes only a modest reduction in autophagic flux, consistent with partial inhibition of COPII, thereby supporting our original interpretation.

We also took steps to improve the integration of our findings with prior literature. Reviewer 2 requested that we strengthen the manuscript by incorporating studies on ERES–ERGIC remodeling (“It would strengthen the manuscript to discuss previous studies…”). We now cite and discuss the studies corresponding to PMIDs 34561617 and 28754694, aligning our observations with mechanistic models of early secretory pathway remodeling. More broadly, Reviewer 1 commented that our discussion “overlooks some important aspects,” and Reviewer 3 asked, “Are the membranes to which ATG2A is recruited a form of ERGIC?” In response, we substantially rewrote the discussion, expanding our integration of existing literature and explicitly addressing models in which ATG2A acts at an ERGIC-derived membrane.

Version 2

Abstract

Autophagosomes form from seed membranes that expand through bulk-lipid transport via the bridge-like lipid transporter ATG2. The origins of the seed membranes and their relationship to the lipid transport machinery are poorly understood. Using proximity labeling and a variety of fluorescence microscopy techniques, we show that ATG2A localizes to extra-Golgi ARFGAP1 puncta during autophagosome biogenesis. ARFGAP1 itself is dispensable during macroautophagy, but among other proteins associating to these membranes, we find that RAB1 is essential. ATG2A co-immunoprecipitates strongly, albeit indirectly, with RAB1A, and siRNA-mediated depletion of RAB1A/B blocks autophagy downstream of LC3B lipidation, similar to ATG2A depletion. Further, when either autophagosome formation or the early secretory pathway is perturbed, ARFGAP1 and RAB1A accumulate at ectopic locations with autophagic machinery. Our results indicate that ATG2A engages a RAB1A complex on select early secretory membranes in support of autophagosome biogenesis.

Introduction

Macroautophagy is a cellular degradative pathway in which a seed vesicle grows to encapsulate a target cargo, closes to form the complete autophagosome, and subsequently fuses with the lysosome. This transient organelle forms de novo in approximately ten minutes and requires the delivery of millions of lipids1. ATG2 proteins function as bridge-like bulk lipid transport proteins2-5 and are essential during the membrane expansion step of autophagosome biogenesis. We propose that ATG2 bridges the ER and the developing phagophore, thereby allowing lipids to cross the cytosol through a long hydrophobic groove3. At the phagophore, lipids are equilibrated between both leaflets of the bilayer via the ATG9 scramblases6-8. As such, ATG9 is essential to macroautophagy and is a key component of the seed membranes that ultimately expand during ATG2-mediated lipid transport9, 10.

In addition to ATG9 vesicles, several other membrane sources have been proposed to be involved in autophagosome biogenesis. In particular, both genetics and biochemistry implicate an essential role for the early secretory pathway. In yeast, some COP-II vesicle and ER exit site (ERES) proteins contribute to starvation-mediated autophagy at a stage that is downstream of formation of the pre-autophagosome structure (PAS)11, 12. Furthermore, yeast ERES naturally colocalize with the autophagic cargo Ape113, and that colocalization increases when autophagosome growth is stalled by depletion of essential factors13, 14. Autophagy proteins themselves collect at ERES; both imaging and cross-linking studies indicate a close association between the two membranes and disruption of ERES activity through depletion of Sec12 dramatically reduces Atg8 puncta formation15. Finally, the direct integration of yeast COPII-derived membranes into developing phagophores is suggested by the detection of the COP-II cargo Axl2-GFP, a transmembrane protein, at both phagophores and mature autophagosomes16.

In human cell lines, genetic or pharmacological disruption of the COPII-related small GTPase SAR1A reduces the frequency of autophagic events17. Following biochemical fractionation of cellular membranes, the ER-to-Golgi Intermediate Compartment (ERGIC) is especially competent for supporting in vitro lipidation by LC3B, the primary marker of autophagic structures17, 18 and clusters of COPII positive membranes can be seen accumulating adjacent to phagophores19. Like in yeast, these studies can be interpreted as a key role for ERES or ERGIC membranes in forming or expanding autophagosome membranes. As starvation-mediated macroautophagy drives substantial remodeling of both the ERES and ERGIC membranes18, 20, including the formation of subdomains with unique protein-dependent organelle-organelle contact sites21, it is possible that the ERES/ERGIC contribution to autophagy represents a distinct subcompartment of these membrane structures.

Presumably, any contribution of an early secretory membrane must ultimately coordinate with recruitment of ATG9 vesicles, in order to ensure that the lipid transport machinery is competent to expand the phagophore membrane. Several studies have postulated an early fusion intermediate in phagophore formation, such as the description of the HyPas compartment where vesicles bearing distinct autophagy proteins (ATG16L1 and FIP200) were observed to coalesce early in autophagosome biogenesis22. Intriguingly, when the lipid transporter ATG2 is genetically depleted from cells, all of the essential autophagy machinery including ATG9 concentrates into large compartments9, 23-25. Correlative light and electron microscopy of these structures indicate they are large assemblies of ATG9 vesicles surrounded by more complex membranes which harbor FIP200 and WIPI2, among other proteins9, 26. These complex membranes often adopt a cup-like shape, leading us to name them “proto-phagophores”9. The origins of the proto-phagophores have been unknown.

In this study, we demonstrate that ATG2A is enriched during autophagosome biogenesis at RAB1/ARFGAP1 positive membranes. Although both proteins can be found within the ERGIC compartment, proximity data and colocalization analyses in secretory pathway disrupted cells suggest that RAB1 and ARFGAP1 membranes involved in ATG2 recruitment are distinct from generic ERGIC membranes. Genetic depletion experiments indicate that RAB1 and ATG2 function at a similar step in autophagosome formation and further these proteins can be isolated together in a complex. Finally, both ARFGAP1 and RAB1 co-localize with proto-phagophores, implying these membranes can recruit much of the key early autophagy machinery and may be dynamically remodeled into cup-like morphologies.

Results

ATG2A is proximal to early secretory membranes

To determine the local proteome adjacent to ATG2A, we used a proximity labeling approach. We fused the biotinylation enzyme APEX2 and a green fluorescent marker to the amino-terminus of ATG2A (APEX2-EGFP-ATG2A) and expressed this construct in ATG2A/B double knockout cells (ATG2 DKO cells). In this way, the APEX2 enzyme will biotinylate cellular proteins in close proximity to the expressed ATG2A fusion protein27. To ensure that the APEX2-EGFP-ATG2A construct is active and correctly localized, we measured the restoration of macroautophagy in these cells by following LC3B flux and p62 clearance (Fig. S1A-C). While both LC3B-II and p62 dramatically accumulate in the ATG2 DKO cells, this accumulation is almost entirely reversed upon re-expression of APEX2-EGFP-ATG2A.

We also validated that biotinylation in these cells depends upon the presence of APEX-EGFP-ATG2A and a full reaction. In cells expressing APEX2-EGFP-ATG2A, biotin is only observed when both the biotin substrate (biotin phenol) and the reaction promoter (hydrogen peroxide) are included (Fig. S1D,E). Likewise, streptavidin-mediated immunoprecipitation revealed a strong enrichment of biotinylated material only when the full reaction conditions were employed (Fig. S1F). Furthermore, the anti-biotin and GFP signals colocalized establishing that biotinylation is favored at sites proximal to ATG2A (Fig. S1D).

Mass spectrometry of proteins isolated by streptavidin-mediated immunoprecipitation was performed using a three state quantitative SILAC experiment27 with light, medium, and heavy isotopes as follows: no hydrogen peroxide as a negative control, a full reaction in complete media, and a full reaction in starvation media and 100 nM bafilomycin. When comparing both the heavy and medium isotope states to the light state, mass spectrometry analysis revealed an abundance of autophagy related proteins in both full media and under starvation (Figs. 1A, S1G, respectively). The majority of these proteins were cargo adaptors such as NBR1, p62 (SQSTM1), and TAX1BP1, suggesting that the target cargo is immediately adjacent to the biogenesis machinery during autophagosome maturation. We also observed a clear enrichment of ERES or COPII proteins including SEC16A, SEC23A, SEC23IP, and SEC24B/C. Additional ERES proteins such as TFG and SEC31A were also enriched in some, but not all, of the three replicates. In addition, ARFGAP1, a COPI effector that is primarily localized to the Golgi but also found in earlier secretory membranes, was one of the two most enriched proteins in both full media and starvation. Both ARFGAP1 and the COPII proteins suggest ATG2 is localizing to the early secretory pathway.

ATG2A is proximal to ARFGAP1 positive membranes.

(A) Proximity labeling of HEK293 ATG2 DKO cells stably expressing APEX2-GFP-ATG2A was performed as previously described27. Three biological replicates were averaged and plotted as the log2 ratio between the full labeling reaction in complete media over cells that were not treated with hydrogen peroxide on the x axis, and -log10 of the p-value on the y axis. The red line demonstrates the p-value cutoff of 0.05. The proteomics revealed an enrichment of autophagy (labeled in green) and early secretory proteins (COPII/ER Exit Site proteins are labeled in orange, ARFGAP1 is labeled in red). (B) Strategy to identify phagophores by live cell imaging. Phagophores are defined as sites where both GFP-ATG2A and TagBFP-LC3B co-localize. (C-D) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2A and TagBFP-LC3B and transfected with mRuby-SEC23A or RFP-SEC23IP/p125a or ARFGAP1-TagRFP. The cells were starved for 2-4 hours prior to imaging. The fraction of phagophores (white arrows) that colocalized with the third fluorescent protein (transfected gene) was quantified and plotted in (D). The gray lines show the cell periphery and the insets focus on representative phagophores. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled for each condition. Maximum intensity projections of confocal images. (E) Global organelle profiling data from https://organelles.sf.czbiohub.org32. Whole organelle profiles were established by pulling down endogenously tagged proteins and performing quantitative proteomics. The enrichment value for each protein were calculated by taking the difference in median log2 LFQ values between the replicates and the null distribution.

ATG2A is recruited to ARFGAP1 membranes during autophagosome biogenesis

As an orthogonal method to assess the proximity of ATG2A to early secretory components, we performed FLASH-PAINT, a super resolution technique with high multiplexing capabilities28. FLASH-PAINT offers sub-organelle resolution while simultaneously allowing us to screen >10 proteins. ATG2 DKO cells stably expressing GFP-ATG2A were starved and then immunostained for autophagic proteins (GFP-ATG2A, p62, and LC3B), ERES proteins (Tango1, SEC24C, SEC31A, SEC16A, TFG), and markers of COPI-positive membranes (ARFGAP1 and COPI). All proteins were concurrently labeled with primary antibodies that were in turn identified by their association with secondary antibodies with unique DNA barcodes (Fig. S1H). A fluorophore conjugated imager DNA strand and adaptor strands were added to target the imager to the correct secondary antibody. Eraser strands were added in between rounds of imaging to outcompete the adaptor and allow for sequential labeling. In this way, the sub-cellular localization of all ten proteins could be established within the same image.

In these images, the Golgi can easily be identified as a perinuclear ribbon staining strongly for both ARFGAP1 and COP1 (magenta and purple respectively, Fig. S1I). In addition, there are bright puncta throughout the cell (pink arrow heads) stained with TFG and SEC31A, indicating that they are ERES. ARFGAP1 and COPI frequently also accumulated at ERES (made more obvious in the top-right inset, where TFG and SEC31 are removed), consistent with previous reports that COPI vesicles facilitate ER to Golgi transport in human cells29-31. GFP-ATG2A is mostly cytosolic, however small accumulations of ATG2 could be found at ERES and were generally observed to be proximal to ARFGAP1/COP1 (inset; brackets). Note that clusters of LC3B and p62 are also present. Occasionally, non-ERES accumulations of p62 could be observed in close proximity to LC3B and ATG2A (lower insets), and here again ATG2 is proximal to ARFGAP1.

Critically, we can directly measure the nominal proximity of each protein relative to the others under investigation by quantifying the median distance between all ten tested proteins within 200 nm of each individual protein (Fig. S1J). All early secretory pathway proteins closely associate with each other, including the COPI proteins ARFGAP1 and β-COP. As the majority of ATG2A was cytosolic, its average distances from other proteins were greater, however, within this cohort of candidate proteins, we found that ATG2A is most closely associated with p62 and ARFGAP1 (15nm and 16nm respectively; top row of Fig. S1J), exactly consistent with the APEX screen.

The APEX and FLASH-PAINT datasets provided a clear but static result that ATG2A is recruited to ARFGAP1 positive structures. To determine whether this recruitment was coincident with phagophore formation, we stably expressed GFP-ATG2A and TagBFP-LC3B in ATG2 DKO cells in order to identify phagophores. ATG2A primarily localizes to phagophores and lipid droplets, whereas LC3B primarily localizes to phagophores, autophagosomes and autolysosomes. Where these two proteins colocalize strongly suggests a growing phagophore in the process of membrane expansion (Figs. 1B). To assess co-localization with phagophores, we expressed candidate early secretory proteins labeled with red fluorophores (Fig. 1 C,D). ARFGAP1-TagRFP colocalized with ∼90% of actively forming phagophores, whereas only ∼30% of phagophores were coincident with the ERES markers mRuby-SEC23A or RFP-p125a/SEC23IP. Further, while the ARFGAP1 signal was largely concentrated in the Golgi, phagophores most frequently colocalized with extra-Golgi ARFGAP1 puncta.

Finally, we compare our results to the subcellular localization of ARFGAP1 established by global organelle profiling available at the database at https://organelles.sf.czbiohub.org32. ARFGAP1 is enriched on autophagic membranes positive for LC3B as well as endolysosomal and early secretory membranes (Fig. 1E). Strikingly, ARFGAP1 was most highly enriched on membranes isolated via RAB1A immunoprecipitation. The autophagic proteins ATG2A, ATG2B, WDR45, and ATG9A, all of which are involved in membrane expansion of the autophagosome, were also highly enriched on RAB1A membranes. These collective results establish that both endogenous ARFGAP1 (organelle isolation, APEX, FLASH PAINT) and overexpressed ARFGAP1 (live imaging) naturally associate with ATG2A, including during phagophore biogenesis.

Autophagic flux depends upon RAB1 but not ARFGAP1

To look for a potential functional role for ARFGAP1 in autophagy, we measured autophagic flux using the HaloTag dropout assay in which liberation of a free HaloTag (HT) is indicative of lysosome-mediated proteolysis of the full-length (FL) protein33. We stably expressed HaloTag7-mGFP-LC3B or HaloTag7-mGFP in WT HEK293 cells and then used siRNA to knockdown various regulators or potential regulators of macroautophagy initiation (Fig. 2A,B, S2A). Despite efficient knockdown, siRNA against ARFGAP1 had very little effect on flux; the ratio of free HaloTag to full-length HaloTag7-mGFP-LC3B following ARFGAP1 knockdown was similar to both nontransfected (NT) and nontargeting siRNA controls. In contrast, knockdown of the autophagy initiation factor FIP200 dramatically suppressed dropout of the HaloTag. We also looked for an interaction between ARFGAP1 and ATG2A, but even when both proteins were overexpressed, we could not detect significant interaction via co-immunoprecipitation (Fig. 3A,B). Thus, it is likely that ARFGAP1 localizes to a key membrane involved in autophagosome biogenesis but is not itself a player in initiating autophagosome formation. To examine whether the other isoforms of ARFGAP1 could redundantly contribute to autophagosome biogenesis, we knocked ARFGAP1, ARFGAP2, and ARFGAP3 down individually and collectively. Knockdown of ARFGAP2, ARFGAP3 and ARFGAP1-3 marginally increased autophagic flux (Fig. S2B,C), suggesting that the ARFGAP proteins could possibly act as negative regulators of autophagy.

Autophagic flux depends upon RAB1 but not ARFGAP1.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that siRNA KD of RAB1 inhibits autophagic flux. In gel fluorescence reveals HaloTag7s moiety bound to the TMR ligand. Indicated bands are full length (FL) HT-mGFP-LC3B protein and HaloTag7 after cleavage (HT) in the lysosome. The ratio between FL and HT demonstrates autophagic flux, quantified in (B). Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were collected for each condition. (C-D) Immunoblot of lysates from three biological replicates demonstrating that siRNA KD of RAB1 results in an accumulation of LC3B-II. The band intensity of LC3B-II was normalized against the intensity of GAPDH for each lane. To normalize each replicate, each individual value was divided by value of the non-transfected sample. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. (E-G) FACS data demonstrating that RAB1A/B KD results in a block in autophagic flux. H4 cells with a heterozygous GFP-mCherry-LC3B endogenous tag were transfected with siRNA for three days, following which they were starved in HBSS for four hours and analyzed by FACS. The ratio between GFP and mCherry is an indicator of autophagic flux, as the GFP signal is quenched in the lysosome. This ratio is displayed in histograms in (E) for one replicate and as averages in (F). The KDs of each protein can be seen in (G).

ATG2A interacts indirectly with RAB1A.

(A-B) Immunoblot showing CoIP between GFP-RAB1A and 3xFLAG-ATG2A compared to the same reaction with other GFP-tagged early secretory proteins. The cells were incubated with either complete media or EBSS for four hours prior to harvesting. For the CoIP, ∼3 mg of cell lysate was used per condition (slight deviations in total amount between replicates, but not between lanes). To quantify the results, the intensity of the 3xFLAG-ATG2A IP signal was ratioed against the input signal. To normalize each replicate, each individual value was divided by the average of the replicate. Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. The media was not a significant source of variation (p = 0.3535). (C-D) A fully reconstituted CoIP reaction between GFP tagged bait proteins and 3xFLAG-ATG2A reveals no direct interaction between RAB1A and ATG2A. GST tagged GFP-GABARAPL1 (GL1) and GFP-RAB1A were purified from E. Coli, during which process the GST tag was proteolytically removed. The purified proteins were then used in a CoIP reaction using GFP trap beads as in (A), but with no cell lysate present in the reaction. BSA was used to block the beads and as a negative control. To quantify the results, the intensity of the 3xFLAG-ATG2A IP signal was ratioed against the input signal. To normalize each replicate, each individual value was divided by the average of the replicate. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001.

As shown in (Fig. 1E), both the ATG2/ATG9 lipid transport apparatus as well as ARFGAP1 are associated to RAB1A-positive membranes. RAB1 proteins mediate ER to Golgi trafficking in the early secretory pathway. The yeast RAB1 homologue, Ypt1, is essential to autophagy34, 35, while in mammals, genetic depletion of either RAB1A or RAB1B modestly reduces autophagic flux36, 37. Thus, we tested the functional and physical interaction of RAB1 proteins with ATG2. Consistent with previous results, siRNA knockdown of either RAB1A or RAB1B in our system partially inhibited flux in the Halo dropout assay (Fig. 2A,B). Strikingly, knockdown of both isoforms simultaneously was nearly as effective as knocking down FIP200, establishing that RAB1 proteins play an essential role in supporting starvation induced autophagic flux. This result was consistent with but more pronounced than previous work examining LC3B flux following RAB1A/B double knock down38. As an orthogonal method to examine the role of RAB1 in LC3B flux, we used H4 neuroblastoma cells with a GFP-mCherry-LC3B heterozygous knock in reporter39.

Following knock down and starvation, we examined the ratio between GFP and mCherry by FACS. The low pH of the lysosome quenches the fluorescence of GFP at a higher rate than mCherry, resulting in a low ratio under starvation conditions as seen in the HBSS and siCNTL conditions (Fig. 2 E-G). In contrast, knock down of FIP200 or RAB1A/B resulted in an increased ratio, confirming that these proteins are necessary for proper autophagic flux. Of note, the RAB1A/B knockdown did not block flux as efficiently as the FIP200 KD. If Rab1 is essential for autophagy, this result suggests that either very few copies of RAB1 are needed to support autophagy or that another gene could be compensating for the lack of the two RAB1 isoforms.

In addition to following tagged LC3B, we also immunoblotted for endogenous LC3B in each of our siRNA experiments. We found that knockdown of either RAB1 isoform, and especially knockdown of both together leads to a clear accumulation of the membrane conjugated form of LC3B, LC3B-II (Fig. 2A,G). As LC3B overexpression can alter endogenous LC3B patterns, we confirmed this increase in a cell line expressing only HaloTag7-mGFP (Fig. 2C,D). Knockdown of RAB1A and RAB1B drove a threefold increase of LC3B-II (Fig. 2D). We also confirmed this phenotype in WT HEK293 cells in a classical LC3B flux assay with bafilomycin (Fig. S2D), where we found elevated LC3B-II levels in all conditions.

Increasing LC3B-II levels without autophagic flux is indicative of a block to macroautophagy at a late stage, downstream of initiation, and is therefore easily distinguished from the block induced by depleting FIP200 (where LC3B-II levels go down). Intriguingly, LC3B-II levels also increase when ATG2 proteins are genetically depleted (Fig. S1A,B), suggesting ATG2A and RAB1A might function concurrently. Indeed, overexpressed GFP-RAB1A could strongly immunoprecipitate flag-tagged ATG2A in our cells, indicating these proteins are in a shared complex (Fig. 2E,F). However, ATG2A and RAB1A (bound either to GDP or GTPγS) do not interact in a fully reconstituted system (Fig. 3C,D), suggesting that the interaction is indirect. In contrast, ATG2A and GABARAPL1 cleanly interact in the in vitro CoIP.

We also tested several other proximity hits from the biotinylation analysis for interaction with flag-tagged ATG2A, including GFP-tagged copies of ERES resident proteins SEC23A, SEC31A, SEC16A, Sar1 and all four homologs of Sec24 (Fig. S3 A-D). We found that ATG2A co-immunoprecipitated weakly with many of them, the strongest of which was SEC24D. Note that this SEC24D interaction was barely above the background of pull-down with GFP alone (Fig. 3A,B). Thus, consistent with proximity and imaging analyses, ATG2 proteins are in the general proximity of the ERES machinery, but protein-protein interactions appear to be limited to RAB1.

Disruption of the early secretory pathway results in the recruitment of autophagic proteins to RAB1A-positive tubular membranes

These results suggest that ATG2 engages an early secretory membrane, distinct from the ERES, in a RAB1-dependent manner. We next tested whether disruption of the early secretory system would impact the localization of ATG2A. To this end, we overexpressed SAR1BH79G, the catalytic dead mutant of the small GTPase SAR1B that recruits the COPII coat. SAR1BH79G binds the membrane and recruits the COPII coat, but is unable to release, thereby trapping COPII onto the membrane. Overexpression of this construct has been shown to reduce autophagy in starved cells17, 36. Consistent with those results, overexpression of SAR1BH79G-mOrange2 diminished autophagic flux of both a bulk cargo (HaloTag7-mGFP reduced ∼20%) and of LC3B (HaloTag7-mGFP-LC3B reduced by ∼30%) by the HaloTag degradation assay33 (Figs 4A,B). In a parallel assay following autophagic flux instead by the changing fluorescence of tandem tagged GFP-mCherry-LC3B, we detect a similar, but statistically insignificant trend. As both assays are likely sensitive to SAR1BH79G expression levels, which vary in our transient transfection system, we also tested how LC3 puncta accumulation correlates with SAR1BH79G expression on a per cell basis. LC3B puncta number were lowest in cells expressing high amounts of SAR1BH79G (Figs. S4A,B), consistent with a block to autophagosome formation when the secretory pathway is disrupted rather than a later impact on autophagosome clearance.

ATG2A, LC3B, ARFGAP1, and RAB1A localize to tubules following disruption of the early secretory pathway.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that SAR1BH79G-mOrange2 (mO) overexpression negatively affects autophagic flux. The ratio between the FL and the free HT represents the autophagic flux in each condition (as in Fig. 2 A/B) and is plotted in (B). Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from four biological replicates were collected for each of the HT-mGFP conditions and five replicates were collected for each of the HT-mGFP-LC3B conditions. (C) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2 that were transfected with SAR1BH79G-mOrange2 and starved for two hours. Note that GFP-ATG2A forms long tubules that do not colocalize with the SAR1BH79G-mOrange2 signal. Maximum intensity projection of a confocal image. Gray line was added to show cell outline. (D-E) Live cell imaging demonstrates the localization of ARFGAP1-TagRFP (D) and SNAP-RAB1A (E) on GFP-ATG2A/TagBFP-LC3B positive tubules following SAR1BH79G overexpression. Arrows denote thin GFP-ATG2A positive tubules whereas arrow heads denote thicker tubules that lack most GFP-ATG2A but contain high levels of LC3B, ARFGAP1, and RAB1A. Maximum intensity projections of confocal images. (F) CLEM-FIBSEM of cells prepared as in (C). GFP-ATG2A fluorescence corresponds to a thin tubular membrane that is connected to a thicker tubule. The small inset in the left image shows a single SEM slice. The turquoise arrow is pointing to the tubule which is segmented and displayed in turquoise in the larger image, whereas the yellow arrows are pointing at adjacent ER tubules which are segmented and displayed in the larger image in yellow. At many points, these ER structures come into very close apposition with the tubule which is denoted in red. The ER structures are removed in the right image for better visualization of the ATG2A positive tubule. Note that only the thin section has corresponding GFP-ATG2A fluorescence.

Unexpectedly, in a small subset of transfected cells (1.5±0.8%; calculated from the images in Fig. S4A,B), overexpression of SAR1BH79G under starvation conditions shifted the localization of GFP-ATG2A from a largely punctate distribution to long, tubular objects that were SAR1BH79G negative (Fig. 4C). GFP-ATG2A decorated these tubules with varying intensities approximating a beads-on-a-string pattern (e.g. arrows in Fig. 4D,E). Fluorescently tagged forms of LC3B, ARFGAP1 and RAB1A also decorated these tubules. However, these three proteins displayed the brightest intensities where ATG2A was dimmer (e.g. arrowheads in Fig. 4D,E). These three proteins, but not ATG2A, also decorated larger structures that appeared to be contiguous with the tubule, including bulbous regions at the tubule termini.

To determine the ultrastructure of these regions, we subjected GFP-ATG2A-expressing cells to correlative light and focused ion beam scanning electron microscopy (CLEM-FIB-SEM). The GFP-ATG2A fluorescence signal corresponded to thin, tubular membranes (Fig. 4F and S4D), which is consistent with the finding that ATG2A preferentially binds highly curved membranes40. The tubule is surrounded by ER elements, many of which appear to be in very close contact (Fig. 4F, S4C). GFP-ATG2A signal was more intense in regions with closely adjacent ER, consistent with a possible contact site localization. The tubule continued into larger structures at either end, which were GFP-ATG2-negative, and morphologically similar with the bulbous structures harboring ARFGAP1, RAB1A and LC3B in fluorescence images.

Thus, although unusual, these rare tubules afforded an opportunity to demonstrate that ATG2, ARFGAP1 and RAB1A target the same membrane, easily resolvable even at confocal resolution. ATG2 collects where ER is closely adjacent, and LC3B accumulates with RAB1A, implying a topological organization with LC3B and RAB1A decorating the same structure, suggesting that RAB1A membranes may be a component or progenitor of the phagophore membrane.

ARFGAP1 and RAB1A accumulate around pre-autophagic structures in ATG2 DKO cells

We previously published that very small, ATG9- and LC3B-positive vesicles accumulate in large clusters with the cargo-adaptor p62 in ATG2 DKO cells9. This cluster is surrounded by a shell of more complicated membranes which we termed “protophagophores” because they colocalize with other autophagy proteins such as WIPI2 and because many of them appear to adopt a cup-like shape in TEM images. Beyond this protophagophore shell, the entire structure is surrounded by ER (illustrated in Fig. 5A). Our work and others have shown that essentially all proteins involved in autophagosome biogenesis accumulate with these clusters, including FIP200, ATG14, ATG9A, ATG16L, and p629, 23, 24, suggesting they are repositories of autophagosome biogenesis material awaiting the arrival of ATG2 proteins. If early secretory membranes contribute directly in any way to the biogenesis event, they may also concentrate at these sites.

ATG2 deletion results in the accumulation of ARFGAP1 and RAB1A on autophagic membranes.

(A) Schematic of the membranes that accumulate at the ATG2 DKO compartment based off the TEM images in Olivas et al9. The core of the compartment contains small vesicles that stain for ATG9A, LC3B, and p62. The periphery of the compartment is comprised of more complicated cup shaped membranes that stain for WIPI2. The entire compartment is enwrapped by ER tubules. (B) Immunofluorescence images demonstrating the distribution of early secretory membranes around the ATG2 DKO compartment which is marked by either p62 (mouse antibody) or LC3B (rabbit antibody). The images presented are single confocal slices of the center of the ATG2 DKO compartment. The gray lines show the cell periphery and the insets focus on the largest accumulation of p62 or LC3B in the cell. (C) Quantification of the images in (C) and in Sup Fig 4A. The enrichment factor is equal to the mean of the given protein at the periphery of the ATG2 DKO compartment compared to its mean throughout the cell. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled together for each of the conditions. (D-E) Immunofluorescence images demonstrating the enrichment of overexpressed GFP-RAB1A (D) or SNAP-RAB1A (E) at the periphery of the ATG2 DKO compartment as determined by the presence of p62 (D) or WIPI2 (E). Presented as a single confocal slice. The gray lines show the periphery of the cell and the insets focus on ATG2 DKO compartments that are distal from the Golgi.

We therefore performed IF against various markers of the early secretory pathway in ATG2 DKO cells to look for colocalization with this compartment (represented by staining for p62 (mouse antibody) or for LC3B (rabbit antibody); Figs. 5B-D). We found that ER exit site markers such as SEC16A, SEC24C, SEC31A, and TFG were not enriched around the compartment, neither was the Golgi protein GM130 (Figs. 5B-C and S5A). ßCOP was modestly enriched, but at levels that by eye, did not suggest encapsulation of the compartment the way autophagy proteins collect. In addition, overexpressed mEmerald-ERGIC53 did not localize to the ATG2 DKO compartment (Fig. S5B). In contrast, ARFGAP1 targeted all of the ATG2 DKO compartments, forming a robust spherical shell at the periphery of the p62 signal in HEK293 ATG2 DKO cells (Fig. 4B-D), or permeating the entire compartment in HeLa ATG2 DKO cells41 (Fig. S5C).

We were unable to find a strong RAB1A signal by immunofluorescence, however, when overexpressed, tagged versions of RAB1A also formed a sphere around the HEK293 ATG2 DKO compartment (Fig. 5D. Both ARFGAP1 and SNAP-RAB1A colocalized with WIPI2 (Fig. 5E), consistent with a model whereby these membranes arrive at the PAS prior to ATG2A recruitment and subsequent autophagosome biogenesis.

ARFGAP1 accumulates at the ATG2 DKO compartment independently of the Golgi

The vast majority of ARFGAP1 is found at the Golgi, but ATG2A specifically localizes to extra-Golgi ARFGAP1 puncta (Fig. 1C, S1J). Consistent with a non-Golgi origin for these membranes, both natural and induced perturbations of Golgi structure have no impact on the recruitment of ARFGAP1 membranes to the p62-rich compartments in ATG2 DKO cells. First, in the course of these studies we observed that ARFGAP1 persistently colocalized with p62 in cells undergoing mitosis (Fig. 6A,B,C), even as the Golgi fragments and dissipates. Second, when cells were treated with brefeldin A to inhibit the formation of the Golgi42, ARFGAP1 again persistently colocalized with p62 (Fig. 6B,C). As brefeldin A specifically targets GBF1, the GEF for the ARF proteins43, this result also suggests that ARFGAP1 localization to the p62 compartment is independent of GTP bound ARF proteins. Third, overexpression of GFP-SidM/DrrA, an enzyme from Legionella Pneumophila which acts as both a GEF44, 45 and a GDF46 for RAB1, fragments the Golgi44 and sequesters RAB1 to the plasma membrane45. We found that following GFP-SidM expression in our cells, ARFGAP1 maintained colocalization with p62, even as the Golgi-associated ARFGAP1 as well as the Golgi-marker GM130 were completely dissipated (Fig. 6D, S6A). Thus, the ARFGAP1 localizing to the p62 compartment is not influenced by dispersal of the Golgi and does not colocalize with other Golgi markers, implying a non-Golgi origin for these structures.

ARFGAP1 accumulation at the ATG2 DKO compartment is independent of the Golgi.

(A) Immunofluorescence microscopy demonstrates that ARFGAP1 colocalizes with p62 in ATG2 DKO cells. Inset 1 highlights a cell undergoing mitosis, in which the colocalization of these two markers persists as denoted by the arrows. Inset 2 shows several cells, a ring-like accumulation of ARFGAP1 around p62 is clearly visible, separate from the Golgi. Maximum intensity projection of a confocal image. (B-C) Immunofluorescence images showing that perturbations which disrupt Golgi formation don’t affect the recruitment of ARFGAP1 to the ATG2 DKO compartment. HEK293 ATG2 DKO cells were treated or not treated with 5 μg/ml Brefeldin A for one hour, following which they were immunostained against ARFGAP1 and p62. The gray lines show the periphery of the cell and the insets focus on ATG2 DKO compartments as seen by the p62 accumulations. Colocalization between ARFGAP1 and p62 was assessed in (C) using the Pearson’s Correlation Coefficient. In the untreated sample, cells were separated into two groups: mitotic and non-mitotic, based on the ARFGAP1 fluorescence pattern. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. (D) Immunofluorescence image demonstrating that Golgi disruption by SidM/DrrA overexpression did not perturb the colocalization between ARFGAP1 and p62. HeLa ATG2 DKO cells were transfected with GFP-SidM and then immunostained against ARFGAP1 and p62. The gray line shows the periphery of the cell and the inset focuses on the largest accumulation of p62.

Finally, we tested whether the GAP domain of ARFGAP1 or the ALPS motif are necessary for its localization to the PAS, by comparing the colocalization of various tagged constructs with GFP-ATG2A and TagBFP-LC3B (Fig. S6A). Disruption of the GAP domain via the well characterized R50K mutation had no impact on its localization to the ATG2/LC3B positive phagophores (Fig. S6B). Mutation of the ALPS domain (L207D, V279D) altered ARFGAP1 distributions in the cell such that specific localizations could no longer be quantified (the redistribution to the cytoplasm created a high background that was difficult to properly threshold). Nonetheless, even in this condition there remained clear accumulations of this construct at phagophores indicating that recruitment is not strongly dependent upon the ALPS motif. As these elements of ARFGAP1 are essential to its Golgi function, these results further suggest that the ARFGAP1/RAB1A membranes involved in autophagosome biogenesis derive from early secretory membranes prior to the Golgi.

Discussion

Here we describe an interaction of ARFGAP1/RAB1-positive membranes with the autophagic biogenesis machinery including the lipid transporter ATG2. These membranes likely derive from a privileged region of the ERGIC consistent with the literature tying these intermediate secretory membranes to autophagy17, 21. In particular, the presence of both ARFGAP1 and RAB1 plus the general proximity to ERES proteins across a variety of assays is consistent with an early secretory origin for these membranes. Intriguingly, ARFGAP1 and RAB1 coalesce around the periphery of the ATG2 DKO compartment exactly where we previously described complex cup-shaped membranes, so called proto-phagophores9, accumulate. We do not detect any other ERES or ERGIC markers at these sites, implying proto-phagophores derive from only a subdomain of the ERGIC which we suspect is remodeled specifically for the purpose of supporting autophagy.

How do the ERGIC-derived RAB1 membranes contribute to autophagosome formation? Ypt1, the yeast homologue of RAB1A, was identified early on as a genetic regulator of autophagy34, and subsequently a number of studies have connected both Ypt1 and RAB1 to autophagosome biogenesis, but the precise mechanism of action remains contentious. In yeast, Ypt1 has been shown to interact with Atg135 and Atg1747, highly suggestive of a role in regulating or recruiting the kinase machinery driving initiation of autophagosome biogenesis. Human RAB1A binds the VPS34 complex favoring conformations that likely support PI3P formation48. Thus, either or both of the early acting complexes in autophagosome biogenesis may engage RAB1 proteins. Multiple studies have also demonstrated that depletion of RAB1 isoforms increases LC3B-II levels49,38, though with differing opinions on whether autophagy is promoted or inhibited in these backgrounds. In this study, we also find that RAB1 depletion elevates LC3B-II levels, phenocopying the change in LC3B-II that occurs when ATG2 is depleted. Further, we present two orthogonal and highly quantitative methods to follow LC3B degradation in the lysosome. Both methods unambiguously show a clear, albeit partial, genetic dependence of LC3B degradation upon RAB1 consistent with RAB1 promoting autophagy. As we detect RAB1A at the ATG2 DKO compartment, RAB1A recruitment appears to precede ATG2A recruitment to the PAS, consistent with an interaction with either the ULK1 or VPS34 complexes, but is not essential to support LC3B-II formation on early autophagy membranes.

We propose that ARFGAP1/RAB1 membranes are directly integrated into the phagophore prior to ATG2 recruitment. The fusion of vesicular sources has been proposed as a general mechanism for phagophore biogenesis50, 51. For example, the HyPas model postulates that FIP200 vesicles from the cis-Golgi fuse with endosomally derived ATG16L1 vesicles to create the nucleating structure for autophagosome formation22. Recent reports, including our own, have also identified ATG9 vesicles as a seed for autophagosome biogenesis, which expand through lipids delivered by ATG29, 10 and then scrambled into both leaflets by ATG96-8. An integration of these models is necessary. Geometric estimates of the surface-to-volume ratio in yeast autophagosomes, gleaned from cryo-electron microscopy tomograms, suggest that as much as 20% of the membrane might derive from fusion rather than lipid transport in order to deliver luminal volume52, which cannot be supplied by lipid transport. We therefore propose a model in which an early fusion step between ATG9A vesicles and ARFGAP1/RAB1 membranes, both of which accumulate in ATG2 DKO cells, primes the nucleating membrane for autophagosome biogenesis.

Materials and Methods

Plasmids and reagents

The following plasmids were used in this study: pLVX Puro GFP-ATG2A (Addgene 250863; Previously published3), pLVX Puro APEX2-GFP-ATG2A (Addgene 250864; This study; a PCR fragment containing flanking EcoRI cutsites around APEX2 was generated from pcDNA3 APEX2-NES (Addgene #49386) and was inserted into GFP-ATG2A by ligation), pCMV 3xFLAG-ATG2A (Addgene 250865; Previously published3), pLVX Neo TagBFP-LC3B (Addgene 250866; This study; pLVX puro TagBFP-LC3B3 and pLVX-EGFP-IRES-Neo (Addgene #128660) were cut with EcoRI and BamHI, following which the TagBFP-LC3B fragment was ligated into the pLVX Neo backbone), pDESt47 Sar1-GFP (Addgene #67409), SAR1B-mOrange2 (Addgene #166899), SAR1BH79G2-mOrange2 (Addgene #166900), SAR1BH79G-3xFLAG (Addgene 250867; This Study; A BamHI-3xFLAG-NotI fragment was generated by oligo annealing, following which it was ligated into the SAR1BH79G-mOrange2 backbone that was cut by the same restriction enzymes), SAR1B-BFP (Addgene 250872; This Study; A BamHI-BFP-NotI fragment was generated by PCR, following which it was ligated into the SAR1B-mOrange2 backbone that was cut by the same restriction enzymes), SAR1BH79G-BFP (Addgene 250873; This Study; A BamHI-BFP-NotI fragment was generated by PCR, following which it was ligated into the SAR1BH79G-mOrange2 backbone that was cut by the same restriction enzymes), mRuby-SEC23A (Addgene 36158), EGFP-SEC23A (Addgene #66609), pEGFP-SEC24A (Addgene 250868; This Study; a XhoI-SEC24A-PacI fragment was generated by PCR from HA-SEC24A which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), pEGFP-SEC24B (Addgene 250869; This Study; a XhoI-SEC24B-PacI fragment was generated by PCR from HA-SEC24B which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), EGFP-SEC24C (Addgene 250870; This Study; a AgeI-EGFP-XhoI fragment was generated by PCR and was ligated into the backbone of pEYFP-SEC24C (Addgene #66608) which was cut by the same restriction enzymes), pEGFP-SEC24D (Addgene #32678), GFP-SEC31A (Addgene #42124), GFP-SEC16A (Addgene #36155), ARFGAP1-TagRFP (This Study; ARFGAP1-TagRFP was a gift from the Rothman lab. A frame shift and a point mutation were corrected by site directed mutagenesis), ARFGAP1-GFP (Gift from the Antonny Lab), pMRX IB HaloTag7-mGFP (Addgene # 184903), pMRX No HaloTag7-mGFP-LC3B (Addgene # 184901), pLVX GFP (Addgene 250871; This Study; a fragment containing EGFP was generated from pLVX GFP-ATG2A and was ligated into pLVX puro (Clonetech, 632159) that was cut with EcoRI using Gibson Assembly), GFP-RAB1A (Gift from the Rothman Lab), SNAP-RAB1A (Gift from the Rothman Lab), pGEX2T GST-GFP-RAB1A (Addgene 250877; this study; a GFP-RAB1A fragment was generated from the GFP-RAB1A plasmid from the Rothman lab and inserted in to the pGEX2T backbone), and pGEX2T GST-GFP-GL1 (Addgene 2508740).

Antibodies for immunoblotting were used at 1:1000 and include the following: anti-ATG2A (Cell Signaling Technology, 15011S), anti-p62 (BD BioSciences, 610832), anti-GAPDH (Abcam, ab9484), anti-LC3B (Cell Signaling Technology, 3868S), anti-RB1CC1 (Thermo Fisher Scientific, 172501AP), anti-RAB1A (Cell Signaling Technology, D3X9S), anti-RAB1B (Proteintech, 17824-1-AP), anti-ARFGAP1 (Proteintech, 13571-1-AP), anti-FLAG (Sigma-Aldrich, F1804-200UG), anti-GFP (Cell Signaling Technology, 2956S), anti-mCherry (Invitrogen, PA5-34974), ECL anti-rabbit IgG horseradish peroxidase-linked (GE Healthcare, NA934V), and ECL anti-mouse IgG horseradish peroxidase-linked (GE Healthcare, NA931V).

Antibodies for immunofluorescence were used at 1:500 and include the following: anti-Biotin (Rockland, 200-301098S), anti-GFP nanobody (Massive Photonics), anti-ARFGAP1 (Proteintech, 13571-1-AP;), anti-SEC31A (BD Biosciences, 611280), anti-SEC24C (Cell Signaling Technology, 14676S), anti-MIA3 (Tango1; Sigma, HPA055922), anti-TFG (Abcam, ab156866), anti-COPI (CMIA10; made by Rothman Lab), anti-LC3B (MBL, PM036), anti-p62 (BD BioSciences, 610832), anti-SEC16A (Abcam, ab70722), anti-SEC24D (Cell Signaling Technology, 14687S), anti-GM130 (BD Biosciences, 610823), anti-ERGIC53 (Abcam, EPR6979), Alexa Fluor 405 goat anti-rabbit IgG (Thermo Fisher Scientific, A31556), Alexa Fluor 405 goat anti-mouse IgG (Thermo Fisher Scientific, A31553), Alexa Fluor 594 goat anti-rabbit IgG (Thermo Fisher Scientific, A11037), Alexa Fluor 647 goat anti-rabbit IgG (Thermo Fisher Scientific, A21245), and Alexa Fluor 647 goat anti-mouse IgG (Thermo Fisher Scientific, A21236).

The following reagents were purchased for FLASH-PAINT: Cy3B-modified DNA oligonucleotides were custom-ordered from IDT. Sodium chloride 5 M (cat: AM9759) were obtained from Ambion. Ultrapure water (cat: 10977-015) was purchased from Invitrogen. µ-Slide 8-well chambers (cat: 80807) were purchased from ibidi. Methanol (cat: 9070-05) was purchased from J.T. Baker. Glycerol (cat: G5516-500ml), protocatechuate 3,4-dioxygenase pseudomonas (PCD) (cat: P8279), 3,4-dihydroxybenzoic acid (PCA) (cat: 37580-25G-F) and (+−)-6-hydroxy-2,5,7,8- tetra-methylchromane-2-carboxylic acid (Trolox) (cat: 238813-5 G) were ordered from Sigma. 1× Phosphate Buffered Saline (PBS) pH 7.2 (cat: 10010-023) was purchased from Gibco. Paraformaldehyde (cat: 15710) were obtained from Electron Microscopy Sciences. Bovine serum albumin (cat: 001-000-162) was ordered from Jackson ImmunoResearch. Triton X-100 (cat: T8787-50ML) was purchased from Sigma-Aldrich. DNA labeled Nanobodies were obtained from Massive Photonics.

SiControl, FIP200, RAB1A, RAB1B, and ARFGAP1 were acquired from Horizon Discovery (D-001810-10-05, L-0082383-00-0005, L-008958-01-0005, L-013321-02-005; Horizon Discovery).

Cell culture

HEK293 cells (ATCC, CRL_1573), H4 (CVCL_1239), and HeLa (CVCL_0030) cells were cultured at 37°C and 5% CO2 in DMEM (Thermo Fisher Scientific, 11965092) supplemented with 10% FBS (Thermo Fisher Scientific, 10438062) and 1% penicillinstreptomycin (Thermo Fisher Scientific, 15140122). Cells were starved by incubation in Earle’s Balanced Salt Solution (Thermo Fisher Scientific, 24010043) for biochemical assays or in Hank’s Balanced Salt Solution (Thermo Fisher Scientific, 14025092) for live cell imaging assays and treated with 0.1 µM bafilomycin A1 (Enzo, BML-CM110-0100) for 2 h.

Lentivirus production and transduction

HEK293 cells were seeded into a 10-cm plate. At 70% confluence, cells were transfected with 4.3 µg psPAX2 (Addgene #12260), 0.43 µg pCMV-VSV-G (Addgene, #8454), and 4.3 µg target plasmid using 20 μl Lipofectamine 3000 (Thermo Fisher Scientific, L3000008). DNA was added to 500 μl Opti-MEM (Thermo Fisher Scientific, 31985070) and Lipofectamine 3000 to 500 μl Opti-MEM in separate 1.5 ml microcentrifuge tubes. The tubes were then mixed and left for 10 min before adding the mixture dropwise to cells in fresh DMEM. After overnight incubation at 37°C, the medium was replaced with fresh DMEM. Cell medium was then collected after 24-48 h and filtered with a 0.45-µm syringe filter (Pall, 4184). Generated virus was either used immediately or was aliquoted and stored at −80°C. HEK293 cells to be transduced were seeded into a 6-well plate. Undiluted virus was added to cells and incubated at 37°C for 24 h. The medium was replaced with fresh DMEM and left for another 24 h. The cells were then treated with 2 µg/ml puromycin (Clontech, 631306), 2 mg/ml Geneticin (Thermo Fisher Scientific, 10131035), or 10 µg/ml Blasticidin (Gibco, A11139-03) for at least 2 wk. For the case of pMRX No HaloTag7-mGFP-LC3B, which lacked a selection marker, the cells were transduced and subsequently sorted by flow assisted cell sorting to normalize GFP expression levels.

Proximity labeling

Proximity labeling was performed as previously described27. ATG2 DKO cells stably expressing APEX2-EGFP-ATG2A were cultured for two weeks in media containing heavy, medium, or light isotopes of arginine and lysine (Cambridge Isotope Laboratories, CLM-2265-H-PK, DLM-2640-PK, CNLM-291-H-PK; Sigma-Aldrich, 608033). All conditions were expanded to 3x15cm plates prior to proximity labeling. Heavy labeled cells were starved in EBSS and treated with 100 nM Bafilomycin A1 (Enzo, BML-CM110-0100) for 2 hrs prior to labeling. All conditions were incubated with 500 μM Biotin Phenol (Biotinyl Tyramide; Sigma-Aldrich, SML2135) for 30 min at 37°C. The heavy and medium isotope conditions were treated with 1 mM H2O2 (Sigma-Aldrich, H1009-100ML) for 1 min at room temperature. All conditions were then aspirated and scraped in quencher solution containing 10 mM sodium azide (Sigma-Aldrich, S2002-100G), 10 mM sodium ascorbate (Sigma-Aldrich, A4034-100G) and 5 mM Trolox (Sigma-Aldrich, 238813-5G) in DPBS. The cells were then spun down at 77xg for 5 min at 4°C and were resuspended in quencher solution. This process was repeated for a total of three washes. The cells were then lysed in RIPA buffer (50 mM Tris, 150 mM NaCl, 0.1% SDS, 0.5% sodium deoxycholate (Sigma-Aldrich, D6750-10G), 1% NP40 (Roche, 11754599001) and 1x protease inhibitor (Roche, 11836170001)) for 30 min prior to centrifugation for 10 min at 18,407xg at 4°C. The supernatants were collected and protein concentration was assessed via bradford assay. Heavy, medium and light cells were mixed at a 1:1:1 ratio for a combined mass of 3-4 mg of protein. The protein mixtures were subsequently incubated with streptavidin magnetic beads (Pierce, 88817) for 2 hrs at 4°C after which they were washed twice with RIPA lysis buffer, once with 1 M KCl, once with 0.1 M Na2CO3, once with 2 M urea in 10 mM Tris (pH 8.0) and twice with RIPA lysis buffer for each sample. All washes were performed at 4°C. The proteins were eluted off of the beads by boiling the samples in 2x LDS sample buffer (Invitrogen, NP0007). The eluted fraction was run on a gel (Invitrogen, NP0341) for a 10 minutes. The resulting gel plug was stained (Invitrogen, 46-5034) and excised and sent to the Yale Keck facility for mass spectrometry analysis.

Mass spectrometry

MS was performed at Yale Mass Spectrometry & Proteomics Resource of the W.M. Keck Foundation Biotechnology Resource Laboratory. Samples were analyzed in biological triplicates. The data were normalized to internal controls and total spectral counts. For differential analysis between the two groups 10E-04 to was added to all values, the data was log2-transformed, and an unpaired Student t-test was performed. Differentially abundant proteins were considered those with an absolute fold change > 1.5 and p <0.05.

FLASH-PAINT Imaging

Cells were fixed with 4% PFA (Electron Microscopy Services, 15710-S) for 30 min. After three washes in 1xPBS for five minutes each, cells were blocked and permeabilized with 3% BSA and 0.25% Triton X-100 at room temperature for 1 h. Next, cells were incubated with the anti-ARFGAP1 antibody, anti- SEC31A antibody, and the GFP-Nanobody in 3% BSA and 0.1% Triton X-100 at 4 °C overnight. Additionally, all other primary antibodies were pre incubated with the corresponding nanobodies (Tables S1) at 4 °C overnight. The next day, after four washes (30 s, 60 s, 2× 5 min) cells were incubated with the nanobodies corresponding to ARFGAP1 antibody and SEC31A antibody for ∼2 h at room temperature. Next, unlabeled excess secondary nanobodies (to block unlabeled epitopes) were added to pre-incubation antibody-nanobody mixes at room temperature for 5 min. Next, the cells were incubated with the pooled antibody-nanobody mix for ∼2.5 h at room temperature. After four washes (30 s, 60 s, 2× 5 min), the sample was post-fixed with 3% PFA and 0.1% GA for 10 min. Finally, samples were washed three times with 1× PBS for 5 min each before adding the imaging solution.

The imaging solution consisted of 1x PBS and 500 mM NaCl and was supplemented with 1× Trolox, (100× Trolox: 100 mg Trolox, 430 μL 100 % Methanol, 345 μL 1M NaOH in 3.2 mL H2O), 1× PCA (40× PCA: 154 mg PCA, 10 mL water and NaOH were mixed and pH was adjusted 9.0), and 1× PCD (100× PCD: 9.3 mg PCD, 13.3 mL of buffer (100 mM Tris-HCl pH 8, 50 mM KCl, 1 mM EDTA, 50 % Glycerol)). All three stock buffers were frozen and stored at -20 °C.

Fluorescence imaging was carried out on an inverted Nikon Eclipse Ti2 microscope (Nikon Instruments) with the Perfect Focus System, attached to a Andor Dragonfly unit. The Dragonfly was used in the BTIRF mode, applying an objective-type TIRF configuration with an oil-immersion objective (Nikon Instruments, Apo SR TIRF 60×, NA 1.49, Oil). As excitation laser, a 561 nm (1W nominal) was used. The beam was coupled into a multimode fiber going through the Andor Borealis unit reshaping the beam from a gaussian profile to a homogenous flat top. As dichroic mirror a CR-DFLY-DMQD-01 was used. Fluorescence light was spectrally filtered with an emission filter (TR-DFLY-F600-050) and imaged on a scientific complementary metal oxide semiconductor (sCMOS) camera (Sona 4BV6X, Andor Technologies) without further magnification, resulting in an effective pixel size of 108 nm. The power at the objective lens was ∼10 % of the power set at the laser.

Imaging was carried out using the corresponding Imager and Adapter (Table S2, S3, S5) in imaging buffer. 25,000 frames were acquired at 25 ms exposure time. The readout bandwidth was set to 540 MHz. Laser power (@561 nm) was set to 80 mW (measured before the back focal plane (BFP) of the objective), corresponding to ∼1.8 kW/cm2 at the sample plane. After imaging the sample was subsequently washed three times with 200 µL each with 1× PBS (on the microscope) followed by an incubation with the corresponding eraser (Table S4; at 20 nM) for ∼ 5 min. This process was repeated iteratively for ten sequential rounds (Table S5).

DNA-PAINT data was reconstructed, postprocessed (drift correction and alignment of imaging rounds) and rendered with the Picasso package53.

Focused ion beam scanning electron microscopy and correlative light electron microscopy

For focused ion beam scanning electron microscopy (FIBSEM) CLEM, ATG2 DKO HEK293 cells overexpressing GFP-ATG2A were plated on 35-mm MatTek dish (MatTek, P35G-1.5-14-CGRD). The cells were starved for two hours prior to pre-fixation in 4% PFA + 0.25% glutaraldehyde and three washes in 1x PBS. Regions of interest by fluorescence light microscopy were selected and their coordinates on the dish were identified using phase contrast. Cells were further fixed with 2.5% glutaraldehyde in 0.1 M sodium cacodylate buffer, postfixed in 2% OsO4 and 1.5% K4Fe(CN)6 (Sigma-Aldrich) in 0.1 M sodium cacodylate buffer, en bloc stained with 2% aqueous uranyl acetate, dehydrated, embedded in Embed 812, and polymerized at 60°C for 48 hrs. Cells of interest were relocated based on the pre-recorded coordinates. Epon block was glued onto the SEM sample mounting aluminum stub and platinum en bloc coating on the sample surface was carried out with the sputter coater (Ted Pella, Inc.). The cell of interest was relocated under SEM imaging based on pre-recorded coordinates and FIB-SEM imaged in a Crossbeam 550 FIB-SEM workstation operating under SmartSEM (Carl Zeiss Microscopy GmbH) and Atlas 5 engine (Fibics, Inc.). The imaging resolution was set at 7 nm/pixel in the X, Y axis with milling being performed at 7 nm/step along the Z axis to achieve an isotropic resolution of 7 nm/voxel. Images were aligned and exported with Atlas 5 (Fibics, Inc.), further processed and segmented with DragonFly Pro software (Object Research Systems [ORS], Inc.).

Expi293 culture and transfection

3xFLAG-ATG2A was expressed and purified from Expi293 suspension cells. Cells were cultured at 37°C, 8% CO2 in vented PETG flasks (Thermo Scientific, Nalgene; 4115-0500) with 125 rpm rotation. Prior to transfection, cells were grown to 3-5 x 106 cells/mL then diluted to 2.5x106 cells/mL and allowed to grow for 18- 24 hours. Upon reaching 4.5-5.5 x 106 cells/mL, the culture was diluted to 3 x 106 cells/mL with fresh media. Cells were transfected with 100 μg plasmid per 100 mL of cultured cells. For every 100 mL of cells, 100 μg DNA was diluted in 5 mL OptiMEM. 300 µL of 1 mg/mL polyethylenimine linear 25k (Polyscience, 23966) was diluted into 5 mL OptiMEM in a separate tube. Both tubes were incubated for 5 minutes at room temperature then combined and further incubated for 30 minutes. The DNA/PEI mixture was then added dropwise into Expi293 cultures. 12-18 hours after transfection, protein expression was enhanced using 700 μL of 500 mM valproic acid. Cells were then harvested after 48-72 hours.

3xFLAG-ATG2A purification

Cell pellets were resuspended in 10 mL purification buffer A (500 mM NaCl, 20 mM HEPES, pH: 8.0, 10% glycerol, 1 mM TCEP) for every 100 mL of original culture volume. Purification buffer A was supplemented with 1x EDTA-free protease inhibitor cocktail (Roche; 11873580001). Cells were then lysed by subjecting pellet resuspensions to 6 freeze/thaw cycles in liquid nitrogen The lysates were then clarified by centrifugation at 18,500 x G, and supernatants were transferred to 150 μL of anti-Flag M2 agarose resin per 100 mL of initial culture volume (prewashed with 10 mL purification buffer A). Lysates were incubated with beads under mild agitation for 2 hours at 4 °C. Lysates and beads were then passed through a gravity filtration column and washed with 4 CVs of purification buffer. Beads were then incubated with 2.5 mM ATP and 5 mM Mg2+ in purification buffer overnight at 4 °C. Columns were then washed with 2 CVs of purification buffer, followed by elution with 50 μL of 1 mg/mL 3x-Flag (DYKDDDDK) peptide (ApexBio Technology; A6001) per 100 mL of initial culture volume. Protein concentration was determined against a BSA standard curve by Coomassie staining prior to subsequent assays. Proteins were flash frozen in liquid nitrogen for storage at -80 °C.

Recombinant protein expression and purification from E. Coli

For all in vitro experiments, GFP-GL1 refers to protein expressed and purified from constructs ending at the reactive glycine (G116) of GL1, as this construct was previously used for lipidation experiments in the absence of ATG4 proteins. In summary, human GFP-GL1 and GFP-RAB1A were cloned in the pGEX-2T GST vector. The proteins were expressed in BL21-Gold (DE3) competent cells (Agilent Technologies; 230132). The cells were cultured in 4 liters of Luria Bertani broth (LB) medium with 50 mg/ml carbenicillin (C1389; Sigma-Aldrich) and induced with 0.5 mM isopropyl β-D-thiogalactopyranoside (IPTG). Bacterial pellets were treated with EDTA-free protease inhibitor tablets in thrombin buffer (20 mM Tris pH 7.5, 100 mM NaCl, 5 mM MgCl2, 2 mM CaCl2, 1 mM DTT). The cells were broken in a cell disrupter, and the lysate was incubated with pre-washed glutathione agarose beads (G4501; Sigma-Aldrich) for 3 h at 4°C. The beads were washed three times and then incubated with thrombin (10 μl of thrombin [T6884; Sigma-Aldrich], 500 μl of thrombin buffer, 0.5 μl of DTT, 500 μl of beads) to cut the proteins from GST tags overnight at 4°C. Purified proteins were stored in 20% glycerol at −80°C.

Lysis, gel electrophoresis, and immunoblotting

Cells were collected by scraping in PBS. The cells were centrifuged at 500 g for 3 min at 4°C, and lysed in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor cocktail (Sigma-Aldrich, 11873580001) on ice for five minutes. Lysates were collected after centrifugation at 16,000 g for 10 min at 4°C. Total protein concentration was determined by a bradford assay (BioRad, 5000006). 50-100 μM DTT (American Bio, AB00490-00010) and 1xLDS (Invitrogen, NP0007) loading buffer were added to each sample, which were then electrophoresed on precast Bis-Tris gels (Thermo Fisher Scientific, NP0341BOX or NP0322BOX). Gels were subsequently transferred to Immobilon-FL PVDF membranes (Sigma-Aldrich, IPFL00010) at 30-35 V for 90 min.

The membranes were blocked with 5% BSA (Sigma-Aldrich, A9647) in PBS-T (PBS containing 0.1% Tween 20; Thermo Scientific, CAS 9005-64-5) for 1 h at room temperature and incubated with primary antibody (diluted 1:1,000 in PBS-T containing 5% BSA [Sigma-Aldrich, A9647] and 0.03% sodium azide (Sigma-Aldrich, S2002)) overnight at 4°C. Membranes were washed three times in PBS-T before incubation with the secondary antibody which was diluted 1:5,000 in PBS-T containing 0.5% Omniblock (American Bio, AB10109-00100) for 1 h. Membranes were then washed three times in PBST and treated with SuperSignal West Femto substrate (Thermo Fisher Scientific, 34096) for 5 min before imaging with the Bio-Rad VersaDoc imaging system.

Immunoprecipitation

For immunoprecipitation experiments, 15 cm plates were seeded and transfected upon reaching 80-90% confluency. The next day, the cells were collected by scraping, washed 1x in PBS, and lysed as previously described. Following quantification of protein levels by a Bradford assay, equal amounts of protein were added to pre-washed GFP-Trap beads (Proteintech, gtma-20). The volume was increased to 500 uL and the bead/lysate slurry was incubated at 4°C for two hours while rotating. The beads were pelleted using a magnetic rack and washed five times in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor. The target proteins were eluted by boiling the beads for 10 min in 2x LDS loading buffer.

Densitometry and quantification

Densitometry quantifications of immunoblots and in-gel fluorescence images were performed using ImageJ software. For LC3B and p62 clearance analysis (Fig. S1 A-C), the band intensity of LC3B or p62 for each condition was normalized against the GAPDH intensity of that lane. To normalize each replicate, each value was divided by the mean value of the entire replicate. This normalization technique was also employed in the co-immunoprecipitation experiments. Statistical significance was determined by two-way ANOVA followed by a Tukey multiple comparison test to compare cell types. To quantify the co-immunoprecipitation reactions, the intensity of the 3xFLAG-ATG2A IP band was ratioed against the intensity of its corresponding input band. Following normalization, statistical significance was determined by a one (Fig. S3 A-D) or two-way (Fig. 2E-F) ANOVA followed by a Dunnett multiple comparison test to comparing each sample to the control lane. For the HaloTag flux assay (Fig. 2A, 2B, 3A, 3B), the Halotag flux rate was calculated as the ratio of the free Halotag band divided by the unprocessed Halotag band. Statistical significance was determined by a one (Fig. 2 A and B) or two-way (Fig 3 A and B) ANOVA followed by a Dunnett multiple comparison test to comparing each sample to the control lane. For one way ANOVA tests, data distributions were assumed to be normal. For two way ANOVA tests, the Geisser-Greenhouse correction was used as equal variability was not assumed. All data are n = 3 or greater and represent biological replicates. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001. All blots were adjusted in contrast for better visualization of the bands in the figures.

LC3B flux analysis

WT, ATG2 DKO and ATG2 DKO + APEX2-GFP-ATG2A HEK293 cells were cultured in 6-well plates to 70-90% confluence. At the beginning of the assay, the media was replaced with fresh DMEM, DMEM containing 100 nM bafilomycin A1, or EBSS containing 100 nM bafilomycin A1. After a 2 hour incubation, cells were harvested by scraping, washed with 1x PBS, and lysed. 20 μg of protein from each sample were analyzed by SDS-PAGE and immunoblotting for ATG2A, p62, GAPDH, and LC3B.

HaloTag flux assay

HEK293 WT stably expressing either HaloTag7-mGFP or HaloTag7-mGFP-LC3B were cultured in 6-well plates to 60-70% confluence prior to transfection with either SAR1B/SAR1BH79G-mOrange2 for one day (Fig. 3A) or siRNA for two days (Fig. 2A). SAR1B transfection was performed using Lipofectamine 3000 (Thermofisher Scientific, L3000015) according to the manufacturer’s protocol. SiRNA transfection was performed using Dharmafect (Horizon Discovery, T-2001-01) according to the manufacturer’s protocol. The cells were incubated with 100 nm TMR-conjugated HaloTag ligand (Promega, G8251) for 30 min. Following two washes with 1xPBS, the cells were starved in EBSS for four hours. The cells were then collected and lysed. Following SDS-PAGE separation, in gel fluorescence was immediately measured with a ChemiDoc MP Imaging System (BioRad). Images were taken at 30 s and 100 s exposures. The gel was then transferred and immunoblotted against mCherry and GAPDH (Fig 3A) or FIP200, RAB1A, RAB1B, ARFGAP1, GAPDH, and LC3B (Fig 2A).

GFP-mCherry-LC3B Flux analysis by FACS

H4 cells and H4 cells with a GFP-mCherry-LC3B heterozygous knock-in tag were seeded in a six well plate. Following a three day siRNA knock down, the cells were subject to a four hour starvation period in HBSS. They were subsequently trypsinized, pelleted, resuspended in HBSS, and passed through a filter prior to FACS data acquisition. The cell population was gated to ensure a single, live cell population. The ratio between the GFP and mCherry intensity was plotted as a histogram using FlowJo 10 and the ratio between the average values of each replicate was plotted using Prism 10. Three biological replicates were combined and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

Immunofluorescence

Cells seeded on coverslips were fixed in 4% PFA (Electron Microscopy Sciences, 15710), washed three times with PBS, and permeabilized in PBS containing 0.1% Saponin (Sigma-Aldrich, 47036) and 3% BSA. Incubation with primary antibodies was done at 4C for 18-24 hrs and secondary antibodies at room temperature for 1 hr at indicated concentrations (see Reagents). After each antibody incubation, cells were washed three times in PBS containing 0.1% Saponin and 3% BSA. The coverslips were mounted on precleaned microscope slides with Fluoromont-G mounting reagent (Southern Biotech, 0100-01). Images were acquired on an inverted Zeiss LSM 880 laser scanning confocal microscope with AiryScan, using Zen acquisition software (Fig S1D) or a Nikon CSU-W1 SoRa 60x oil objective, using the Nikon Elements software (Fig. 1C, 3C-F, S3A, S3C, 4A, 4C, 4E, 4F). All images were processed using ImageJ software. Images are displayed either as maximum intensity projections or single confocal slices as stated in the figure legends.

To assess the enrichment factor of early secretory proteins around the ATG2 DKO compartment (Fig. 4D), the phagophore channel which was denoted by either LC3B or p62 was masked and duplicated. Following outlier removal, one masked image was dilated and the other eroded. Following an XOR function, the resulting signal formed a ring around the ATG2 DKO compartment and the early secretory protein channel(s) were measured using this ROI. Separately, the early secretory protein channel was masked using a manual threshold and measured. The enrichment factor was calculated as (Meancompartment-Mincell)/(Meancell-Mincell). Three biological replicates were combined and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

Live cell imaging

Cells were plated on poly-d-lysine coated MatTek dishes (MatTek, P35GC-1.5-14-C). The following day the cells were transfected using Lipofectamine 3000 (Thermo Fisher Scientific, L3000015) according to the manufacturer’s protocol for 18-24 hrs. The cells were starved in HBSS (Thermo Fisher Scientific, 14025092) for 2-4 h and imaged in the same solution at 37°C. All images were analyzed using ImageJ and are displayed as maximum intensity projections.

For phagophore quantification (Fig. 1 C and D), each channel was masked using a manual threshold and outliers were removed. The ATG2A and LC3B channels were merged using the AND function and quantified as the total number of phagophores. The phagophore channel was then merged using the AND function with the transfected gene and quantified as the number of positive phagophores. Three biological replicates were combined and the ratio of positive over total was plotted and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values were considering adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

eLife Assessment

This study provides important insights into the crosstalk between ATG2A with components of the early secretory pathway. While the mechanisms governing autophagic membrane expansion remain yet to be fully understood, in this study the authors employ an elegant proximity labelling approach and identify two ER-Golgi intermediate compartment (ERGIC)-localized proteins. Through a series of complementary experiments combining microscopy and biochemistry, the authors identify ARFGAP1 and Rab1A as components of early autophagic membranes, which accumulate at the periphery of pre-autophagosomal structures induced by loss of ATG2. The overall study is well executed and the evidence supporting the claims is convincing.

Reviewer #1 (Public review):

Summary:

D. Fuller et al. set out to study the molecular partners that cooperate with ATG2A, a lipid transfer protein essential for phagophore elongation, during the process of autophagy. Through a series of experiments combining microscopy and biochemistry, the authors identify ARFGAP1 and Rab1A as components of early autophagic membranes, which accumulate at the periphery of aberrant pre-autophagosomal structures induced by loss of ATG2. While ARFGAP1 has no apparent function in autophagy, the authors show that RAB1A is implicated in autophagy, although the precise mechanisms are not explored in the manuscript.

Strengths:

The work presented by Fuller et al. provides new insights into the composition of early autophagic membranes. The authors provide a series of MS experiments identifying proteins in close proximity to ATG2A, which is a valuable dataset for the field. Furthermore, they show for the first time the interaction between ATG2A and RAB1A both in fed and starved conditions, which extends the characterisation of the pre-autophagosomal structures observed in ATG2 DKO cells.

Weaknesses / Specific comments:

(1) The authors claim that Rab1A/B knockdown phenocopies the LC3-II accumulation observed in ATG2 DKO cells. While LC3-II accumulation is consistent with this interpretation, depletion of many autophagy-related proteins can give rise to a similar phenotype, even when they function at distinct stages of the autophagic cascade. Therefore, LC3-II accumulation alone is insufficient to support phenocopying in my vew. Immunofluorescence analyses demonstrating comparable cellular phenotypes-such as membrane accumulation of pre-autophagosomal structures-following Rab1 knockdown should be provided. Moreover, p62 does not accumulate upon Rab1 depletion, suggesting that loss of Rab1 does not fully phenocopy ATG2 deficiency. Consequently, it remains unclear whether Rab1A depletion truly phenocopies ATG2A depletion with respect to autophagy progression or the accumulation of pre-autophagosomal structures.

(2) Interpretation of the significance of the data

(2.1) The significance statement asserts that "this study elucidates the role of early secretory membranes in autophagosome biogenesis." While the data convincingly demonstrate an association between the RAB1A GTPase and ATG2A, the study does not provide mechanistic insight into how this interaction functionally contributes to autophagy. As presented, the findings support a correlative relationship rather than a defined role in autophagosome biogenesis.

(2.2) The title states that ATG2A "engages" Rab1A- and ARFGAP1-positive membranes during autophagosome formation. However, both Rab1A and ARFGAP1 are shown to localize to pre-autophagosomal structures independently of ATG2A. In the absence of evidence demonstrating a functional or causal dependency, the term "engages" appears overstated. A more descriptive term, such as "associates," would more accurately reflect the data.

(2.3) In the Discussion, the authors state that previous studies have reported increased LC3-II levels following knockdown of Rab1 proteins (refs. 38 and 49). However, it is unclear where this observation is documented in the cited references.

(3) Some concerns remain in specific figures, as outlined below:

• Quantification is missing in Fig S2D.

• The authors claim: "siRNA against ARFGAP1 had very little effect" but the quantification and blots show actually no effect.

• Conclusions drawn from KD experiments in Fig. S2 should be interpreted with caution, as knockdown efficiency is very low, particularly for ARFGAP1/3 in the triple knockdown.

• In New Fig. 4, the representative blot is not representative of the results showed in the quantification as previously noted.

Reviewer #2 (Public review):

The mechanisms governing autophagic membrane expansion remain incompletely understood. ATG2 is known to function as a lipid transfer protein critical for this process; however, how ATG2 is coordinated with the broader autophagic machinery and endomembrane systems has remained elusive. In this study, the authors employ an elegant proximity labeling approach and identify two ER-Golgi intermediate compartment (ERGIC)-localized proteins-Rab1 and ARFGAP1-as novel regulators of ATG2 during autophagic membrane expansion.

Their findings support a model in which autophagosome formation occurs within a specialized subdomain of the ER that is enriched in both ER exit sites (ERES) and ERGIC, providing valuable mechanistic insight. The overall study is well executed and offers an important contribution to our understanding of autophagy. I support its publication in eLife and offer the following minor comments for clarification and improvement.

Specific Comments

(1) Integration with Prior Literature

The data convincingly implicate the ERES-ERGIC interface in autophagosome biogenesis. It would strengthen the manuscript to discuss previous studies reporting ERES and ERGIC remodeling and formation of ERERS-ERGIC contact sites (PMID: 34561617; PMID: 28754694) in the context of the current findings.

(2) Figure Labeling

The font size in Figure 1A and Supplementary Figure S1G is too small for comfortable reading. Please consider enlarging the labels to improve clarity.

(3) Experimental Conditions

In Figures 2A-C and Figure 4, it is unclear how the cells were treated. Were they starved in EBSS? Please include this information in the corresponding figure legends.

(4) LC3 Lipidation vs. Cleavage

In Figure 2A, ARFGAP1 knockdown appears to reduce LC3 lipidation without affecting Halo-LC3 cleavage. Clarifying this observation would help readers better understand the functional specificity of ARFGAP1 in the pathway.

(5) Use of HT-mGFP in Figure 2C

Please clarify why the assay in Figure 2C was performed in the presence of HT-mGFP. Explaining the rationale would aid interpretation of the results.

(6) FIB-SEM Imaging

For the FIB-SEM images in Figures 3 and S3, directly labeling the cellular structures in the images would greatly facilitate interpretation for the reader.

(7) Supplementary Figures

Many of the supplemental figures are high quality and contain key data. If space permits, I suggest moving these into the main figures. In particular, the FLASH-PAINT experiment could be presented as part of Figure 1.

(8) Text Revision for Clarity

In line 242, the phrase "but protein-protein interactions appear to be limited to RAB1" would benefit from clarification. A more precise formulation could be: "but stable protein-protein interactions appear to be limited to RAB1."

(9) COPII Inhibition Strategy

The authors used the dominant-active SAR1(H79G) mutant to inhibit COPII function. While this is effective in in vitro budding assays, the GDP-locked mutant SAR1(T39N) has been shown to be more effective in blocking COPII-mediated trafficking in cells. Including SAR1(T39N) in the analysis would provide stronger support for the conclusions.

Reviewer #3 (Public review):

The manuscript by Fuller et al describes a crosstalk between ARTG2A with components of the early secretory pathway, namely RAB1A and ARFGAP1. They show that ATG2A is recruited to membranes positive for RAB1A, which they also show to interact with ATG2A. In agreement with earlier findings by other groups, silencing RAB1A negatively affects autophagy. While ARFGAP1 was also found on ATG2A positive membranes, silencing ARFGAP1 had no impact autophagy. Notably, these ARFGAP1 positive membranes are not Golgi membranes.

The findings are interesting and the data are in general of good quality. I think the story is good enough to be published in eLife and I have the following questions, which the authors may attend to:

(1) Are the membranes to which ATG2A is recruited a form of ERGIC?

(2) Figure 3A/B: Is it possible to show a better example? The difference is barely detectable by eye. Since Immunoblotting is not really a quantitative method, I think that such a weak effect is prone to be wrong. Is there another tool/assay to validate this result?

(3) Is the curvature-sensitive region of ARFGAP1 required for its co-localization with ATG2A?

(4) What does Rab1A do? What is its effector? Or does the GTPase itself remodel the membrane?

(5) What about Arf1? It appears that this role of ARFGAP1 is unrelated to Arf1 and COPI? Thus, one would predict that Arf1 does not localize to these structures and does not affect ATG2A function

(6) Does ARFGAP1 promote fission of the membrane from its donor compartment?

(7) What are ARFGAP1 and Rab1A recruited to? What is the lipid composition, or protein that recruits these two players to regulate autophagy?

Comments on the latest version:

The revisions carried out by the authors are fine. The new data on ArfGAP1 and about the indirectness of the ATG2A and Rab1A interaction improve both clarity and strength of the manuscript. I have no further comments.

Author response:

The following is the authors’ response to the previous reviews

(1) Interpretation of LC3-II accumulation and phenocopying

Therefore, LC3-II accumulation alone is insufficient to support phenocopying in my view.

We agree with this assessment. Upon reconsideration, we concluded that LC3-II accumulation alone does not justify the use of the term "phenocopying." We have therefore removed this language from the manuscript.

(2) Strength of conclusions regarding autophagosome biogenesis

As presented, the findings support a correlative relationship rather than a defined role in autophagosome biogenesis.

We agree that our original wording overstated the strength of the conclusions. To better reflect the data, we revised the text to state that our findings "expound upon" rather than "elucidate" the role of these membranes in autophagosome biogenesis.

(3) Title wording

The title states that ATG2A ‘engages’ Rab1A- and ARFGAP1-positive membranes during autophagosome formation... A more descriptive term, such as ‘associates,’ would more accurately reflect the data.

We appreciate this suggestion. We revised the title to avoid implying a causal dependency. The title now states that ATG2A interacts with Rab1A- and ARFGAP1-positive membranes, emphasizing the membrane association observed in our study rather than a direct interaction with the proteins themselves.

(4) ARFGAP1 knockdown phenotype

The authors claim: ‘siRNA against ARFGAP1 had very little effect’ but the quantification and blots show actually no effect.

We agree that the original wording was imprecise. The sentence has been revised to state:

“siRNA against ARFGAP1 had no effect on flux.”

This more accurately reflects both the data and our original interpretation.

(5) Interpretation of ARFGAP knockdown experiments

Conclusions drawn from KD experiments in Fig. S2 should be interpreted with caution, as knockdown efficiency is very low, particularly for ARFGAP1/3 in the triple knockdown.

We agree with this caution and have revised the text accordingly. The manuscript now states:

“Knockdown of ARFGAP2, ARFGAP3 and ARFGAP1-3 marginally increased autophagic flux (Fig. S2B,C), suggesting either no role or a minor role as negative regulators of autophagy.”

This wording more appropriately reflects the limitations of the experiment.

(6) Discussion of ERGIC/ERES remodeling literature

It would strengthen the manuscript to discuss previous studies reporting ERES and ERGIC remodeling and formation of ERES-ERGIC contact sites (PMID: 34561617; PMID: 28754694).

We appreciate the suggestion. These studies were already cited and discussed in the original submission, and therefore no additional changes were required.

(7) Figure readability

The font size in Figure 1A and Supplementary Figure S1G is too small for comfortable reading.

We agree and have enlarged the labels in both figures to improve readability.

(8) Clarification of starvation conditions in figure legends

In Figures 2A-C and Figure 4, it is unclear how the cells were treated. Were they starved in EBSS?

We have updated the corresponding figure legends to explicitly state the starvation conditions used in these experiments.

(9) Interpretation of ARFGAP1 knockdown and LC3 lipidation

In Figure 2A, ARFGAP1 knockdown appears to reduce LC3 lipidation without affecting Halo-LC3 cleavage.

We do not observe a reproducible reduction in LC3 lipidation following ARFGAP1 knockdown and therefore do not believe this conclusion is supported by the data. No changes were made in response to this comment.

(10) Clarification of protein-protein interaction statement

The phrase ‘but protein-protein interactions appear to be limited to RAB1’ would benefit from clarification.

We agree and have adopted the suggested wording. The manuscript now states:

“but stable protein-protein interactions appear to be limited to RAB1.”

We thank the reviewers again for their constructive feedback and for helping us improve the clarity and accuracy of the manuscript.

Version 3

Abstract

Autophagosomes form from seed membranes that expand through bulk-lipid transport via the bridge-like lipid transporter ATG2. The origins of the seed membranes and their relationship to the lipid transport machinery are poorly understood. Using proximity labeling and a variety of fluorescence microscopy techniques, we show that ATG2A localizes to extra-Golgi ARFGAP1 puncta during autophagosome biogenesis. ARFGAP1 itself is dispensable during macroautophagy, but among other proteins associating to these membranes, we find that RAB1 is essential. ATG2A co-immunoprecipitates strongly, albeit indirectly, with RAB1A, and siRNA-mediated depletion of RAB1A/B blocks autophagy downstream of LC3B lipidation, similar to ATG2A depletion. Further, when either autophagosome formation or the early secretory pathway is perturbed, ARFGAP1 and RAB1A accumulate at ectopic locations with autophagic machinery. Our results indicate that ATG2A engages a RAB1A complex on select early secretory membranes in support of autophagosome biogenesis.

Introduction

Macroautophagy is a cellular degradative pathway in which a seed vesicle grows to encapsulate a target cargo, closes to form the complete autophagosome, and subsequently fuses with the lysosome. This transient organelle forms de novo in approximately ten minutes and requires the delivery of millions of lipids1. ATG2 proteins function as bridge-like bulk lipid transport proteins2–5 and are essential during the membrane expansion step of autophagosome biogenesis. We propose that ATG2 bridges the ER and the developing phagophore, thereby allowing lipids to cross the cytosol through a long hydrophobic groove3. At the phagophore, lipids are equilibrated between both leaflets of the bilayer via the ATG9 scramblases6–8. As such, ATG9 is essential to macroautophagy and is a key component of the seed membranes that ultimately expand during ATG2-mediated lipid transport9, 10.

In addition to ATG9 vesicles, several other membrane sources have been proposed to be involved in autophagosome biogenesis. In particular, both genetics and biochemistry implicate an essential role for the early secretory pathway. In yeast, some COP-II vesicle and ER exit site (ERES) proteins contribute to starvation-mediated autophagy at a stage that is downstream of formation of the pre-autophagosome structure (PAS)11, 12. Furthermore, yeast ERES naturally colocalize with the autophagic cargo Ape113, and that colocalization increases when autophagosome growth is stalled by depletion of essential factors13, 14. Autophagy proteins themselves collect at ERES; both imaging and cross-linking studies indicate a close association between the two membranes and disruption of ERES activity through depletion of Sec12 dramatically reduces Atg8 puncta formation15. Finally, the direct integration of yeast COPII-derived membranes into developing phagophores is suggested by the detection of the COP-II cargo Axl2-GFP, a transmembrane protein, at both phagophores and mature autophagosomes16.

In human cell lines, genetic or pharmacological disruption of the COPII-related small GTPase SAR1A reduces the frequency of autophagic events17. Following biochemical fractionation of cellular membranes, the ER-to-Golgi Intermediate Compartment (ERGIC) is especially competent for supporting in vitro lipidation by LC3B, the primary marker of autophagic structures17, 18 and clusters of COPII positive membranes can be seen accumulating adjacent to phagophores19. Like in yeast, these studies can be interpreted as a key role for ERES or ERGIC membranes in forming or expanding autophagosome membranes. As starvation-mediated macroautophagy drives substantial remodeling of both the ERES and ERGIC membranes18, 20, including the formation of subdomains with unique protein-dependent organelle-organelle contact sites21, it is possible that the ERES/ERGIC contribution to autophagy represents a distinct subcompartment of these membrane structures.

Presumably, any contribution of an early secretory membrane must ultimately coordinate with recruitment of ATG9 vesicles, in order to ensure that the lipid transport machinery is competent to expand the phagophore membrane. Several studies have postulated an early fusion intermediate in phagophore formation, such as the description of the HyPas compartment where vesicles bearing distinct autophagy proteins (ATG16L1 and FIP200) were observed to coalesce early in autophagosome biogenesis22. Intriguingly, when the lipid transporter ATG2 is genetically depleted from cells, all of the essential autophagy machinery including ATG9 concentrates into large compartments9, 23–25. Correlative light and electron microscopy of these structures indicate they are large assemblies of ATG9 vesicles surrounded by more complex membranes which harbor FIP200 and WIPI2, among other proteins9, 26. These complex membranes often adopt a cup-like shape, leading us to name them “proto-phagophores”9. The origins of the proto-phagophores have been unknown.

In this study, we demonstrate that ATG2A is enriched during autophagosome biogenesis at RAB1/ARFGAP1-positive membranes. Although both proteins can be found within the ERGIC compartment, proximity data and colocalization analyses in secretory pathway disrupted cells suggest that RAB1 and ARFGAP1 membranes involved in ATG2 recruitment are distinct from generic ERGIC membranes. Genetic depletion experiments indicate that RAB1 and ATG2 function at a similar step in autophagosome formation and further these proteins can be isolated together in a complex. Finally, both ARFGAP1 and RAB1 co-localize with proto-phagophores, implying these membranes can recruit much of the key early autophagy machinery and may be dynamically remodeled into cup-like morphologies.

Results

ATG2A is proximal to early secretory membranes

To determine the local proteome adjacent to ATG2A, we used a proximity labeling approach. We fused the biotinylation enzyme APEX2 and a green fluorescent marker to the amino-terminus of ATG2A (APEX2-EGFP-ATG2A) and expressed this construct in ATG2A/B double knockout cells (ATG2 DKO cells). In this way, the APEX2 enzyme will biotinylate cellular proteins in close proximity to the expressed ATG2A fusion protein27. To ensure that the APEX2-EGFP-ATG2A construct is active and correctly localized, we measured the restoration of macroautophagy in these cells by following LC3B flux and p62 clearance (Fig. S1A-C). While both LC3B-II and p62 dramatically accumulate in the ATG2 DKO cells, this accumulation is almost entirely reversed upon re-expression of APEX2-EGFP-ATG2A.

We also validated that biotinylation in these cells depends upon the presence of APEX-EGFP-ATG2A and a full reaction. In cells expressing APEX2-EGFP-ATG2A, biotin is only observed when both the biotin substrate (biotin phenol) and the reaction promoter (hydrogen peroxide) are included (Fig. S1D,E). Likewise, streptavidin-mediated immunoprecipitation revealed a strong enrichment of biotinylated material only when the full reaction conditions were employed (Fig. S1F). Furthermore, the anti-biotin and GFP signals colocalized establishing that biotinylation is favored at sites proximal to ATG2A (Fig. S1D).

Mass spectrometry of proteins isolated by streptavidin-mediated immunoprecipitation was performed using a three state quantitative SILAC experiment27 with light, medium, and heavy isotopes as follows: no hydrogen peroxide as a negative control, a full reaction in complete media, and a full reaction in starvation media and 100 nM bafilomycin. When comparing both the heavy and medium isotope states to the light state, mass spectrometry analysis revealed an abundance of autophagy related proteins in both full media and under starvation (Figs. 1A, S1G, respectively). The majority of these proteins were cargo adaptors such as NBR1, p62 (SQSTM1), and TAX1BP1, suggesting that the target cargo is immediately adjacent to the biogenesis machinery during autophagosome maturation. We also observed a clear enrichment of ERES or COPII proteins including SEC16A, SEC23A, SEC23IP, and SEC24B/C. Additional ERES proteins such as TFG and SEC31A were also enriched in some, but not all, of the three replicates. In addition, ARFGAP1, a COPI effector that is primarily localized to the Golgi but also found in earlier secretory membranes, was one of the two most enriched proteins in both full media and starvation. Both ARFGAP1 and the COPII proteins suggest ATG2 is localizing to the early secretory pathway.

ATG2A is proximal to ARFGAP1-positive membranes.

(A) Proximity labeling of HEK293 ATG2 DKO cells stably expressing APEX2-GFP-ATG2A was performed as previously described27. Three biological replicates were averaged and plotted as the log2 ratio between the full labeling reaction in complete media over cells that were not treated with hydrogen peroxide on the x axis, and -log10 of the p-value on the y axis. The red line demonstrates the p-value cutoff of 0.05. The proteomics revealed an enrichment of autophagy (labeled in green) and early secretory proteins (COPII/ER Exit Site proteins are labeled in orange, ARFGAP1 is labeled in red). (B) Strategy to identify phagophores by live cell imaging. Phagophores are defined as sites where both GFP-ATG2A and TagBFP-LC3B co-localize. (C-D) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2A and TagBFP-LC3B and transfected with mRuby-SEC23A or RFP-SEC23IP/p125a or ARFGAP1-TagRFP. The cells were starved for 2-4 hours prior to imaging. The fraction of phagophores (white arrows) that colocalized with the third fluorescent protein (transfected gene) was quantified and plotted in (D). The gray lines show the cell periphery and the insets focus on representative phagophores. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled for each condition. Maximum intensity projections of confocal images. (E) Global organelle profiling data from https://organelles.sf.czbiohub.org32. Whole organelle profiles were established by pulling down endogenously tagged proteins and performing quantitative proteomics. The enrichment value for each protein was calculated by taking the difference in median log2 LFQ values between the replicates and the null distribution.

ATG2A is recruited to ARFGAP1 membranes during autophagosome biogenesis

As an orthogonal method to assess the proximity of ATG2A to early secretory components, we performed FLASH-PAINT, a super resolution technique with high multiplexing capabilities28. FLASH-PAINT offers sub-organelle resolution while simultaneously allowing us to screen >10 proteins. ATG2 DKO cells stably expressing GFP-ATG2A were starved and then immunostained for autophagic proteins (GFP-ATG2A, p62, and LC3B), ERES proteins (Tango1, SEC24C, SEC31A, SEC16A, TFG), and markers of COPI-positive membranes (ARFGAP1 and COPI). All proteins were concurrently labeled with primary antibodies that were in turn identified by their association with secondary antibodies with unique DNA barcodes (Fig. S1I). A fluorophore conjugated imager DNA strand and adaptor strands were added to target the imager to the correct secondary antibody. Eraser strands were added in between rounds of imaging to outcompete the adaptor and allow for sequential labeling. In this way, the sub-cellular localization of all ten proteins could be established within the same image.

In these images, the Golgi can easily be identified as a perinuclear ribbon staining strongly for both ARFGAP1 and COP1 (magenta and purple respectively, Fig. S1J). In addition, there are bright puncta throughout the cell (pink arrow heads) stained with TFG and SEC31A, indicating that they are ERES. ARFGAP1 and COPI frequently also accumulated at ERES (made more obvious in the top-right inset, where TFG and SEC31 are removed), consistent with previous reports that COPI vesicles facilitate ER to Golgi transport in human cells29–31. GFP-ATG2A is mostly cytosolic, however small accumulations of ATG2 could be found at ERES and were generally observed to be proximal to ARFGAP1/COP1 (inset; brackets). Note that clusters of LC3B and p62 are also present. Occasionally, non-ERES accumulations of p62 could be observed in close proximity to LC3B and ATG2A (lower insets), and here again ATG2 is proximal to ARFGAP1.

Critically, we can directly measure the nominal proximity of each protein relative to the others under investigation by quantifying the median distance between all ten tested proteins within 200 nm of each individual protein (Fig. S1K). All early secretory pathway proteins closely associate with each other, including the COPI proteins ARFGAP1 and β-COP. As the majority of ATG2A was cytosolic, its average distances from other proteins were greater, however, within this cohort of candidate proteins, we found that ATG2A is most closely associated with p62 and ARFGAP1 (15nm and 16nm respectively; top row of Fig. S1K), exactly consistent with the APEX screen.

The APEX and FLASH-PAINT datasets provided a clear but static result that ATG2A is recruited to ARFGAP1 positive structures. To determine whether this recruitment was coincident with phagophore formation, we stably expressed GFP-ATG2A and TagBFP-LC3B in ATG2 DKO cells in order to identify phagophores. ATG2A primarily localizes to phagophores and lipid droplets, whereas LC3B primarily localizes to phagophores, autophagosomes and autolysosomes. Where these two proteins colocalize suggests a growing phagophore in the process of membrane expansion (Figs. 1B). To assess co-localization with phagophores, we expressed candidate early secretory proteins labeled with red fluorophores (Fig. 1C,D; see Fig. S1H for analysis strategy). ARFGAP1-TagRFP colocalized with ∼90% of actively forming phagophores, whereas only ∼30% of phagophores were coincident with the ERES markers mRuby-SEC23A or RFP-p125a/SEC23IP. Further, while the ARFGAP1 signal was largely concentrated in the Golgi, phagophores most frequently colocalized with extra-Golgi ARFGAP1 puncta.

Finally, we compare our results to the subcellular localization of ARFGAP1 established by global organelle profiling available at the database at https://organelles.sf.czbiohub.org32. ARFGAP1 is enriched on autophagic membranes positive for LC3B as well as endolysosomal and early secretory membranes (Fig. 1E). Strikingly, ARFGAP1 was most highly enriched on membranes isolated via RAB1A immunoprecipitation. The autophagic proteins ATG2A, ATG2B, WDR45, and ATG9A, all of which are involved in membrane expansion of the autophagosome, were also highly enriched on RAB1A membranes. These collective results establish that both endogenous ARFGAP1 (organelle isolation, APEX, FLASH PAINT) and overexpressed ARFGAP1 (live imaging) naturally associate with ATG2A, including during phagophore biogenesis.

Autophagic flux depends upon RAB1 but not ARFGAP1

To look for a potential functional role for ARFGAP1 in autophagy, we measured autophagic flux using the HaloTag dropout assay in which liberation of a free HaloTag (HT) is indicative of lysosome-mediated proteolysis of the full-length (FL) protein33. We stably expressed HaloTag7-mGFP-LC3B or HaloTag7-mGFP in WT HEK293 cells and then used siRNA to knockdown various regulators or potential regulators of macroautophagy initiation (Fig. 2A,B, S2A). Despite efficient knockdown, siRNA against ARFGAP1 had no effect on flux; the ratio of free HaloTag to full-length HaloTag7-mGFP-LC3B following ARFGAP1 knockdown was similar to both nontransfected (NT) and nontargeting siRNA controls. In contrast, knockdown of the autophagy initiation factor FIP200 dramatically suppressed dropout of the HaloTag. We also looked for an interaction between ARFGAP1 and ATG2A, but even when both proteins were overexpressed, we could not detect significant interaction via co-immunoprecipitation (co-IP; Fig. 3A,B). Thus, it is likely that ARFGAP1 localizes to a key membrane involved in autophagosome biogenesis but is not itself a player in initiating autophagosome formation. To examine whether the other isoforms of ARFGAP1 could redundantly contribute to autophagosome biogenesis, we knocked ARFGAP1, ARFGAP2, and ARFGAP3 down individually and collectively. Knockdown of ARFGAP2, ARFGAP3 and ARFGAP1-3 marginally increased autophagic flux (Fig. S2B,C), suggesting either no role or a minor role as negative regulators of autophagy.

Autophagic flux depends upon RAB1 but not ARFGAP1.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that siRNA KD of RAB1 inhibits autophagic flux in cells starved in EBSS for four hours. In gel fluorescence reveals HaloTag7s moiety bound to the TMR ligand. Indicated bands are full length (FL) HT-mGFP-LC3B protein and HaloTag7 after cleavage (HT) in the lysosome. The ratio between FL and HT demonstrates autophagic flux, quantified in (B). Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were collected for each condition. (C-D) Immunoblot of lysates from three biological replicates demonstrating that siRNA KD of RAB1 results in an accumulation of LC3B-II in cells starved in EBSS for four hours. The band intensity of LC3B-II was normalized against the intensity of GAPDH for each lane. To normalize each replicate, each individual value was divided by value of the non-transfected sample. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. (E-G) FACS data demonstrating that RAB1A/B KD results in a block in autophagic flux. H4 cells with a heterozygous GFP-mCherry-LC3B endogenous tag were transfected with siRNA for three days, following which they were starved in HBSS for four hours and analyzed by FACS. The ratio between GFP and mCherry is an indicator of autophagic flux, as the GFP signal is quenched in the lysosome. This ratio is displayed in histograms in (E) for one replicate and as averages in (F). The KDs of each protein can be seen in (G).

ATG2A interacts indirectly with RAB1A.

(A-B) Immunoblot showing CoIP between GFP-RAB1A and 3xFLAG-ATG2A compared to the same reaction with other GFP-tagged early secretory proteins. The cells were incubated with either complete media or EBSS for four hours prior to harvesting. For the CoIP, ∼3 mg of cell lysate was used per condition (slight deviations in total amount between replicates, but not between lanes). To quantify the results, the intensity of the 3xFLAG-ATG2A IP signal was ratioed against the input signal. To normalize each replicate, each individual value was divided by the average of the replicate. Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. The media was not a significant source of variation (p = 0.3535). (C-D) A fully reconstituted CoIP reaction between GFP tagged bait proteins and 3xFLAG-ATG2A reveals no direct interaction between RAB1A and ATG2A. GST tagged GFP-GABARAPL1 (GL1) and GFP-RAB1A were purified from E. Coli, during which process the GST tag was proteolytically removed. The purified proteins were then used in a CoIP reaction using GFP trap beads as in (A), but with no cell lysate present in the reaction. BSA was used to block the beads and as a negative control. To quantify the results, the intensity of the 3xFLAG-ATG2A IP signal was ratioed against the input signal. To normalize each replicate, each individual value was divided by the average of the replicate. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001.

As shown in (Fig. 1E), both the ATG2/ATG9 lipid transport apparatus as well as ARFGAP1 are associated to RAB1A-positive membranes. RAB1 proteins mediate ER to Golgi trafficking in the early secretory pathway. The yeast RAB1 homologue, Ypt1, is essential to autophagy34, 35, while in mammals, genetic depletion of either RAB1A or RAB1B modestly reduces autophagic flux36, 37. Thus, we tested the functional and physical interaction of RAB1 proteins with ATG2. Consistent with previous results, siRNA knockdown of either RAB1A or RAB1B in our system partially inhibited flux in the Halo dropout assay (Fig. 2A,B). Strikingly, knockdown of both isoforms simultaneously was nearly as effective as knocking down FIP200, establishing that RAB1 proteins play an essential role in supporting starvation induced autophagic flux. This result was consistent with but more pronounced than previous work examining LC3B flux following RAB1A/B double knock down38. As an orthogonal method to examine the role of RAB1 in LC3B flux, we used H4 neuroglioma cells with a GFP-mCherry-LC3B heterozygous knock in reporter39.

Following knock down and starvation, we examined the ratio between GFP and mCherry by FACS. The low pH of the lysosome quenches the fluorescence of GFP at a higher rate than mCherry, resulting in a low ratio under starvation conditions as seen in the HBSS and siCNTL conditions (Fig. 2 E-G). In contrast, knock down of FIP200 or RAB1A/B resulted in an increased ratio, confirming that these proteins are necessary for proper autophagic flux. Of note, the RAB1A/B knockdown did not block flux as efficiently as the FIP200 KD. If RAB1 is essential for autophagy, this result suggests that either very few copies of RAB1 are needed to support autophagy or that another gene could be compensating for the lack of the two RAB1 isoforms.

In addition to following tagged LC3B, we also immunoblotted for endogenous LC3B in each of our siRNA experiments. We found that knockdown of either RAB1 isoform, and especially knockdown of both together leads to a clear accumulation of the membrane conjugated form of LC3B, LC3B-II (Fig. 2A,G). As LC3B overexpression can alter endogenous LC3B patterns, we confirmed this increase in a cell line expressing only HaloTag7-mGFP (Fig. 2C,D). Knockdown of RAB1A and RAB1B drove a threefold increase of LC3B-II (Fig. 2D). We also confirmed this phenotype in WT HEK293 cells in a classical LC3B flux assay with bafilomycin (Fig. S2D), where we found elevated LC3B-II levels in all conditions.

Increasing LC3B-II levels without autophagic flux is indicative of a block to macroautophagy at a late stage, downstream of initiation, and is therefore easily distinguished from the block induced by depleting FIP200 (where LC3B-II levels go down). Intriguingly, LC3B-II levels also increase when ATG2 proteins are genetically depleted (Fig. S1A,B), suggesting ATG2A and RAB1A might function concurrently. Indeed, overexpressed GFP-RAB1A could strongly immunoprecipitate flag-tagged ATG2A in our cells, indicating these proteins are in a shared complex (Fig. 2E,F). However, ATG2A and RAB1A (bound either to GDP or GTPγS) do not interact in a fully reconstituted system (Fig. 3C,D), suggesting that the interaction is indirect. In contrast, ATG2A and GABARAPL1 cleanly interact in the in vitro CoIP.

We also tested several other proximity hits from the biotinylation analysis for interaction with flag-tagged ATG2A, including GFP-tagged copies of ERES resident proteins SEC23A, SEC31A, SEC16A, Sar1 and all four homologs of Sec24 (Fig. S3 A-D). We found that ATG2A co-immunoprecipitated weakly with many of them, the strongest of which was SEC24D. Note that this SEC24D interaction was barely above the background of pull-down with GFP alone (Fig. 3A,B). Thus, consistent with proximity and imaging analyses, ATG2 proteins are in the general proximity of the ERES machinery, but stable protein-protein interactions appear to be limited to RAB1.

Disruption of the early secretory pathway results in the recruitment of autophagic proteins to RAB1A-positive tubular membranes

These results suggest that ATG2 engages an early secretory membrane, distinct from the ERES, in a RAB1-dependent manner. We next tested whether disruption of the early secretory system would impact the localization of ATG2A. To this end, we overexpressed SAR1BH79G, the catalytic dead mutant of the small GTPase SAR1B that recruits the COPII coat. SAR1BH79G binds the membrane and recruits the COPII coat, but is unable to release, thereby trapping COPII onto the membrane. Overexpression of this construct has been shown to reduce autophagy in starved cells17, 36. Consistent with those results, overexpression of SAR1BH79G-mOrange2 diminished autophagic flux of both a bulk cargo (HaloTag7-mGFP reduced ∼20%) and of LC3B (HaloTag7-mGFP-LC3B reduced by ∼30%) by the HaloTag degradation assay33 (Figs. 4A,B). In a parallel assay following autophagic flux instead by the changing fluorescence of tandem tagged GFP-mCherry-LC3B, we detect a similar, but statistically insignificant trend. As both assays are likely sensitive to SAR1BH79G expression levels, which vary in our transient transfection system, we also tested how LC3 puncta accumulation correlates with SAR1BH79G expression on a per cell basis. LC3B puncta number were lowest in cells expressing high amounts of SAR1BH79G (Figs. S4A,B), consistent with a block to autophagosome formation when the secretory pathway is disrupted rather than a later impact on autophagosome clearance.

ATG2A, LC3B, ARFGAP1, and RAB1A localize to tubules following disruption of the early secretory pathway.

(A-B) In gel fluorescence image (top panel) and immunoblots demonstrating that SAR1BH79G-mOrange2 (mO) overexpression negatively affects autophagic flux in cells starved in EBSS for four hours. The ratio between the FL and the free HT represents the autophagic flux in each condition (as in Fig. 2 A/B) and is plotted in (B). Statistical significance was assessed by two way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from four biological replicates were collected for each of the HT-mGFP conditions and five replicates were collected for each of the HT-mGFP-LC3B conditions. (C) Live cell imaging of ATG2 DKO cells stably expressing GFP-ATG2 that were transfected with SAR1BH79G-mOrange2 and starved for two hours. Note that GFP-ATG2A forms long tubules that do not colocalize with the SAR1BH79G-mOrange2 signal. Maximum intensity projection of a confocal image. Gray line was added to show cell outline. (D-E) Live cell imaging demonstrates the localization of ARFGAP1-TagRFP (D) and SNAP-RAB1A (E) on GFP-ATG2A/TagBFP-LC3B positive tubules following SAR1BH79G overexpression. Arrows denote thin GFP-ATG2A positive tubules whereas arrow heads denote thicker tubules that lack most GFP-ATG2A but contain high levels of LC3B, ARFGAP1, and RAB1A. Maximum intensity projections of confocal images. (F) CLEM-FIBSEM of cells prepared as in (C). GFP-ATG2A fluorescence corresponds to a thin tubular membrane that is connected to a thicker tubule. The small inset in the left image shows a single SEM slice. The arrowheads are pointing to the tubule which is segmented and displayed in turquoise in the larger image, whereas the arrows are pointing at adjacent ER tubules which are segmented and displayed in the larger image in yellow. At many points, these ER structures come into very close apposition with the tubule which is denoted in red. The ER structures are removed in the right image for better visualization of the ATG2A positive tubule.

Unexpectedly, in a small subset of transfected cells (1.5±0.8%; calculated from the images in Fig. S4A,B), overexpression of SAR1BH79G under starvation conditions shifted the localization of GFP-ATG2A from a largely punctate distribution to long, tubular objects that were SAR1BH79G negative (Fig. 4C). GFP-ATG2A decorated these tubules with varying intensities approximating a beads-on-a-string pattern (e.g. arrows in Fig. 4D,E). Fluorescently tagged forms of LC3B, ARFGAP1 and RAB1A also decorated these tubules. However, these three proteins displayed the brightest intensities where ATG2A was dimmer (e.g. arrowheads in Fig. 4D,E). These three proteins, but not ATG2A, also decorated larger structures that appeared to be contiguous with the tubule, including bulbous regions at the tubule termini.

To determine the ultrastructure of these regions, we subjected GFP-ATG2A-expressing cells to correlative light and focused ion beam scanning electron microscopy (CLEM-FIB-SEM). The GFP-ATG2A fluorescence signal corresponded to thin, tubular membranes (Fig. 4F and S4D), which is consistent with the finding that ATG2A preferentially binds highly curved membranes40. The tubule is surrounded by ER elements, many of which appear to be in very close contact (Fig. 4F, S4C). GFP-ATG2A signal was more intense in regions with closely adjacent ER, consistent with a possible contact site localization. The tubule continued into larger structures at either end, which were GFP-ATG2-negative, and morphologically similar with the bulbous structures harboring ARFGAP1, RAB1A and LC3B in fluorescence images.

Thus, although unusual, these rare tubules afforded an opportunity to demonstrate that ATG2, ARFGAP1 and RAB1A target the same membrane, easily resolvable even at confocal resolution. ATG2 collects where ER is closely adjacent, and LC3B accumulates with RAB1A, implying a topological organization with LC3B and RAB1A decorating the same structure, suggesting that RAB1A membranes may be a component or progenitor of the phagophore membrane.

ARFGAP1 and RAB1A accumulate around pre-autophagic structures in ATG2 DKO cells

We previously published that very small, ATG9- and LC3B-positive vesicles accumulate in large clusters with the cargo-adaptor p62 in ATG2 DKO cells9. This cluster is surrounded by a shell of more complicated membranes which we termed “protophagophores” because they colocalize with other autophagy proteins such as WIPI2 and because many of them appear to adopt a cup-like shape in TEM images. Beyond this protophagophore shell, the entire structure is surrounded by ER (illustrated in Fig. 5A). Our work and others have shown that essentially all proteins involved in autophagosome biogenesis accumulate with these clusters, including FIP200, ATG14, ATG9A, ATG16L, and p629, 23, 24, suggesting they are repositories of autophagosome biogenesis material awaiting the arrival of ATG2 proteins. If early secretory membranes contribute directly in any way to the biogenesis event, they may also concentrate at these sites.

ATG2 deletion results in the accumulation of ARFGAP1 and RAB1A on autophagic membranes.

(A) Schematic of the membranes that accumulate at the ATG2 DKO compartment based on the TEM images in Olivas et al9. The core of the compartment contains small vesicles that stain for ATG9A, LC3B, and p62. The periphery of the compartment is comprised of more complicated cup shaped membranes that stain for WIPI2. The entire compartment is enwrapped by ER tubules. (B) Immunofluorescence images demonstrating the distribution of early secretory membranes around the ATG2 DKO compartment which is marked by either p62 (mouse antibody) or LC3B (rabbit antibody). The images presented are single confocal slices of the center of the ATG2 DKO compartment. The gray lines show the cell periphery and the insets focus on the largest accumulation of p62 or LC3B in the cell. (C) Quantification of the images in (B) and in (Fig. S4A). The enrichment factor is equal to the mean of the given protein at the periphery of the ATG2 DKO compartment compared to its mean throughout the cell. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. Data from three biological replicates were pooled together for each of the conditions. (D-E) Immunofluorescence images demonstrating the enrichment of overexpressed GFP-RAB1A (D) or SNAP-RAB1A (E) at the periphery of the ATG2 DKO compartment as determined by the presence of p62 (D) or WIPI2 (E). Presented as a single confocal slice. The gray lines show the periphery of the cell and the insets focus on ATG2 DKO compartments that are distal from the Golgi.

We therefore performed IF against various markers of the early secretory pathway in ATG2 DKO cells to look for colocalization with this compartment (represented by staining for p62 (mouse antibody) or for LC3B (rabbit antibody); Figs. 5B-D). We found that ER exit site markers such as SEC16A, SEC24C, SEC31A, and TFG were not enriched around the compartment, neither was the Golgi protein GM130 (Figs. 5B-C and S5A). ßCOP was modestly enriched, but at levels that by eye, did not suggest encapsulation of the compartment the way autophagy proteins collect. In addition, overexpressed mEmerald-ERGIC53 did not localize to the ATG2 DKO compartment (Fig. S5B). In contrast, ARFGAP1 targeted all of the ATG2 DKO compartments, forming a robust spherical shell at the periphery of the p62 signal in HEK293 ATG2 DKO cells (Fig. 4B-D), or permeating the entire compartment in HeLa ATG2 DKO cells41 (Fig. S5C).

We were unable to find a strong RAB1A signal by immunofluorescence, however, when overexpressed, tagged versions of RAB1A also formed a sphere around the HEK293 ATG2 DKO compartment (Fig. 5D. Both ARFGAP1 and SNAP-RAB1A colocalized with WIPI2 (Fig. 5E), consistent with a model whereby these membranes arrive at the PAS prior to ATG2A recruitment and subsequent autophagosome biogenesis.

ARFGAP1 accumulates at the ATG2 DKO compartment independently of the Golgi

The vast majority of ARFGAP1 is found at the Golgi, but ATG2A specifically localizes to extra-Golgi ARFGAP1 puncta (Fig. 1C, S1J). Consistent with a non-Golgi origin for these membranes, both natural and induced perturbations of Golgi structure have no impact on the recruitment of ARFGAP1 membranes to the p62-rich compartments in ATG2 DKO cells. First, in the course of these studies we observed that ARFGAP1 persistently colocalized with p62 in cells undergoing mitosis (Fig. 6A,B,C), even as the Golgi fragments and dissipates. Second, when cells were treated with brefeldin A to inhibit the formation of the Golgi42, ARFGAP1 again persistently colocalized with p62 (Fig. 6B,C). As brefeldin A specifically targets GBF1, the GEF for the ARF proteins43, this result also suggests that ARFGAP1 localization to the p62 compartment is independent of GTP bound ARF proteins. Third, overexpression of GFP-SidM/DrrA, an enzyme from Legionella Pneumophila which acts as both a GEF44, 45 and a GDF46 for RAB1, fragments the Golgi44 and sequesters RAB1 to the plasma membrane45. We found that following GFP-SidM expression in our cells, ARFGAP1 maintained colocalization with p62, even as the Golgi-associated ARFGAP1 as well as the Golgi-marker GM130 were completely dissipated (Fig. 6D, S6A). Thus, the ARFGAP1 localizing to the p62 compartment is not influenced by dispersal of the Golgi and does not colocalize with other Golgi markers, implying a non-Golgi origin for these structures.

ARFGAP1 accumulation at the ATG2 DKO compartment is independent of the Golgi.

(A) Immunofluorescence microscopy demonstrates that ARFGAP1 colocalizes with p62 in ATG2 DKO cells. Inset 1 highlights a cell undergoing mitosis, in which the colocalization of these two markers persists as denoted by the arrows. Inset 2 shows several cells, a ring-like accumulation of ARFGAP1 around p62 is clearly visible, separate from the Golgi. Maximum intensity projection of a confocal image. (B-C) Immunofluorescence images showing that perturbations which disrupt Golgi formation don’t affect the recruitment of ARFGAP1 to the ATG2 DKO compartment. HEK293 ATG2 DKO cells were treated or not treated with 5 μg/ml Brefeldin A for one hour, following which they were immunostained against ARFGAP1 and p62. The gray lines show the periphery of the cell and the insets focus on ATG2 DKO compartments as seen by the p62 accumulations. Colocalization between ARFGAP1 and p62 was assessed in (C) using the Pearson’s Correlation Coefficient. In the untreated sample, cells were separated into two groups: mitotic and non-mitotic, based on the ARFGAP1 fluorescence pattern. Statistical significance was assessed by one way ANOVA. *, adjusted P value <0.05. **, adjusted P value <0.01. ***, adjusted P value <0.001. ****, adjusted P value <0.0001. (D) Immunofluorescence image demonstrating that Golgi disruption by SidM/DrrA overexpression did not perturb the colocalization between ARFGAP1 and p62. HeLa ATG2 DKO cells were transfected with GFP-SidM and then immunostained against ARFGAP1 and p62. The gray line shows the periphery of the cell and the inset focuses on the largest accumulation of p62.

Finally, we tested whether the GAP domain of ARFGAP1 or the ALPS motif are necessary for its localization to the PAS, by comparing the colocalization of various tagged constructs with GFP-ATG2A and TagBFP-LC3B (Fig. S6A). Disruption of the GAP domain via the well characterized R50K mutation had no impact on its localization to the ATG2/LC3B-positive phagophores (Fig. S6B). Mutation of the ALPS domain (L207D, V279D) altered ARFGAP1 distributions in the cell such that specific localizations could no longer be quantified (the redistribution to the cytoplasm created a high background that was difficult to properly threshold). Nonetheless, even in this condition there remained clear accumulations of this construct at phagophores indicating that recruitment is not strongly dependent upon the ALPS motif. As these elements of ARFGAP1 are essential to its Golgi function, these results further suggest that the ARFGAP1/RAB1A membranes involved in autophagosome biogenesis derive from early secretory membranes prior to the Golgi.

Discussion

Here we describe an interaction of ARFGAP1/RAB1-positive membranes with the autophagic biogenesis machinery including the lipid transporter ATG2. These membranes likely derive from a privileged region of the ERGIC consistent with the literature tying these intermediate secretory membranes to autophagy17, 21. In particular, the presence of both ARFGAP1 and RAB1 plus the general proximity to ERES proteins across a variety of assays is consistent with an early secretory origin for these membranes. Intriguingly, ARFGAP1 and RAB1 coalesce around the periphery of the ATG2 DKO compartment exactly where we previously described complex cup-shaped membranes, so called proto-phagophores9, accumulate. We do not detect any other ERES or ERGIC markers at these sites, implying proto-phagophores derive from only a subdomain of the ERGIC which we suspect is remodeled specifically for the purpose of supporting autophagy.

How do the ERGIC-derived RAB1 membranes contribute to autophagosome formation? Ypt1, the yeast homologue of RAB1A, was identified early on as a genetic regulator of autophagy34, and subsequently a number of studies have connected both Ypt1 and RAB1 to autophagosome biogenesis, but the precise mechanism of action remains contentious. In yeast, Ypt1 has been shown to interact with Atg135 and Atg1747, highly suggestive of a role in regulating or recruiting the kinase machinery driving initiation of autophagosome biogenesis. Human RAB1A binds the VPS34 complex favoring conformations that likely support PI3P formation48. Thus, either or both of the early acting complexes in autophagosome biogenesis may engage RAB1 proteins. In this study, we find that RAB1 depletion elevates LC3B-II levels, suggesting that RAB1 acts either downstream or independently from the lipidation machinery. Further, we present two orthogonal and highly quantitative methods to follow LC3B degradation in the lysosome. Both methods unambiguously show a clear, albeit partial, genetic dependence of LC3B degradation upon RAB1 consistent with RAB1 promoting autophagy. As we detect RAB1A at the ATG2 DKO compartment, RAB1A recruitment appears to precede ATG2A recruitment to the PAS, consistent with an interaction with either the ULK1 or VPS34 complexes, but is not essential to support LC3B-II formation on early autophagy membranes.

We propose that ARFGAP1/RAB1 membranes are directly integrated into the phagophore prior to ATG2 recruitment. The fusion of vesicular sources has been proposed as a general mechanism for phagophore biogenesis49, 50. For example, the HyPas model postulates that FIP200 vesicles from the cis-Golgi fuse with endosomally derived ATG16L1 vesicles to create the nucleating structure for autophagosome formation22. Recent reports, including our own, have also identified ATG9 vesicles as a seed for autophagosome biogenesis, which expand through lipids delivered by ATG29, 10 and then scrambled into both leaflets by ATG96–8. An integration of these models is necessary. Geometric estimates of the surface-to-volume ratio in yeast autophagosomes, gleaned from cryo-electron microscopy tomograms, suggest that as much as 20% of the membrane might derive from fusion rather than lipid transport in order to deliver luminal volume51, which cannot be supplied by lipid transport. We therefore propose a model in which an early fusion step between ATG9A vesicles and ARFGAP1/RAB1 membranes, both of which accumulate in ATG2 DKO cells, primes the nucleating membrane for autophagosome biogenesis.

Materials and Methods

Plasmids and reagents

The following plasmids were used in this study: pLVX Puro GFP-ATG2A (Addgene 250863; Previously published3), pLVX Puro APEX2-GFP-ATG2A (Addgene 250864; This study; a PCR fragment containing flanking EcoRI cutsites around APEX2 was generated from pcDNA3 APEX2-NES (Addgene #49386) and was inserted into GFP-ATG2A by ligation), pCMV 3xFLAG-ATG2A (Addgene 250865; Previously published3), pLVX Neo TagBFP-LC3B (Addgene 250866; This study; pLVX puro TagBFP-LC3B3 and pLVX-EGFP-IRES-Neo (Addgene #128660) were cut with EcoRI and BamHI, following which the TagBFP-LC3B fragment was ligated into the pLVX Neo backbone), pDESt47 Sar1-GFP (Addgene #67409), SAR1B-mOrange2 (Addgene #166899), SAR1BH79G2-mOrange2 (Addgene #166900), SAR1BH79G-3xFLAG (Addgene 250867; This Study; A BamHI-3xFLAG-NotI fragment was generated by oligo annealing, following which it was ligated into the SAR1BH79G-mOrange2 backbone that was cut by the same restriction enzymes), SAR1B-BFP (Addgene 250872; This Study; A BamHI-BFP-NotI fragment was generated by PCR, following which it was ligated into the SAR1B-mOrange2 backbone that was cut by the same restriction enzymes), SAR1BH79G-BFP (Addgene 250873; This Study; A BamHI-BFP-NotI fragment was generated by PCR, following which it was ligated into the SAR1BH79G-mOrange2 backbone that was cut by the same restriction enzymes), mRuby-SEC23A (Addgene 36158), EGFP-SEC23A (Addgene #66609), pEGFP-SEC24A (Addgene 250868; This Study; a XhoI-SEC24A-PacI fragment was generated by PCR from HA-SEC24A which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), pEGFP-SEC24B (Addgene 250869; This Study; a XhoI-SEC24B-PacI fragment was generated by PCR from HA-SEC24B which was a generous gift from the Kundu lab. The fragment was ligated into the backbone of pEGFP-SEC24D, which was cut by the same restriction enzymes), EGFP-SEC24C (Addgene 250870; This Study; a AgeI-EGFP-XhoI fragment was generated by PCR and was ligated into the backbone of pEYFP-SEC24C (Addgene #66608) which was cut by the same restriction enzymes), pEGFP-SEC24D (Addgene #32678), GFP-SEC31A (Addgene #42124), GFP-SEC16A (Addgene #36155), ARFGAP1-TagRFP (This Study; ARFGAP1-TagRFP was a gift from the Rothman lab. A frame shift and a point mutation were corrected by site directed mutagenesis), ARFGAP1-GFP (Gift from the Antonny Lab), pMRX IB HaloTag7-mGFP (Addgene # 184903), pMRX No HaloTag7-mGFP-LC3B (Addgene # 184901), pLVX GFP (Addgene 250871; This Study; a fragment containing EGFP was generated from pLVX GFP-ATG2A and was ligated into pLVX puro (Clonetech, 632159) that was cut with EcoRI using Gibson Assembly), GFP-RAB1A (Gift from the Rothman Lab), SNAP-RAB1A (Gift from the Rothman Lab), pGEX2T GST-GFP-RAB1A (Addgene 250877; this study; a GFP-RAB1A fragment was generated from the GFP-RAB1A plasmid from the Rothman lab and inserted in to the pGEX2T backbone), and pGEX2T GST-GFP-GL1 (Addgene 2508740).

Antibodies for immunoblotting were used at 1:1000 and include the following: anti-ATG2A (Cell Signaling Technology, 15011S), anti-p62 (BD BioSciences, 610832), anti-GAPDH (Abcam, ab9484), anti-LC3B (Cell Signaling Technology, 3868S), anti-RB1CC1 (Thermo Fisher Scientific, 172501AP), anti-RAB1A (Cell Signaling Technology, D3X9S), anti-RAB1B (Proteintech, 17824-1-AP), anti-ARFGAP1 (Proteintech, 13571-1-AP), anti-FLAG (Sigma-Aldrich, F1804-200UG), anti-GFP (Cell Signaling Technology, 2956S), anti-mCherry (Invitrogen, PA5-34974), ECL anti-rabbit IgG horseradish peroxidase-linked (GE Healthcare, NA934V), and ECL anti-mouse IgG horseradish peroxidase-linked (GE Healthcare, NA931V).

Antibodies for immunofluorescence were used at 1:500 and include the following: anti-Biotin (Rockland, 200-301098S), anti-GFP nanobody (Massive Photonics), anti-ARFGAP1 (Proteintech, 13571-1-AP;), anti-SEC31A (BD Biosciences, 611280), anti-SEC24C (Cell Signaling Technology, 14676S), anti-MIA3 (Tango1; Sigma, HPA055922), anti-TFG (Abcam, ab156866), anti-COPI (CMIA10; made by Rothman Lab), anti-LC3B (MBL, PM036), anti-p62 (BD BioSciences, 610832), anti-SEC16A (Abcam, ab70722), anti-SEC24D (Cell Signaling Technology, 14687S), anti-GM130 (BD Biosciences, 610823), anti-ERGIC53 (Abcam, EPR6979), Alexa Fluor 405 goat anti-rabbit IgG (Thermo Fisher Scientific, A31556), Alexa Fluor 405 goat anti-mouse IgG (Thermo Fisher Scientific, A31553), Alexa Fluor 594 goat anti-rabbit IgG (Thermo Fisher Scientific, A11037), Alexa Fluor 647 goat anti-rabbit IgG (Thermo Fisher Scientific, A21245), and Alexa Fluor 647 goat anti-mouse IgG (Thermo Fisher Scientific, A21236).

The following reagents were purchased for FLASH-PAINT: Cy3B-modified DNA oligonucleotides were custom-ordered from IDT. Sodium chloride 5 M (cat: AM9759) were obtained from Ambion. Ultrapure water (cat: 10977-015) was purchased from Invitrogen. µ-Slide 8-well chambers (cat: 80807) were purchased from ibidi. Methanol (cat: 9070-05) was purchased from J.T. Baker. Glycerol (cat: G5516-500ml), protocatechuate 3,4-dioxygenase pseudomonas (PCD) (cat: P8279), 3,4-dihydroxybenzoic acid (PCA) (cat: 37580-25G-F) and (+−)-6-hydroxy-2,5,7,8-tetra-methylchromane-2-carboxylic acid (Trolox) (cat: 238813-5 G) were ordered from Sigma. 1× Phosphate Buffered Saline (PBS) pH 7.2 (cat: 10010-023) was purchased from Gibco. Paraformaldehyde (cat: 15710) were obtained from Electron Microscopy Sciences. Bovine serum albumin (cat: 001-000-162) was ordered from Jackson ImmunoResearch. Triton X-100 (cat: T8787-50ML) was purchased from Sigma-Aldrich. DNA labeled Nanobodies were obtained from Massive Photonics.

SiControl, FIP200, RAB1A, RAB1B, and ARFGAP1 were acquired from Horizon Discovery (D-001810-10-05, L-0082383-00-0005, L-008958-01-0005, L-013321-02-005; Horizon Discovery).

Cell culture

HEK293 cells (ATCC, CRL_1573), H4 (CVCL_1239), and HeLa (CVCL_0030) cells were cultured at 37°C and 5% CO2 in DMEM (Thermo Fisher Scientific, 11965092) supplemented with 10% FBS (Thermo Fisher Scientific, 10438062) and 1% penicillin-streptomycin (Thermo Fisher Scientific, 15140122). Cells were starved by incubation in Earle’s Balanced Salt Solution (Thermo Fisher Scientific, 24010043) for biochemical assays or in Hank’s Balanced Salt Solution (Thermo Fisher Scientific, 14025092) for live cell imaging assays and treated with 0.1 µM bafilomycin A1 (Enzo, BML-CM110-0100) for 2 h.

Lentivirus production and transduction

HEK293 cells were seeded into a 10-cm plate. At 70% confluence, cells were transfected with 4.3 µg psPAX2 (Addgene #12260), 0.43 µg pCMV-VSV-G (Addgene, #8454), and 4.3 µg target plasmid using 20 μl Lipofectamine 3000 (Thermo Fisher Scientific, L3000008). DNA was added to 500 μl Opti-MEM (Thermo Fisher Scientific, 31985070) and Lipofectamine 3000 to 500 μl Opti-MEM in separate 1.5 ml microcentrifuge tubes. The tubes were then mixed and left for 10 min before adding the mixture dropwise to cells in fresh DMEM. After overnight incubation at 37°C, the medium was replaced with fresh DMEM. Cell medium was then collected after 24-48 h and filtered with a 0.45-µm syringe filter (Pall, 4184). Generated virus was either used immediately or was aliquoted and stored at −80°C. HEK293 cells to be transduced were seeded into a 6-well plate. Undiluted virus was added to cells and incubated at 37°C for 24 h. The medium was replaced with fresh DMEM and left for another 24 h. The cells were then treated with 2 µg/ml puromycin (Clontech, 631306), 2 mg/ml Geneticin (Thermo Fisher Scientific, 10131035), or 10 µg/ml Blasticidin (Gibco, A11139-03) for at least 2 wk. For the case of pMRX No HaloTag7-mGFP-LC3B, which lacked a selection marker, the cells were transduced and subsequently sorted by flow assisted cell sorting to normalize GFP expression levels.

Proximity labeling

Proximity labeling was performed as previously described27. ATG2 DKO cells stably expressing APEX2-EGFP-ATG2A were cultured for two weeks in media containing heavy, medium, or light isotopes of arginine and lysine (Cambridge Isotope Laboratories, CLM-2265-H-PK, DLM-2640-PK, CNLM-291-H-PK; Sigma-Aldrich, 608033). All conditions were expanded to 3×15cm plates prior to proximity labeling. Heavy labeled cells were starved in EBSS and treated with 100 nM Bafilomycin A1 (Enzo, BML-CM110-0100) for 2 hrs prior to labeling. All conditions were incubated with 500 μM Biotin Phenol (Biotinyl Tyramide; Sigma-Aldrich, SML2135) for 30 min at 37°C. The heavy and medium isotope conditions were treated with 1 mM H2O2 (Sigma-Aldrich, H1009-100ML) for 1 min at room temperature. All conditions were then aspirated and scraped in quencher solution containing 10 mM sodium azide (Sigma-Aldrich, S2002-100G), 10 mM sodium ascorbate (Sigma-Aldrich, A4034-100G) and 5 mM Trolox (Sigma-Aldrich, 238813-5G) in DPBS. The cells were then spun down at 77xg for 5 min at 4°C and were resuspended in quencher solution. This process was repeated for a total of three washes. The cells were then lysed in RIPA buffer (50 mM Tris, 150 mM NaCl, 0.1% SDS, 0.5% sodium deoxycholate (Sigma-Aldrich, D6750-10G), 1% NP40 (Roche, 11754599001) and 1x protease inhibitor (Roche, 11836170001)) for 30 min prior to centrifugation for 10 min at 18,407xg at 4°C. The supernatants were collected and protein concentration was assessed via Bradford assay. Heavy, medium and light cells were mixed at a 1:1:1 ratio for a combined mass of 3-4 mg of protein. The protein mixtures were subsequently incubated with streptavidin magnetic beads (Pierce, 88817) for 2 hrs at 4°C after which they were washed twice with RIPA lysis buffer, once with 1 M KCl, once with 0.1 M Na2CO3, once with 2 M urea in 10 mM Tris (pH 8.0) and twice with RIPA lysis buffer for each sample. All washes were performed at 4°C. The proteins were eluted off of the beads by boiling the samples in 2x LDS sample buffer (Invitrogen, NP0007). The eluted fraction was run on a gel (Invitrogen, NP0341) for a 10 minutes. The resulting gel plug was stained (Invitrogen, 46-5034) and excised and sent to the Yale Keck facility for mass spectrometry analysis.

Mass spectrometry

MS was performed at Yale Mass Spectrometry & Proteomics Resource of the W.M. Keck Foundation Biotechnology Resource Laboratory. Samples were analyzed in biological triplicates. The data were normalized to internal controls and total spectral counts. For differential analysis, 10E-04 was added to all values prior to log2-transformation, and an unpaired Student t-test was performed. Differentially abundant proteins were considered those with an absolute fold change > 1.5 and p <0.05.

FLASH-PAINT Imaging

Cells were fixed with 4% PFA (Electron Microscopy Services, 15710-S) for 30 min. After three washes in 1x PBS for five minutes each, cells were blocked and permeabilized with 3% BSA and 0.25% Triton X-100 at room temperature for 1 h. Next, cells were incubated with the anti-ARFGAP1 antibody, anti-SEC31A antibody, and the GFP-Nanobody in 3% BSA and 0.1% Triton X-100 at 4 °C overnight. Additionally, all other primary antibodies were pre incubated with the corresponding nanobodies (Tables S1) at 4 °C overnight. The next day, after four washes (30 s, 60 s, 2× 5 min) cells were incubated with the nanobodies corresponding to ARFGAP1 antibody and SEC31A antibody for ∼2 h at room temperature. Next, unlabeled excess secondary nanobodies (to block unlabeled epitopes) were added to pre-incubation antibody-nanobody mixes at room temperature for 5 min. Next, the cells were incubated with the pooled antibody-nanobody mix for ∼2.5 h at room temperature. After four washes (30 s, 60 s, 2× 5 min), the sample was post-fixed with 3% PFA and 0.1% GA for 10 min. Finally, samples were washed three times with 1× PBS for 5 min each before adding the imaging solution.

The imaging solution consisted of 1x PBS and 500 mM NaCl and was supplemented with 1× Trolox, (100× Trolox: 100 mg Trolox, 430 μL 100 % Methanol, 345 μL 1M NaOH in 3.2 mL H2O), 1× PCA (40× PCA: 154 mg PCA, 10 mL water and NaOH were mixed and pH was adjusted 9.0), and 1× PCD (100× PCD: 9.3 mg PCD, 13.3 mL of buffer (100 mM Tris-HCl pH 8, 50 mM KCl, 1 mM EDTA, 50 % Glycerol)). All three stock buffers were frozen and stored at -20 °C.

Fluorescence imaging was carried out on an inverted Nikon Eclipse Ti2 microscope (Nikon Instruments) with the Perfect Focus System, attached to a Andor Dragonfly unit. The Dragonfly was used in the BTIRF mode, applying an objective-type TIRF configuration with an oil-immersion objective (Nikon Instruments, Apo SR TIRF 60×, NA 1.49, Oil). As excitation laser, a 561 nm (1W nominal) was used. The beam was coupled into a multimode fiber going through the Andor Borealis unit reshaping the beam from a gaussian profile to a homogenous flat top. As dichroic mirror a CR-DFLY-DMQD-01 was used. Fluorescence light was spectrally filtered with an emission filter (TR-DFLY-F600-050) and imaged on a scientific complementary metal oxide semiconductor (sCMOS) camera (Sona 4BV6X, Andor Technologies) without further magnification, resulting in an effective pixel size of 108 nm. The power at the objective lens was ∼10 % of the power set at the laser.

Imaging was carried out using the corresponding Imager and Adapter (Table S2, S3, S5) in imaging buffer. 25,000 frames were acquired at 25 ms exposure time. The readout bandwidth was set to 540 MHz. Laser power (@561 nm) was set to 80 mW (measured before the back focal plane (BFP) of the objective), corresponding to ∼1.8 kW/cm2 at the sample plane. After imaging the sample was subsequently washed three times with 200 µL each with 1× PBS (on the microscope) followed by an incubation with the corresponding eraser (Table S4; at 20 nM) for ∼ 5 min. This process was repeated iteratively for ten sequential rounds (Table S5).

DNA-PAINT data was reconstructed, postprocessed (drift correction and alignment of imaging rounds) and rendered with the Picasso package52.

Focused ion beam scanning electron microscopy and correlative light electron microscopy

For focused ion beam scanning electron microscopy (FIBSEM) CLEM, ATG2 DKO HEK293 cells overexpressing GFP-ATG2A were plated on 35-mm MatTek dish (MatTek, P35G-1.5-14-CGRD). The cells were starved for two hours prior to pre-fixation in 4% PFA + 0.25% glutaraldehyde and three washes in 1x PBS. Regions of interest by fluorescence light microscopy were selected and their coordinates on the dish were identified using phase contrast. Cells were further fixed with 2.5% glutaraldehyde in 0.1 M sodium cacodylate buffer, postfixed in 2% OsO4 and 1.5% K4Fe(CN)6 (Sigma-Aldrich) in 0.1 M sodium cacodylate buffer, en bloc stained with 2% aqueous uranyl acetate, dehydrated, embedded in Embed 812, and polymerized at 60°C for 48 hrs. Cells of interest were relocated based on the pre-recorded coordinates. Epon block was glued onto the SEM sample mounting aluminum stub and platinum en bloc coating on the sample surface was carried out with the sputter coater (Ted Pella, Inc.). The cell of interest was relocated under SEM imaging based on pre-recorded coordinates and FIB-SEM imaged in a Crossbeam 550 FIB-SEM workstation operating under SmartSEM (Carl Zeiss Microscopy GmbH) and Atlas 5 engine (Fibics, Inc.). The imaging resolution was set at 7 nm/pixel in the X, Y axis with milling being performed at 7 nm/step along the Z axis to achieve an isotropic resolution of 7 nm/voxel. Images were aligned and exported with Atlas 5 (Fibics, Inc.), further processed and segmented with DragonFly Pro software (Object Research Systems [ORS], Inc.).

Expi293 culture and transfection

3xFLAG-ATG2A was expressed and purified from Expi293 suspension cells. Cells were cultured at 37°C, 8% CO2 in vented PETG flasks (Thermo Scientific, Nalgene; 4115-0500) with 125 rpm rotation. Prior to transfection, cells were grown to 3-5 x 106 cells/mL then diluted to 2.5×106 cells/mL and allowed to grow for 18-24 hours. Upon reaching 4.5-5.5 x 106 cells/mL, the culture was diluted to 3 x 106 cells/mL with fresh media. Cells were transfected with 100 μg plasmid per 100 mL of cultured cells. For every 100 mL of cells, 100 μg DNA was diluted in 5 mL OptiMEM. 300 µL of 1 mg/mL polyethylenimine linear 25k (Polyscience, 23966) was diluted into 5 mL OptiMEM in a separate tube. Both tubes were incubated for 5 minutes at room temperature then combined and further incubated for 30 minutes. The DNA/PEI mixture was then added dropwise into Expi293 cultures. 12-18 hours after transfection, protein expression was enhanced using 700 μL of 500 mM valproic acid. Cells were then harvested after 48-72 hours.

3xFLAG-ATG2A purification

Cell pellets were resuspended in 10 mL purification buffer A (500 mM NaCl, 20 mM HEPES, pH: 8.0, 10% glycerol, 1 mM TCEP) for every 100 mL of original culture volume. Purification buffer A was supplemented with 1x EDTA-free protease inhibitor cocktail (Roche; 11873580001). Cells were then lysed by subjecting pellet resuspensions to 6 freeze/thaw cycles in liquid nitrogen The lysates were then clarified by centrifugation at 18,500 x G, and supernatants were transferred to 150 μL of anti-Flag M2 agarose resin per 100 mL of initial culture volume (prewashed with 10 mL purification buffer A). Lysates were incubated with beads under mild agitation for 2 hours at 4 °C. Lysates and beads were then passed through a gravity filtration column and washed with 4 CVs of purification buffer. Beads were then incubated with 2.5 mM ATP and 5 mM Mg2+ in purification buffer overnight at 4 °C. Columns were then washed with 2 CVs of purification buffer, followed by elution with 50 μL of 1 mg/mL 3x-Flag (DYKDDDDK) peptide (ApexBio Technology; A6001) per 100 mL of initial culture volume. Protein concentration was determined against a BSA standard curve by Coomassie staining prior to subsequent assays. Proteins were flash frozen in liquid nitrogen for storage at -80 °C.

Recombinant protein expression and purification from E. Coli

For all in vitro experiments, GFP-GL1 refers to protein expressed and purified from constructs ending at the reactive glycine (G116) of GL1, as this construct was previously used for lipidation experiments in the absence of ATG4 proteins. In summary, human GFP-GL1 and GFP-RAB1A were cloned in the pGEX-2T GST vector. The proteins were expressed in BL21-Gold (DE3) competent cells (Agilent Technologies; 230132). The cells were cultured in 4 liters of Luria Bertani broth (LB) medium with 50 mg/ml carbenicillin (C1389; Sigma-Aldrich) and induced with 0.5 mM isopropyl β-D-thiogalactopyranoside (IPTG). Bacterial pellets were treated with EDTA-free protease inhibitor tablets in thrombin buffer (20 mM Tris pH 7.5, 100 mM NaCl, 5 mM MgCl2, 2 mM CaCl2, 1 mM DTT). The cells were broken in a cell disrupter, and the lysate was incubated with pre-washed glutathione agarose beads (G4501; Sigma-Aldrich) for 3 h at 4°C. The beads were washed three times and then incubated with thrombin (10 μl of thrombin [T6884; Sigma-Aldrich], 500 μl of thrombin buffer, 0.5 μl of DTT, 500 μl of beads) to cut the proteins from GST tags overnight at 4°C. Purified proteins were stored in 20% glycerol at −80°C.

Lysis, gel electrophoresis, and immunoblotting

Cells were collected by scraping in PBS. The cells were centrifuged at 500 g for 3 min at 4°C, and lysed in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor cocktail (Sigma-Aldrich, 11873580001) on ice for five minutes. Lysates were collected after centrifugation at 16,000 g for 10 min at 4°C. Total protein concentration was determined by a Bradford assay (BioRad, 5000006). 50-100 μM DTT (American Bio, AB00490-00010) and 1x LDS (Invitrogen, NP0007) loading buffer were added to each sample, which were then electrophoresed on precast Bis-Tris gels (Thermo Fisher Scientific, NP0341BOX or NP0322BOX). Gels were subsequently transferred to Immobilon-FL PVDF membranes (Sigma-Aldrich, IPFL00010) at 30-35 V for 90 min.

The membranes were blocked with 5% BSA (Sigma-Aldrich, A9647) in PBS-T (PBS containing 0.1% Tween 20; Thermo Scientific, CAS 9005-64-5) for 1 h at room temperature and incubated with primary antibody (diluted 1:1,000 in PBS-T containing 5% BSA [Sigma-Aldrich, A9647] and 0.03% sodium azide (Sigma-Aldrich, S2002)) overnight at 4°C. Membranes were washed three times in PBS-T before incubation with the secondary antibody which was diluted 1:5,000 in PBS-T containing 0.5% Omniblock (American Bio, AB10109-00100) for 1 h. Membranes were then washed three times in PBST and treated with SuperSignal West Femto substrate (Thermo Fisher Scientific, 34096) for 5 min before imaging with the Bio-Rad VersaDoc imaging system.

Immunoprecipitation

For immunoprecipitation experiments, 15 cm plates were seeded and transfected upon reaching 80-90% confluency. The next day, the cells were collected by scraping, washed 1x in PBS, and lysed as previously described. Following quantification of protein levels by a Bradford assay, equal amounts of protein were added to pre-washed GFP-Trap beads (Proteintech, gtma-20). The volume was increased to 500 uL and the bead/lysate slurry was incubated at 4°C for two hours while rotating. The beads were pelleted using a magnetic rack and washed five times in lysis buffer (10 mM Tris pH 7.5, 150 mM NaCl, 0.5 mM EDTA, 0.5% NP-40) with EDTA-free protease inhibitor. The target proteins were eluted by boiling the beads for 10 min in 2x LDS loading buffer.

Densitometry and quantification

Densitometry quantifications of immunoblots and in-gel fluorescence images were performed using ImageJ software. For LC3B and p62 clearance analysis (Fig. S1 A-C), the band intensity of LC3B or p62 for each condition was normalized against the GAPDH intensity of that lane. To normalize each replicate, each value was divided by the mean value of the entire replicate. This normalization technique was also employed in the co-IP experiments. Statistical significance was determined by two-way ANOVA followed by a Tukey multiple comparison test to compare cell types. To quantify the co-IP reactions, the intensity of the 3xFLAG-ATG2A IP band was ratioed against the intensity of its corresponding input band. Following normalization, statistical significance was determined by a one (Fig. S3 A-D) or two-way (Fig. 2E-F) ANOVA followed by a Dunnett’s multiple-comparisons test comparing each sample with the control lane. For the HaloTag flux assay (Fig. 2A, 2B, 3A, 3B), the Halotag flux rate was calculated as the ratio of the free Halotag band divided by the unprocessed Halotag band. Statistical significance was determined by a one (Fig. 2 A and B) or two-way (Fig 3 A and B) ANOVA followed by a Dunnett multiple comparison test to comparing each sample to the control lane. For one way ANOVA tests, parametric tests were used for all comparisons. For two way ANOVA tests, the Geisser-Greenhouse correction was used as equal variability was not assumed. All data are n = 3 or greater and represent biological replicates. All data were plotted with mean ± SD, and all significance values represent adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001. All blots were adjusted in contrast for better visualization of the bands in the figures.

LC3B flux analysis

WT, ATG2 DKO and ATG2 DKO + APEX2-GFP-ATG2A HEK293 cells were cultured in 6-well plates to 70-90% confluence. At the beginning of the assay, the media was replaced with fresh DMEM, DMEM containing 100 nM bafilomycin A1, or EBSS containing 100 nM bafilomycin A1. After a 2 hour incubation, cells were harvested by scraping, washed with 1x PBS, and lysed. 20 μg of protein from each sample were analyzed by SDS-PAGE and immunoblotting for ATG2A, p62, GAPDH, and LC3B.

HaloTag flux assay

HEK293 WT stably expressing either HaloTag7-mGFP or HaloTag7-mGFP-LC3B were cultured in 6-well plates to 60-70% confluence prior to transfection with either SAR1B/SAR1BH79G-mOrange2 for one day (Fig. 3A) or siRNA for two days (Fig. 2A). SAR1B transfection was performed using Lipofectamine 3000 (Thermofisher Scientific, L3000015) according to the manufacturer’s protocol. SiRNA transfection was performed using Dharmafect (Horizon Discovery, T-2001-01) according to the manufacturer’s protocol. The cells were incubated with 100 nm TMR-conjugated HaloTag ligand (Promega, G8251) for 30 min. Following two washes with 1x PBS, the cells were starved in EBSS for four hours. The cells were then collected and lysed. Following SDS-PAGE separation, in gel fluorescence was immediately measured with a ChemiDoc MP Imaging System (BioRad). Images were taken at 30 s and 100 s exposures. The gel was then transferred and immunoblotted against mCherry and GAPDH (Fig 3A) or FIP200, RAB1A, RAB1B, ARFGAP1, GAPDH, and LC3B (Fig 2A).

GFP-mCherry-LC3B Flux analysis by FACS

H4 cells and H4 cells with a GFP-mCherry-LC3B heterozygous knock-in tag were seeded in a six well plate. Following a three day siRNA knock down, the cells were subject to a four hour starvation period in HBSS. They were subsequently trypsinized, pelleted, resuspended in HBSS, and passed through a filter prior to FACS data acquisition. The cell population was gated to ensure a single, live cell population. The ratio between the GFP and mCherry intensity was plotted as a histogram using FlowJo 10 and the ratio between the average values of each replicate was plotted using Prism 10. Three biological replicates were combined and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values represented adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

Immunofluorescence

Cells seeded on coverslips were fixed in 4% PFA (Electron Microscopy Sciences, 15710), washed three times with PBS, and permeabilized in PBS containing 0.1% Saponin (Sigma-Aldrich, 47036) and 3% BSA. Incubation with primary antibodies was done at 4C for 18-24 hrs and secondary antibodies at room temperature for 1 hr at indicated concentrations (see Reagents). After each antibody incubation, cells were washed three times in PBS containing 0.1% Saponin and 3% BSA. The coverslips were mounted on precleaned microscope slides with Fluoromont-G mounting reagent (Southern Biotech, 0100-01). Images were acquired on an inverted Zeiss LSM 880 laser scanning confocal microscope with AiryScan, using Zen acquisition software (Fig S1D) or a Nikon CSU-W1 SoRa 60x oil objective, using the Nikon Elements software (Fig. 1C, 3C-F, S3A, S3C, 4A, 4C, 4E, 4F). All images were processed using ImageJ software. Images are displayed either as maximum intensity projections or single confocal slices as stated in the figure legends.

To assess the enrichment factor of early secretory proteins around the ATG2 DKO compartment (Fig. 4D), the phagophore channel which was denoted by either LC3B or p62 was masked and duplicated. Following outlier removal, one masked image was dilated and the other eroded. Following an XOR function, the resulting signal formed a ring around the ATG2 DKO compartment and the early secretory protein channel(s) were measured using this ROI. Separately, the early secretory protein channel was masked using a manual threshold and measured. The enrichment factor was calculated as (Meancompartment-Mincell)/(Meancell-Mincell). Three biological replicates were combined and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance represent considering adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.

Live cell imaging

Cells were plated on poly-d-lysine coated MatTek dishes (MatTek, P35GC-1.5-14-C). The following day the cells were transfected using Lipofectamine 3000 (Thermo Fisher Scientific, L3000015) according to the manufacturer’s protocol for 18-24 hrs. The cells were starved in HBSS (Thermo Fisher Scientific, 14025092) for 2-4 h and imaged in the same solution at 37°C. All images were analyzed using ImageJ and are displayed as maximum intensity projections.

For phagophore quantification (Fig. 1 C and D), each channel was masked using a manual threshold and outliers were removed. The ATG2A and LC3B channels were merged using the AND function and quantified as the total number of phagophores. The phagophore channel was then merged using the AND function with the transfected gene and quantified as the number of positive phagophores. Three biological replicates were combined and the ratio of positive over total was plotted and statistically assessed using a one-way ANOVA followed by a Dunnett multiple comparison test. All data were plotted with mean ± SD, and all significance values represent adjusted P value based on grouped analyses in Prism 10 (GraphPad). Asterisks indicate significance: *, P < 0.05; **, P < 0.01; ***, P < 0.001; ****, P < 0.0001.